# Affine processes {#affine}

The following code produces a figure showing simulated path of an AR(1) process $x_t$, and $x_t^2$.

## A simple Gaussian AR(1) generates nonlinear transforms such as squared states

This example shows how a simple Gaussian AR(1) generates nonlinear transforms such as squared states.

```{r fig_QTSM1, fig.width=8, fig.height=4}
# Purpose: This example shows how a simple Gaussian AR(1) generates nonlinear transforms such as
# squared states
set.seed(123)

T <- 200

phi <- .9
mu <- 2 * (1 - phi)
sigma <- 1

# Start from the unconditional mean.
x.t <- mu / (1 - phi)
x <- NULL

for (t in 1:T) {
  x.t <- mu + phi * x.t + sigma * rnorm(1)
  x <- c(x, x.t)
}

par(mfrow = c(1, 2), plt = c(.15, .95, .1, .85))
plot(x, type = "l", xlab = "", ylab = "", las = 1,
     main = expression(paste("(a) ", x[t], sep = "")), lwd = 2)
grid()
plot(x^2, type = "l", xlab = "", ylab = "", las = 1,
     main = expression(paste("(b) ", {x[t]}^2, sep = "")), lwd = 2)
grid()
```

The following code produces a figure illustrating ARG processes.

## Standard ARG and ARG0 processes

This example contrasts a standard ARG process with its ARG0 counterpart.

```{r fig_ARG1, fig.width=8, fig.height=4}
# Purpose: This example contrasts a standard ARG process with its ARG0 counterpart
library(DTAM)
set.seed(123)

# Simulate two ARG specifications and compare their paths.
par(mfrow = c(1, 2), plt = c(.15, .95, .1, .85))

W <- simul_ARG(200, mu = .05, nu = 1, rho = .95, alpha = 0)
plot(W, type = "l", xlab = "", ylab = "", las = 1,
     main = expression(paste("(a) ", ARG, " process", sep = "")), lwd = 2)
grid()

W <- simul_ARG(200, mu = .5, nu = 0, rho = .95, alpha = .1)
plot(W, type = "l", xlab = "", ylab = "", las = 1,
     main = expression(paste("(b) ", ARG[0], " process", sep = "")), lwd = 2)
grid()
```

The following code simulates a Poisson process.

## The jump-like dynamics generated by a compound Poisson process

This example illustrates the jump-like dynamics generated by a compound Poisson process.

```{r fig_Poisson1, fig.width=8, fig.height=3}
# Purpose: This example illustrates the jump-like dynamics generated by a compound Poisson
# process
library(DTAM)
set.seed(123)

# Simulate and plot a compound Poisson path.
par(mfrow = c(1, 1), plt = c(.1, .95, .15, .95))
W <- simul_compound_poisson(200, Gamma = 1, Pi = 0.95, lambda = .1)
plot(W, type = "l", las = 1, lwd = 2, xlab = "", ylab = "")
grid()
```

The following code simulate a recursive Gaussian-ARG process.

## Recursive Gaussian-ARG model with crisis and GDP-growth components

This example builds a two-factor recursive Gaussian-ARG system with a crisis component and GDP-growth component.

```{r fig_bivarARGGaussian, fig.width=8, fig.height=5}
# Purpose: This example builds a two-factor recursive Gaussian-ARG system with a crisis
# component and GDP-growth component
library(DTAM)
set.seed(123)

mu <- .01
beta <- .8 / mu
alpha <- .05
rho <- .6
sigma <- .005
kappa <- .03 * (1 - rho)

# Calibrate the recursive Gaussian-ARG system.
# Save the parameterization for the next examples.
model <- list(
  mu = mu, beta = beta, alpha = alpha, rho = rho, sigma = sigma, kappa = kappa
)

par(mfrow = c(2, 1), plt = c(.1, .95, .15, .7))
nb_periods <- 400

# Simulate and plot the nonnegative crisis factor.
W1 <- simul_ARG(nb_periods, mu = mu, nu = 0, rho = beta * mu, alpha = alpha)
plot(W1, type = "l", xlab = "", ylab = "", las = 1,
     main = expression(paste("(a) ", w[1*','*t], " process", sep = "")), lwd = 2)
grid()

mean_W1 <- alpha * mu / (1 - beta * mu)
mean_W2 <- (-mean_W1 + kappa) / (1 - rho)

# Simulate GDP growth conditional on the crisis factor.
W2 <- mean_W2
for (t in 2:nb_periods) {
  W2 <- c(W2, -W1[t] + kappa + rho * W2[t - 1] + sigma * rnorm(1))
}
plot(W2, type = "l", xlab = "", ylab = "", las = 1,
     main = expression(paste("(b) ", w[2*','*t], " process", sep = "")), lwd = 2)
grid()
```

The following code uses the previous model to compute forecasts of GDP (GDP growth is $w_{2,t}$ of the previous process, and $w_{1,t}$ is a crisis regime).

## GDP forecasts under different crisis states

This example shows how the previous model can be used to produce horizon-by-horizon GDP forecasts under different crisis states.

```{r fig_GDP, fig.width=8, fig.height=4}
# Purpose: This example shows how the previous model can be used to produce horizon-by-horizon
# GDP forecasts under different crisis states
psi <- function(U, model) {
  # U is a 2 x K matrix collecting the transform arguments.
  mu <- model$mu
  beta <- model$beta
  alpha <- model$alpha
  rho <- model$rho
  sigma <- model$sigma
  kappa <- model$kappa

  u1 <- U[1, ]
  u2 <- U[2, ]

  a1 <- mu * (u1 - u2) / (1 - mu * (u1 - u2)) * beta
  a2 <- u2 * rho
  b <- mu * (u1 - u2) / (1 - mu * (u1 - u2)) * alpha + u2 * kappa + u2^2 * sigma^2 / 2

  list(a = rbind(a1, a2), b = b)
}

# Three cases: current crisis only, current + future crises, and no future crises.
U <- matrix(c(0, 1,
              -10000, 1,
              -10000, 0), 2, 3)

# Set the forecast horizon and the two initial states.
H <- 10
X <- matrix(c(0, mean_W2,
              .01, mean_W2), 2, 2)
AB <- reverse_MHLT(psi, U, H = H, psi.parameterization = model)

# Recover the three conditional GDP forecasts from the transform coefficients.
GDP_fcst_1 <- c(AB$B[, 1, ]) + t(matrix(AB$A[, 1, ], 2, H)) %*% X[, 1]
GDP_fcst_2 <- c(AB$B[, 1, ]) + t(matrix(AB$A[, 1, ], 2, H)) %*% X[, 2]
GDP_fcst_3 <- c(AB$B[, 2, ] - AB$B[, 3, ]) +
  t(matrix(AB$A[, 2, ] - AB$A[, 3, ], 2, H)) %*% X[, 1]

# Plot the forecast profiles.
par(plt = c(.15, .95, .2, .95))
plot(exp(GDP_fcst_1), type = "l", lwd = 2, las = 1,
     xlab = "horizon h", ylab = "Expected GDP", ylim = c(.97, 1.35))
lines(exp(GDP_fcst_2), col = "grey", lwd = 2)
lines(exp(GDP_fcst_3), lwd = 2, lty = 2)
grid()

legend("topleft",
       legend = c(
         expression(paste(E, "[", Y[t+h], "|",
                          w[1*','*t], " = 0, ",
                          w[2*','*t], "=0.03]", sep = "")),
         expression(paste(E, "[", Y[t+h], "|",
                          w[1*','*t], " = 0.01, ",
                          w[2*','*t], "=0.03]", sep = "")),
         expression(paste(E, "[", Y[t+h], "|",
                          w[1*','*t+1:t+h], " = 0, ",
                          w[2*','*t], "=0.03]", sep = ""))
       ),
       lty = c(1, 1, 2),
       col = c("black", "grey", "black"),
       lwd = c(2, 2, 2), bty = "n")
```

The following code displays the conditional pdf of GDP at different horizons.

## Laplace transforms can recover the full conditional distribution of future GDP growth

This example illustrates how Laplace transforms can recover the full conditional distribution of future GDP growth.

```{r fig_CDFGP, fig.width=8, fig.height=4}
# Purpose: This example illustrates how Laplace transforms can recover the full conditional
# distribution of future GDP growth
library(DTAM)

# Here the variable of interest is cumulated GDP growth.
# If indic_no_crisis is TRUE, future crisis realizations are shut down.
psi_logGDP <- function(u, model) {
  H <- model$H
  X <- model$X

  if (model$indic_no_crisis) {
    U <- rbind(-10000, 1 * c(u))
  } else {
    U <- rbind(0, 1 * c(u))
  }
  AB <- reverse_MHLT(psi, U, H = H, psi.parameterization = model)

  if (model$indic_no_crisis) {
    U <- matrix(c(-10000, 0), 2, 1)
    AB_1w0 <- reverse_MHLT(psi, U, H = H, psi.parameterization = model)
  }

  if (model$indic_no_crisis) {
    psi <- exp(t(AB$A[, , H] - AB_1w0$A[, , H]) %*% X + c(AB$B[, , H] - AB_1w0$B[, , H]))
  } else {
    psi <- exp(t(AB$A[, , H]) %*% X + c(AB$B[, , H]))
  }

  psi
}

model$H <- 10
gamma <- seq(-.1, .45, by = .001)

# Recover densities under alternative crisis assumptions.
model$indic_no_crisis <- FALSE
model$X <- X[, 1]
pdf.dy_with_crisis_w1ini0 <- make_pdf(model, gamma, psi_logGDP)

model$indic_no_crisis <- FALSE
model$X <- X[, 2]
pdf.dy_with_crisis_w1iniNot0 <- make_pdf(model, gamma, psi_logGDP)

model$indic_no_crisis <- TRUE
pdf.dy_no_crisis <- make_pdf(model, gamma, psi_logGDP)

# Plot the three conditional distributions.
par(mfrow = c(1, 1))
par(plt = c(.15, .95, .2, .95))
plot(gamma, pdf.dy_with_crisis_w1ini0, type = "l", main = "dy", lwd = 2,
     ylim = c(0, max(pdf.dy_with_crisis_w1ini0,
                     pdf.dy_with_crisis_w1iniNot0,
                     pdf.dy_no_crisis)),
     xlab = expression(paste(log(Y[t+h] / Y[t]), sep = "")), ylab = "", las = 1)
grid()
lines(gamma, pdf.dy_with_crisis_w1iniNot0, col = "grey", lwd = 2, lty = 1)
lines(gamma, pdf.dy_no_crisis, col = "black", lwd = 2, lty = 2)
abline(h = 0, col = "grey")

legend("topleft",
       legend = c(
         expression(paste("log(", Y[t+h], "/", Y[t], ") | ",
                          w[1*','*t], " = 0, ",
                          w[2*','*t], "=0.03", sep = "")),
         expression(paste("log(", Y[t+h], "/", Y[t], ") | ",
                          w[1*','*t], " = 0.01, ",
                          w[2*','*t], "=0.03", sep = "")),
         expression(paste("log(", Y[t+h], "/", Y[t], ") | ",
                          w[1*','*t+1:t+h], " = 0, ",
                          w[2*','*t], "=0.03", sep = ""))
       ),
       lty = c(1, 1, 2),
       col = c("black", "grey", "black"),
       lwd = c(2, 2, 2), bty = "n")
```
