# Term structure of interest rates: definitions, instruments, and yield-curve conventions {#IntroTSM}

## Several standard yield objects can be derived from a single discount curve

This example shows how several standard yield objects can be derived from a single discount curve.

```{r fig_conventions, fig.width=7, fig.height=4.5}
# Purpose: This example shows how several standard yield objects can be derived from a single
# discount curve

# Nelson-Siegel-Svensson primitives and derived yield curves.
nss_yield <- function(h, theta) {
  b0 <- theta["beta0"]; b1 <- theta["beta1"]; b2 <- theta["beta2"]; b3 <- theta["beta3"]
  l1 <- theta["lambda1"]; l2 <- theta["lambda2"]

  x1 <- (1 - exp(-l1 * h)) / (l1 * h)
  x2 <- x1 - exp(-l1 * h)
  x3 <- (1 - exp(-l2 * h)) / (l2 * h) - exp(-l2 * h)

  b0 + b1 * x1 + b2 * x2 + b3 * x3
}

disc <- function(h, theta) exp(-h * nss_yield(h, theta))

# Curves derived from the discount function.
zc_yield <- function(h, theta) nss_yield(h, theta)

# simple forward rate between h and h+dh under continuous compounding
fwd_rate <- function(h, dh, theta) {
  -(log(disc(h + dh, theta)) - log(disc(h, theta))) / dh
}

# Par yield for maturity h with p coupons per year (cashflows at i/p)
par_yield <- function(h, theta, p) {
  n <- round(h * p)              # number of coupon periods
  times <- (1:n) / p
  df <- disc(times, theta)
  dfT <- disc(h, theta)
  p * (1 - dfT) / sum(df)
}

# Swap rate: same as par yield when fixed leg schedule matches
swap_rate <- function(h, theta, p) {
  par_yield(h, theta, p)
}

# YTM of a coupon bond priced off the same discount curve
ytm_from_price <- function(price, c, h, p) {
  n <- round(h * p)
  times <- (1:n) / p

  f <- function(y) {
    dfy <- exp(-y * times)
    sum((c / p) * dfy) + exp(-y * h) - price
  }
  uniroot(f, c(-0.05, 0.30))$root
}

coupon_bond_price <- function(c, h, theta, p) {
  n <- round(h * p)
  times <- (1:n) / p
  sum((c / p) * disc(times, theta)) + disc(h, theta)
}

# Choose parameters and maturity grids.

theta <- c(
  beta0 = 0.03, beta1 = -0.015, beta2 = 0.02, beta3 = -0.005,
  lambda1 = 1.2, lambda2 = 3.5
)

p <- 2  # number of payments per period -- 2 = semiannual payments

Hmax <- 20

# Dense grid for ZC + forward (exists for all maturities)
h_dense <- seq(0.25, Hmax, by = 0.25)

# Payment-date grid for par/swap (must have h*p integer)
h_pay <- seq(1/p, Hmax, by = 1/p)

# Compute the implied curves.
y_zc  <- sapply(h_dense, zc_yield, theta = theta)
y_fwd <- sapply(h_dense, fwd_rate,  dh = 1/12, theta = theta)  # 1-month fwd

y_par  <- sapply(h_pay, par_yield,  theta = theta, p = p)
y_swap <- sapply(h_pay, swap_rate, theta = theta, p = p)

# Coupon-bond yields for two different coupon levels.

mats <- seq(1,Hmax,by=1)
coupons <- rep(.1,length(mats))
ytm_pts <- mapply(function(h, cpn) {
  pr <- coupon_bond_price(cpn, h, theta, p)
  ytm_from_price(pr, cpn, h, p)
}, mats, coupons)

mats2 <- seq(1,Hmax,by=1)
coupons2 <- rep(0.01,length(mats2))
ytm_pts2 <- mapply(function(h, cpn) {
  pr <- coupon_bond_price(cpn, h, theta, p)
  ytm_from_price(pr, cpn, h, p)
}, mats2, coupons2)



# Plot the resulting term-structure objects together.

par(plt=c(.14,.98,.18,.97))

plot(h_dense, 100*y_zc, type = "l", lwd = 2,las=1,
     ylim = 100 * c(.025, .034),
     xlab = "Maturity h (in years)", ylab = "Annualized %",
     main = "Multiple yield curves from one discount curve")
grid()

lines(h_dense, 100*y_fwd, lwd = 3, lty = 3)
lines(h_pay,  100*y_par,  lwd = 2, lty = 2)
# lines(h_pay,  y_swap, lwd = 2, lty = 4)

points(mats,  100*ytm_pts,  pch = 16, cex = 0.8)
points(mats2, 100*ytm_pts2, pch = 15, cex = 0.8,col="darkgrey")

legend("bottomright",
       legend = c("Zero-coupon yield", "Forward (1m)",
                  "Par yield", "Coupon-bond YTM (cpn = 10%)",
                  "Coupon-bond YTM (cpn = 1%)"),
       col=c(rep("black",4),"darkgrey"),
       lty = c(1, 3, 2, NA, NA),
       lwd = c(2, 3, 2, NA, NA),
       pch = c(NA, NA, NA, 16, 15), bty = "n")


```


## Clean and dirty prices

This example illustrates the distinction between clean and dirty prices for a coupon bond.

```{r fig_clean_dirty_prices, fig.width=7, fig.height=3.8}
# Purpose: This example illustrates the distinction between clean and dirty prices for a coupon
# bond.

# Bond and discount-curve assumptions.
r_flat <- 0.03
coupon_rate <- 0.04
coupon_freq <- 2
coupon_period <- 1 / coupon_freq
coupon_payment <- coupon_rate / coupon_freq
maturity <- 5

# Monthly settlement dates over the life of the bond.
settlement_grid <- seq(0, maturity - 1 / 12, by = 1 / 12)
coupon_dates <- seq(coupon_period, maturity, by = coupon_period)

dirty_price <- clean_price <- accrued_interest <- numeric(length(settlement_grid))

for (i in seq_along(settlement_grid)) {
  s <- settlement_grid[i]
  remaining_dates <- coupon_dates[coupon_dates > s]
  cash_flows <- rep(coupon_payment, length(remaining_dates))
  cash_flows[remaining_dates == maturity] <- cash_flows[remaining_dates == maturity] + 1

  dirty_price[i] <- sum(cash_flows * exp(-r_flat * (remaining_dates - s)))

  previous_coupon <- ifelse(any(coupon_dates <= s), max(coupon_dates[coupon_dates <= s]), 0)
  next_coupon <- min(coupon_dates[coupon_dates > s])
  accrued_interest[i] <- coupon_payment * (s - previous_coupon) / (next_coupon - previous_coupon)
  clean_price[i] <- dirty_price[i] - accrued_interest[i]
}

# Plot clean and dirty prices.
op <- par(mar = c(4.1, 4.1, 0.8, 0.8), las = 1, cex.axis = 1.05)
ylim <- range(c(dirty_price, clean_price))

plot(
  settlement_grid, dirty_price,
  type = "l", lwd = 2, col = "black",
  xlab = "time since issuance (years)", ylab = "price",
  ylim = ylim
)
grid()
abline(v = coupon_dates[coupon_dates < maturity], col = "grey85", lty = 3)
lines(settlement_grid, clean_price, col = "grey55", lwd = 2, lty = 2)
legend(
  "bottomleft",
  legend = c("dirty price", "clean price"),
  col = c("black", "grey55"),
  lty = c(1, 2), lwd = 2, bty = "n",
  cex = 1
)
par(op)
```


## Treasury issuance and outstanding debt by maturity

This example documents the maturity composition of Treasury issuance and the stock of debt outstanding.

```{r SIFMA, fig.width = 10, fig.height = 5}
# Purpose: This example documents the maturity composition of Treasury issuance and the stock of
# debt outstanding
library(tidyverse)
library(zoo)
library(patchwork)
library(readxl)

theme_baseR <- function() {
  theme_classic(base_size = 12) +
    theme(
      panel.grid.major.y = element_line(color = "grey60", linewidth = 0.4,
                                        linetype = 3),
      panel.grid.minor.y = element_line(color = "grey60", linewidth = 0.2,
                                        linetype = 3),
      panel.grid.major.x = element_line(color = "grey60", linewidth = 0.3,
                                        linetype = 3),
      panel.grid.minor.x = element_line(color = "grey60", linewidth = 0.3,
                                        linetype = 3),
      axis.ticks = element_line(linewidth = 0.6),
      panel.grid = element_line(),
      strip.text = element_text(face = "bold", size = 14),
      strip.background = element_blank(),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.6),
      plot.margin = margin(10, 10, 10, 10)
    )
}

# Load and reshape Treasury issuance data.
ta_us_treasury_sifma <- read_excel("data/ta-us-treasury-sifma.xlsx", sheet = "Issuance", skip = 5)
ta_us_treasury_sifma <- ta_us_treasury_sifma %>%
  as_tibble() %>%
  dplyr::select(-...5, -...9, -...13)

names(ta_us_treasury_sifma) <-
  c("date", "Bills_Gross Issuance", "Bills_Gross Retirement", "Bills_net",
    "Notes_Gross Issuance", "Notes_Gross Retirement", "Notes_net",
    "Bonds_Gross Issuance", "Bonds_Gross Retirement", "Bonds_net",
    "Total_Gross Issuance", "Total_Gross Retirement", "Total_net"
  )

ta_us_treasury_sifma <- ta_us_treasury_sifma %>%
  mutate(date = as.yearmon(as.Date(date))) %>%
  pivot_longer(cols = 2:13)

ta_us_treasury_sifma <- ta_us_treasury_sifma %>%
  mutate(type = str_extract(name, "^[^_]+"),
         variable = str_replace(name, "^[^_]*_", ""),
         year = year(date))

# Load and reshape the outstanding-debt data.
ta_us_treasury_sifma_OS <- read_excel(
  "data/ta-us-treasury-sifma.xlsx",
  sheet = "Outstanding",
  skip = 3
)
ta_us_treasury_sifma_OS <- (ta_us_treasury_sifma_OS[1:25,]) %>%
  mutate(TIPS = as.numeric(str_replace(TIPS, "-", "NA")),
         FRN = as.numeric(str_replace(FRN, "-", "NA"))) %>%
  rename(year = ...1)

ta_us_treasury_sifma_OS <- ta_us_treasury_sifma_OS %>%
  mutate(variable = "outstanding") %>%
  pivot_longer(cols = 2:7) %>%
  rename(type = name)


# Plot net issuance by instrument type.
p <- ggplot(ta_us_treasury_sifma %>%
              filter(variable == "net", type != "Total") %>%
              group_by(type) %>%
              mutate(y = rollmean(value,k=12, align = "right", fill = NA),
                     type = factor(type, levels = c("Bonds", "Notes", "Bills"))) ,
            aes(x = date, y = y, col = type, group = type, fill = type)) +
  geom_area() +
  theme_baseR() +
  labs(x = "", y = "USD Billions", title = "(a) Net issuance") +
  scale_color_manual(values = rev(c("black", "grey60", "grey80"))) +
  scale_fill_manual(values = rev(c("black", "grey60", "grey80"))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 11),
        plot.title = element_text(hjust = .5, face="bold"),
        legend.position = "bottom",
        legend.title = element_blank()
  )

ta_us_treasury_sifma_OS <- ta_us_treasury_sifma_OS %>%
  mutate(date = as.Date(paste0(year,"-01-01")))

# Plot outstanding amounts and composition.
q <- ggplot() +
  geom_area(data = ta_us_treasury_sifma_OS %>% filter(type != "Total") %>%
              mutate(type = factor(type, levels = rev(c("TIPS","FRN","Bonds",  "Bills", "Notes")))),
            aes(x = date, y = value, col = type, group = type, fill = type),
            position = "fill", alpha = .6
  ) +
  geom_line(data = ta_us_treasury_sifma_OS %>% filter(type == "Total") %>%
              mutate(type = "Total (right axis)"),
            aes(x = date, y = value/max(value), col = type, group = type, fill = type),
            lwd = 2)+
  theme_baseR() +
  scale_y_continuous(
    name = "Share",      # left axis
    sec.axis = sec_axis(
      ~ . * 20973.13,
      name = "Total amount"
    )
  ) +
  scale_color_manual(values = c( "grey90", "grey70", "grey50", "grey30","black", "black"))+
  scale_fill_manual(values = c( "grey90", "grey70", "grey50", "grey30","black", "black")) +
  labs(x = "", title = "(b) Outstanding amounts") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 11),
        plot.title = element_text(hjust = .5, face="bold"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.box = "horizontal"
  ) +
  guides(color = guide_legend(nrow = 1))

p + q

```
