# Shadow-rate framework and extensions {#ShadowRate}


## Shadow-rate and unconstrained Gaussian term structures in the simple Wu-Xia setting

This example compares shadow-rate and unconstrained Gaussian term structures in the simple Wu-Xia setting.

```{r fig_SR1, fig.width=7, fig.height=3.5}
# Purpose: This example compares shadow-rate and unconstrained Gaussian term structures in the
# simple Wu-Xia setting
library(DTAM)

# Specify the Gaussian shadow-rate state dynamics.
rho <- .96
Phi <- as.matrix(rho)
mu <- matrix(0.03*(1-rho),1,1)
Sigma <- (1-rho^2) * (.02)^2 * matrix(1)
model <- list(mu = mu, Phi = Phi, Sigma = Sigma,
              n_w=1)

# Simulate the latent shadow rate.
TT <- 250
set.seed(1234)
s <- simul_GaussianVAR(model,nb.sim = TT)

par(mfrow=c(1,1))
par(plt=c(.1,.95,.13,.95))
plot(s,type="l",col="white",xlab="",ylab="",las=1)
# grid()
abline(h=0,col="grey",lty=2)
lines(s,lwd=2,col="lightgrey")

delta0 <- 0
delta1 <- matrix(1,1,1)

# Compute shadow-rate forward rates with the Wu-Xia approximation.
H <- 40 # maximum maturity
res <- compute_F_Shadow_Gaussian(s,
                                 psi.parameterization = model,
                                 ell_bar=0,
                                 b=delta0,
                                 a=delta1,
                                 c=matrix(0,1,1),
                                 H = H-1,
                                 indic_WuXia = TRUE)
f <- cbind(pmax(0,s),res$F)
y <- t(apply(f,1,cumsum))/t(matrix(1:H,H,TT))

AB_GATSM <- compute_AB_classical(xi0=0,xi1=matrix(1),H = H,
                                 psi = psi_GaussianVAR,psi.parameterization = model)
y_noSR <- t(matrix(AB_GATSM$b,H,TT)) + s %*% matrix(AB_GATSM$a,ncol=H)

all_t <- seq(1,TT, by=H/2)
for(t in all_t){
  lines((t+1):(t+H),y[t,],type="l",col="black",lwd=3)
  lines((t+1):(t+H),y_noSR[t,],type="l",col="darkgrey",lwd=3,lty=3)
}


legend("bottomright",
       c(expression(paste("Shadow rate (",s[t],")",sep="")),
         expression(paste("Nominal rates (",r[t*','*h],")",sep="")),
         expression(paste("Nominal rates if ",r[t]," = ",s[t],sep=""))),
       lty=c(1,1,3),
       lwd=c(2,3,3),
       pt.cex = 1,
       col= c("grey","black","darkgrey"),
       seg.len = 3)

```

## The lower bound changes conditional rate volatility across maturities

This example shows how the lower bound changes conditional rate volatility across maturities.

```{r fig_SR1_vol, fig.width=7, fig.height=3.5}
# Purpose: This example shows how the lower bound changes conditional rate volatility across
# maturities
library(expm)

# Set the maturities and helper function used in the truncated-normal moments.
all_h <- c(1,3,6)

g <- function(x){
  x*pnorm(x) + dnorm(x)
}

max_y <- 100*res$all_sigma_n[max(all_h)]

# Compute and plot the conditional volatility of observed rates.
par(mfrow=c(1,1))
par(plt=c(.15,.95,.15,.95))
plot(s,type="l",
     ylim=c(0,max_y),
     col="white",xlab="",ylab="Conditional volat., in percent",
     las=1)

grid()
count <- 0
for(h in rev(all_h)){
  count <- count + 1
  
  mu1t <- res$all_b_bar_n[h] + s %*% t(model$Phi%^%h) %*% delta1
  sigma1 <- res$all_sigma_n[h] 
  ratio <- mu1t/sigma1
  var_rt <- (mu1t^2 + sigma1^2)*pnorm(ratio) +
    mu1t*sigma1*dnorm(ratio) - sigma1^2*g(ratio)^2
  vol_rt <- 100*sqrt(var_rt)
  
  lines(vol_rt,lwd=2,lty=length(all_h) + 1 - count)
}


# Add maturity labels.
legend_labels <- paste("h =", all_h)

legend("bottomright",legend = legend_labels,
       lty=1:length(all_h),
       lwd=2,
       col= "black",
       seg.len = 3)

```


## The shadow-rate idea in a small New Keynesian model with nominal and real curves

This example embeds the shadow-rate idea in a small New Keynesian model with nominal and real curves.

```{r fig_SR2, fig.width=7, fig.height=4, results='hide'}
# Purpose: This example embeds the shadow-rate idea in a small New Keynesian model with nominal
# and real curves
library(DTAM)

r_star <- .03
pi_star <- .02
rho <- .5
phi <- .8
beta <- .2
alpha <- 1.5
gamma <- 0.5
sigma_r  <- .005
sigma_pi <- .005
sigma_y  <- .005
kappa <- .2

A0 <- diag(4)
A0[1,3] <- (rho-1)*alpha
A0[1,4] <- (rho-1)*gamma
# A0[4,3] <- -beta

A0_1 <- solve(A0)

mu <- A0_1 %*% matrix(c((1-rho)*(r_star - alpha*pi_star),0,0,
                        beta*(r_star-pi_star)),ncol=1)
Phi <- A0_1 %*% matrix(c(rho,1,0,-beta,
                         0,0,0,0,
                         0,0,1,beta,
                         0,0,kappa,phi+beta*kappa),4,4)
Sigma12 <- A0_1 %*% matrix(c(sigma_r,0,0,0,
                             0,0,sigma_pi,0,
                             0,0,0,sigma_y),4,3)
Sigma <- Sigma12 %*% t(Sigma12)

model <- list(mu = mu,Phi = Phi,
              Sigma12 = Sigma12,Sigma=Sigma,
              n_w=4)

print(abs(eigen(Phi)$values))

TT <- 250
set.seed(123)
W <- simul_GaussianVAR(model,nb.sim = TT)

par(mfrow=c(1,1))
par(plt=c(.1,.98,.1,.98))
plot(W[,1],
     ylim=c(min(W)-.01,max(W)+.03),
     xlab="",ylab="",
     # ylim=c(-.1,.2),
     col="white",las=1)
abline(h=0,col="black",lty=3)
# grid()
lines(W[,1],col="lightgrey",lwd=2,lty=1)
lines(W[,3],col="lightgrey",lwd=2,lty=2)
lines(W[,4],col="lightgrey",lwd=2,lty=3)


# Compute shadow-rate forward rates with the Wu-Xia approximation.
H <- 40 # maximum maturity

res <- compute_F_Shadow_Gaussian(W,
                                 psi.parameterization = model,
                                 ell_bar=0,
                                 b=0,
                                 a=matrix(c(0,1,0,0),4,1),
                                 c=matrix(c(0,0,1,0),4,1),
                                 H = H)
f_nominal <- res$F_c_equal_0
f_real <- res$F
y_nominal <- t(apply(f_nominal,1,cumsum))/t(matrix(1:H,H,TT))
y_real    <- t(apply(f_real,1,cumsum))/t(matrix(1:H,H,TT))

AB_GATSM <- 
  compute_AB_classical(xi0=0,
                       xi1=matrix(c(1,0,0,0),ncol=1),
                       kappa0 = 0, kappa1 = matrix(c(0,0,-1,0),ncol=1),
                       H = H,
                       psi = psi_GaussianVAR,psi.parameterization = model)
y_noSR_nominal <- t(matrix(AB_GATSM$b[,1,],H,TT)) + W %*% matrix(AB_GATSM$a[,1,],ncol=H)
y_noSR_real    <- t(matrix(AB_GATSM$b[,2,],H,TT)) + W %*% matrix(AB_GATSM$a[,2,],ncol=H)


all_t <- seq(1,TT,by=H+10)

for(t in all_t){
  lines((t+1):(t+H),y_nominal[t,],type="l",col="black",lwd=3)
  lines((t+1):(t+H),y_real[t,],type="l",col="darkgrey",lwd=3)
  lines((t+1):(t+H),y_noSR_nominal[t,],type="l",col="black",lwd=3,lty=3)
  lines((t+1):(t+H),y_noSR_real[t,],type="l",col="darkgrey",lwd=3,lty=3)
}

legend("topright",
       c(expression(paste("Shadow rate (",s[t],")",sep="")),
         expression(paste("Inflation (",pi[t],")",sep="")),
         expression(paste("Output gap (",y[t],")",sep=""))),
       lty=c(1,2,3),
       lwd=c(2,2,2),
       pt.cex = 1,
       col= "lightgrey",
       seg.len = 3)

legend("topleft",
       c(expression(paste("Nominal rates (",r[t*','*h],")",sep="")),
         expression(paste("Nominal rates if ",r[t]," = ",s[t],sep="")),
         expression(paste("Real rates (",tilde(r)[t*','*h],")",sep="")),
         expression(paste("Real rates if ",r[t]," = ",s[t],sep=""))),
       lty=c(1,3,1,3),
       lwd=c(3,3,3,3),
       pt.cex = 1,
       col= c("black","black","darkgrey","darkgrey"),
       seg.len = 3)

```


## Shadow-rate pricing of defaultable bond spreads

This example uses a shadow-rate pricing device to illustrate defaultable bond spreads in a structural credit model.

```{r fig_SR3, fig.width=8, fig.height=6.5}
# Purpose: This example uses a shadow-rate pricing device to illustrate defaultable bond spreads
# in a structural credit model
library(DTAM)

# Specify output gap's (z_t) dynamics:
rho <- .8
sigma <- .02

gamma <- 2    # RRA coefficient
delta <- .995 # subjective time discount
mu    <- .015  # average consumption growth
kappa <- 2    # sensitivity of default intensity to output gap


Phi <- matrix(c(rho,1,0,0),2,2)
Sigma12 <- matrix(c(sigma,0),2,1)
Sigma <- Sigma12 %*% t(Sigma12)
model <- list(mu = matrix(0,2,1), Phi = Phi, 
              Sigma = Sigma,Sigma12 = Sigma12,
              n_w=2)

# Simulate z:
TT <- 100
set.seed(123)
W <- simul_GaussianVAR(model,nb.sim = TT)

# plot(W[,1],type="l")

mu_c1 <- matrix(c(1,-1),2,1)
dc <- mu + W %*% mu_c1

# plot(cumsum(dc),type="l")

H <- 10 # maximum maturity
res_default <- compute_F_Shadow_Gaussian(W,
                                         psi.parameterization = model,
                                         ell_bar=0,
                                         b=0,
                                         a=matrix(c(-kappa,0),2,1),
                                         c=matrix(c(-gamma,gamma),2,1),
                                         H = H)
res_rf <- compute_F_Shadow_Gaussian(W,
                                    psi.parameterization = model,
                                    ell_bar=0,
                                    b=0,
                                    a=.00001*matrix(c(-kappa,0),2,1),
                                    c=matrix(c(-gamma,gamma),2,1),
                                    H = H)
f_PD <- res_default$F_c_equal_0
f_bond <- res_default$F
f_rf   <- res_rf$F
y_PD   <- t(apply(f_PD,1,cumsum))/t(matrix(1:H,H,TT))
y_rf   <- -log(delta) + gamma*mu + t(apply(f_rf,1,cumsum))/t(matrix(1:H,H,TT))
y_bond <- -log(delta) + gamma*mu + t(apply(f_bond,1,cumsum))/t(matrix(1:H,H,TT))

par(mfrow=c(2,1))
par(plt=c(.11,.98,.13,.8))

plot(W[,1],col="white",xlab="",ylab="",las=1,
     main = expression(paste("(a) Output gap (",z[t],")",sep="")))
grid()
abline(h=0,col="grey")
lines(W[,1],lwd=2)

plot(y_bond[,H],type="l",col="white",las=1,
     ylim=c(0,.13),xlab="",ylab="",
     main=expression(paste("(b) Bond yields",sep="")))
grid()
lines(y_rf[,H],col="darkgrey",lwd=2)
lines(y_bond[,H],col="black",lwd=2,lty=1)
lines(y_bond[,H]-y_rf[,H],col="black",lwd=2,lty=3)

legend("topright",
       c(expression(paste("Risk-free rate (",tilde(r)[t*','*h],")",sep="")),
         expression(paste("Defaultable-bond rate (",{r^d}[t*','*h],")",sep="")),
         expression(paste("Credit spread (",{r^d}[t*','*h]-tilde(r)[t*','*h],")",sep=""))),
       lty=c(1,1),
       lwd=c(2,2),
       pt.cex = 1,
       col= c("darkgrey","black"),
       seg.len = 3)

# Compute the cumulative probability of default from cumulative intensities.
PD <- 1 - exp(-t(apply(f_PD,1,cumsum)))

```

## Shadow default intensity and credit spreads in a fiscal-limits setting

This example studies shadow default intensity and credit spreads in a fiscal-limits setting.

```{r fig_SR4, fig.width=8, fig.height=8}
# Purpose: This example studies shadow default intensity and credit spreads in a fiscal-limits
# setting
library(DTAM)

r_star <- .03
ell_star <- 1.1
d_star <- .9
sigma_r   <- .005
sigma_ell <- .04
sigma_d   <- .04
rho_r   <- .9
rho_ell <- .98
rho_d   <- .98
kappa <- .5

H <- 5 # bond maturity


# Specify state vector's dynamics (w_t = (r_t,ell_t,d_t,r_{t-1})'):
Phi <- diag(c(rho_r,rho_ell,rho_d,0))
Phi[4,1] <- 1
mu <- (diag(4) - Phi) %*% c(r_star,
                            ell_star,d_star,r_star)
Sigma12 <- diag(c(sigma_r,sigma_ell,sigma_d,0))
Sigma <- Sigma12 %*% t(Sigma12)
model <- list(mu = mu, Phi = Phi, 
              Sigma = Sigma,Sigma12 = Sigma12,
              n_w=4)

# Simulate model:
TT <- 100
set.seed(123)
W <- simul_GaussianVAR(model,nb.sim = TT)

par(mfrow=c(2,2))
par(plt=c(.15,.95,.15,.85))

plot(W[,1],type="l",col="white",ylim=c(.01,.07),
     xlab="",ylab="",las=1,
     main=expression(paste("(a) Short-term interest rate (",r[t],")",sep="")))
grid()
lines(W[,1],lwd=2,col="black")

plot(W[,2],ylim=c(0,2),col="white",xlab="",ylab="",las=1,
     main=expression(paste("(b) Debt and fiscal limit",sep="")))
grid()
lines(W[,2],type="l",ylim=c(0,2),col="black",lwd=2,lty=1)
lines(W[,3],type="l",ylim=c(0,2),col="black",lwd=2,lty=3)
lines(W[,2]-W[,3],type="l",ylim=c(0,2),col="darkgrey",lwd=2,lty=1)

legend("topright",
       c(expression(paste("Debt-to-GDP (",d[t],")",sep="")),
         expression(paste("Fiscal limit (",l[t],")",sep="")),
         expression(paste("Fiscal space (",l[t]," - ",d[t],")",sep=""))),
       lty=c(3,1,1),
       lwd=c(2,2,2),
       pt.cex = 1,
       col= c("black","black","darkgrey"),
       seg.len = 3)

a <- matrix(c(0,-kappa,kappa,0),4,1)
c <- matrix(c(0,0,0,-1),4,1)

res_default <- compute_F_Shadow_Gaussian(W,
                                         psi.parameterization = model,
                                         ell_bar=0,
                                         b=0,
                                         a=a,
                                         c=c,
                                         H = H)
res_rf <- compute_F_Shadow_Gaussian(W,
                                    psi.parameterization = model,
                                    ell_bar=0,
                                    b=0,
                                    a=.00001*a,
                                    c=c,
                                    H = H)

# res_default <- compute_F_Shadow_affine(W,psi_GaussianVAR,
#                                          psi.parameterization = model,
#                                          ell_bar=0,
#                                          b=0,
#                                          a=.5*matrix(c(0,-1,1,1),4,1),
#                                          c=matrix(c(-1,0,0,0),4,1),
#                                          H = H)
# res_rf <- compute_F_Shadow_affine(W,psi_GaussianVAR,
#                                     psi.parameterization = model,
#                                     ell_bar=0,
#                                     b=0,
#                                     a=.00001*matrix(c(0,-1,1,1),4,1),
#                                     c=matrix(c(-1,0,0,0),4,1),
#                                     H = H)


f_PD <- res_default$F_c_equal_0
f_bond <- res_default$F
f_rf   <- res_rf$F
y_PD   <- t(apply(f_PD,1,cumsum))/t(matrix(1:H,H,TT))
y_rf   <- t(apply(f_rf,1,cumsum))/t(matrix(1:H,H,TT))
y_bond <- t(apply(f_bond,1,cumsum))/t(matrix(1:H,H,TT))

plot(y_bond[,H],type="l",col="white",las=1,
     ylim=c(0.01,.1),xlab="",ylab="",
     main=expression(paste("(c) Long-term interest rates",sep="")))
grid()
lines(y_rf[,H],col="darkgrey",lwd=2)
lines(y_bond[,H],col="black",lwd=2,lty=3)

legend("topright",
       c(expression(paste("Risk-free (",tilde(r)[t*','*h],")",sep="")),
         expression(paste("Defaultable (",{r^d}[t*','*h],")",sep=""))),
       lty=c(1,3),
       lwd=c(2,2),
       pt.cex = 1,
       col= c("darkgrey","black"),
       seg.len = 3)

# Check pricing formula ------------------------------------------------------
# n <- length(model$mu)
# nb.replic <- 10000
# all_P <- NULL
# for(i in 1:nb.replic){
#   sum_c <- 0
#   sum_max <- 0
#   w <- W
#   for(h in 1:H){
#     w <- t(matrix(model$mu,n,TT)) + w %*% t(model$Phi) +
#       matrix(rnorm(TT*n),TT,n) %*% t(model$Sigma12)
#     sum_c <- sum_c + w %*% c
#     sum_max <- sum_max + pmax(0,w %*% a)
#   }
#   P <- exp(sum_c - sum_max)
#   all_P <- cbind(all_P,P)
# }
# P_simul <- apply(all_P,1,mean)
# y_simul <- -log(P_simul)/H
# lines(y_simul,col="red",lwd=2,lty=3)
# -----------------------------------------------------------------------------

# Compute the cumulative probability of default from cumulative intensities.
PD <- 1 - exp(-t(apply(f_PD,1,cumsum)))

plot(10000*(y_bond[,H]-y_rf[,H]),type="l",col="white",las=1,
     ylim=10000*c(0,.045),xlab="",ylab="",
     main=expression(paste("(d) Credit spread (in bps)",sep="")))
grid()
lines(10000*y_PD[,H],col="grey",lwd=2)
lines(10000*(y_bond[,H]-y_rf[,H]),col="black",lwd=2,lty=3)

legend("topleft",
       c(expression(paste("Credit spread (",{r^d}[t*','*h]," - ",tilde(r)[t*','*h],")",sep="")),
         expression(paste("Average default intensity (annualized)",sep=""))),
       lty=c(3,1),
       lwd=c(2,2),
       pt.cex = 1,
       col= c("black","darkgrey"),
       seg.len = 3)


```

## Non-Gaussian shadow-rate model with jump-like factors

This example illustrates a non-Gaussian shadow-rate model with jump-like factors and compares constrained and unconstrained curves.

```{r fig_SR5, fig.width=8, fig.height=8}
# Purpose: This example illustrates a non-Gaussian shadow-rate model with jump-like factors and
# compares constrained and unconstrained curves
set.seed(7)

model <- (list(
  alpha = matrix(c(.0,.0),ncol=1),
  nu = matrix(c(.2,.2),ncol=1),
  mu = matrix(c(.01,.01),ncol=1),
  beta = NaN,
  n_w = 2
))
model$beta <- .95*diag(1/c(model$mu))

TT = 150

simX <- t(simul_VARG(model,nb_periods = TT-1))
DeltaNplus <- rpois(n = TT,simX[,1])
DeltaNminus <- rpois(n = TT,simX[,2])
Nplus  <- cumsum(DeltaNplus)
Nminus <- cumsum(DeltaNminus)

par(mfrow=c(2,1))
par(plt=c(.1,.9,.15,.85))

scaled_simX <- simX * 10

maxN <- max(Nplus,Nminus)+2

plot(1:TT,rep(0,TT),las=1,col="white",
     ylim=c(0,maxN),xlab="",ylab="",
     main=expression(paste("(a) Simulated state variables",sep="")))
grid()
Nvalues <- 1:maxN
ylabels_z <- seq(0,maxN/10,by=.2)
axis(side=4,at=10*ylabels_z,labels=ylabels_z,las=1)
lines(scaled_simX[,1],col="black",lwd=2)
lines(scaled_simX[,2],col="darkgrey",lwd=2)

lines(Nplus,col="black",lwd=3,lty=3)
lines(Nminus,col="darkgrey",lwd=3,lty=3)

legend("topleft",
       c(expression(paste(N[t]^{'+'}," (lhs)",sep="")),
         expression(paste(N[t]^{'-'}," (lhs)",sep="")),
         expression(paste(z[t]^{'+'}," (rhs)",sep="")),
         expression(paste(z[t]^{'-'}," (rhs)",sep=""))),
       lty=c(3,3,1,1),
       lwd=c(3,3,2,2),
       pt.cex = 1,
       col= c("black","darkgrey","black","darkgrey"),
       seg.len = 3)

# Determine state vector
W <- cbind(simX,
           Nplus,Nminus)

# Shadow rate specification:
xi0 <- .0025
xi1 <- matrix(c(0,0,.25/100,-.25/100),ncol=1)

s <- xi0 + W %*% xi1
plot(0*s,col="white",type="l",lwd=2,las=1,ylim=c(-.01,.025),
     xlab="",ylab="",
     main=expression(paste("(b) Term structures of interest rates (w/ and w/o lower bound)")))
grid()
abline(h=0,col="black",lty=2)
lines(s,col="darkgrey",lwd=2)

model$n_w <- 4 # update number of factors.


HH <- 30 # Maximum bond maturity

# Compute bond prices without LB:
res <- compute_AB_classical(xi0, xi1, kappa0 = NaN, kappa1 = NaN,
                            H=HH, psi=psi_VARG_Poisson,
                            psi.parameterization=model)
yields <- matrix(1,TT,1) %*% matrix(res$b,nrow=1) +
  W %*% matrix(res$a,dim(res$a)[1],dim(res$a)[3])

# Compute bond prices without LB:
F_affine <- compute_F_Shadow_affine(W=W,psi_VARG_Poisson,
                                    psi.parameterization=model,
                                    ell_bar=0,b=xi0,a=xi1,c=0*xi1,
                                    H=HH,
                                    eps = 10^(-6), # to compute dG
                                    max_x = 1000,
                                    dx_statio = 1,
                                    min_dx = 1e-06,
                                    nb_x1 = 500)
f_bond <- cbind(pmax(0,s),F_affine$F[,1:(HH-1)])
yields_SR <- t(apply(f_bond,1,cumsum))/t(matrix(1:HH,HH,TT))

all_ts <- seq(5,TT-HH,by=HH+5)
for(t in all_ts){
  lines((t+1):(t+HH),yields[t,],col="black",lwd=3,lty=3)
  lines((t+1):(t+HH),yields_SR[t,],col="black",lwd=3)
}

legend("topleft",
       c(expression(s[t]),
         expression(paste(r[t*','*h]," when ",r[t]," = max(",s[t],",0)",sep="")),
         expression(paste(r[t*','*h]," when ",r[t]," = ",s[t],sep=""))),
       lty=c(1,1,3),
       lwd=c(2,3,3),
       pt.cex = 1,
       col= c("darkgrey","black","black"),
       seg.len = 3)


```



## Shadow rate with ARG stochastic volatility

This example combines a shadow rate with stochastic volatility driven by an ARG factor.

```{r fig_SR6, fig.width=8, fig.height=8}
# Purpose: This example combines a shadow rate with stochastic volatility driven by an ARG
# factor

set.seed(1)

TT <- 200

psi_AR_CH <- function(u,psi.parameterization){
  # Laplace transform of a process defined as follows:
  # s_{t+1} = mu + phi·s_{t} + sqrt(z_t)·eps_{t+1}, with eps_t ~ N(0,1) and
  # z_t ARG(nu,beta,mu_z)
  
  mu  <- psi.parameterization$mu
  phi <- psi.parameterization$phi
  
  nu   <- psi.parameterization$nu
  mu_z <- psi.parameterization$mu_z
  beta <- psi.parameterization$beta
  
  u <- matrix(u,nrow=2)
  k <- dim(u)[2]
  
  u_s <- u[1,]
  u_z <- u[2,]
  
  b <- u_s*mu - nu*log(1 - mu_z*u_z)
  a <- matrix(NaN,2,k)
  a[1,] <- u_s * phi
  a[2,] <- beta*mu_z*u_z/(1 - mu_z*u_z) + u_s^2/2
  
  return(list(a = a, b = b))
}

Es    <- .02
phi   <- .95
phi_z <- .95
mu_z  <- .000015
nu    <- .3
model <- list(mu = Es*(1-phi), phi = phi,
              nu = nu, mu_z = mu_z, beta = phi_z/mu_z,
              n_w = 2)

sim_z <- simul_ARG(nb.sim=TT,
                   mu = model$mu_z,
                   nu = model$nu,
                   rho = phi_z)

Es <- model$mu/(1 - model$phi)
s <- Es
sim_s <- s
for(t in 2:TT){
  s <- model$mu + model$phi * s + sqrt(sim_z[t])*rnorm(1)
  sim_s <- c(sim_s,s)
}
W <- cbind(sim_s,sim_z)
s <- W[,1]

# Shadow rate specification:
xi0 <- 0
xi1 <- matrix(c(1,0),ncol=1)

HH <- 20 # Maximum bond maturity

# Compute bond prices without LB:
res <- compute_AB_classical(xi0, xi1, kappa0 = NaN, kappa1 = NaN,
                            H=HH, psi=psi_AR_CH,
                            psi.parameterization=model)
yields <- matrix(1,TT,1) %*% matrix(res$b,nrow=1) +
  W %*% matrix(res$a,dim(res$a)[1],dim(res$a)[3])

# Compute bond prices with LB:
F_affine <- compute_F_Shadow_affine(W=W,psi_AR_CH,
                                    psi.parameterization=model,
                                    ell_bar=0,b=xi0,a=xi1,c=0*xi1,
                                    H=HH,
                                    eps = 10^(-6), # to compute dG
                                    max_x = 1000,
                                    dx_statio = 1,
                                    min_dx = 1e-06,
                                    nb_x1 = 500)
f_bond <- cbind(pmax(0,s),F_affine$F[,1:(HH-1)])
yields_SR <- t(apply(f_bond,1,cumsum))/t(matrix(1:HH,HH,TT))


par(mfrow=c(2,1))
par(plt=c(.1,.95,.15,.85))
plot(sim_s,type="l",las=1,col="white",xlab="",ylab="",
     main=expression(paste("(a) State variables (",s[t]," and ",z[t],")",sep="")))
grid()
abline(h=0,col="grey")
lines(sim_s,col="grey",lwd=2,lty=1)
lines(3*sqrt(sim_z),col="black",lwd=2)

# Compute conditional variance of r_t:
mu1t <- model$mu + model$phi * sim_s
sigma1 <- sqrt(sim_z)
ratio <- mu1t/sigma1
var_rt <- (mu1t^2 + sigma1^2)*pnorm(ratio) +
  mu1t*sigma1*dnorm(ratio) - sigma1^2*g(ratio)^2
lines(3*sqrt(var_rt),lwd=2,lty=3,col="black")

legend("topleft",
       c(expression(s[t]),
         expression(paste(sqrt(z[t])," (x3)",sep="")),
         expression(paste(sqrt(Var(r[t]))," (x3)",sep=""))
       ),
       lty=c(1,1,3),
       lwd=c(2,2,2),
       pt.cex = 1,
       col= c("grey","black","black"),
       seg.len = 3)


plot(0*s,col="white",type="l",lwd=2,las=1,ylim=c(-.03,.08),
     xlab="",ylab="",
     main=expression(paste("(b) Term structures of interest rates (w/ and w/o lower bound)")))
grid()
abline(h=0,col="grey")
lines(s,col="darkgrey",lwd=2)

all_ts <- seq(1,TT,by=HH+1)
for(t in all_ts){
  lines((t+1):(t+HH),yields[t,],col="black",lwd=3,lty=3)
  lines((t+1):(t+HH),yields_SR[t,],col="black",lwd=3)
}

legend("topleft",
       c(expression(s[t]),
         expression(paste(r[t*','*h]," when ",r[t]," = max(",s[t],",0)",sep="")),
         expression(paste(r[t*','*h]," when ",r[t]," = ",s[t],sep=""))),
       lty=c(1,1,3),
       lwd=c(2,3,3),
       pt.cex = 1,
       col= c("darkgrey","black","black"),
       seg.len = 3)


```

## Affine-transform methods to a simple climate-finance model

This example applies affine-transform methods to a simple climate-finance model.

```{r fig_SR7, fig.width=9, fig.height=4}
# Purpose: This example applies affine-transform methods to a simple climate-finance model

library(DTAM)

time_step_yrs <- 5
HH <- 200 # max maturity used to evaluate inter-temporal utility

model <- list(mu_c = .015*time_step_yrs,
              gamma = 1.45,
              et = .0075*time_step_yrs, # initial growth rate of carbon concentrations
              E0 = 2170,
              Et = 3315,
              delta = 0.95,
              mu_e = 0.001,
              rho_e = .8,
              rho = .99999,
              S = 3,
              alpha = .0018*time_step_yrs*1,
              beta  = .0018*time_step_yrs*1,
              tau_star = 3,
              rho_star = .9,
              stdv_tau_star = .5,
              Ct = 60*10^12, # Global consumption in 2025, in USD
              n_w = 3)

psi_IAM <- function(u, psi.parameterization){
  # This function returns the conditional Laplace transform of
  # the state vector of the macro-climate model.
  # The state vector is w_t = (e_t,tau_t,tau_t*)'.
  
  mu_e  <- psi.parameterization$mu_e
  rho_e <- psi.parameterization$rho_e
  rho   <- psi.parameterization$rho
  S     <- psi.parameterization$S
  alpha <- psi.parameterization$alpha
  beta  <- psi.parameterization$beta
  tau_star   <- psi.parameterization$tau_star
  rho_star   <- psi.parameterization$rho_star
  sigma_star <- sqrt(1-rho_star^2)*psi.parameterization$stdv_tau_star
  
  u <- matrix(u,nrow=3)
  k <- dim(u)[2]
  
  u_e    <- matrix(u[1,],nrow=1)
  u_tau  <- matrix(u[2,],nrow=1)
  u_star <- matrix(u[3,],nrow=1)
  
  a <- matrix(0,3,k)
  a[1,] <- rho_e*u_e/(1 - mu_e*u_e) + S/log(2)*u_tau
  a[2,] <- rho*u_tau
  a[3,] <- rho_star*u_star
  b <- (u_star*tau_star*(1-rho_star) + u_star^2*sigma_star^2/2)
  
  return(list(a = a, b = b))
}

# Compute distribution of the date of net-zero emissions -----------------------
u1 <- matrix(c(-1000,0,0),ncol=1)
res <- reverse_MHLT(psi_IAM,u1=u1,u2=0*u1,H=40,psi.parameterization = model)
A <- matrix(res$A,nrow=3)
B <- c(res$B)
P_net_zero <- exp(B + model$et*A[1,])

# Compute distribution of atmospheric carbon concentration in 2100 -------------
all_h <- c(2060,2080,2100)
step <- .025

values_tau <- seq(1.8,4.5,by=step)
PDF <- matrix(0,length(values_tau),length(all_h))

psi_tau <- function(u,model){
  # This function returns E_t[exp(u·tau_{t+h})]
  # The horizon h is one of the 'model' entries.
  U <- rbind(0,1*c(u),0)
  res <- reverse_MHLT(psi_IAM,
                      u1 = U,u2 = 0*U,
                      H = h,model)
  Xt <- matrix(c(model$et,model$S/log(2)*log(model$Et/model$E0),model$tau_star),ncol=1)
  psi <- exp(res$B[1,1,h] + t(res$A[,,model$h]) %*% Xt)
  return(psi)
}

count_h <- 0
for(h in all_h){
  count_h <- count_h + 1
  model$h <- (h - 2025)/time_step_yrs
  PDF[,count_h] <- make_pdf(model,values.of.variable = values_tau,
                            psi=psi_tau, max_x = 500, min_x = 10^(-5), 
                            nb_x = 1000)/step
}

par(mfrow=c(1,2))
par(plt=c(.12,.95,.2,.85))
plot(2025+time_step_yrs*(1:(length(P_net_zero)-1)),diff(P_net_zero),
     col="white",xlab="Year",ylab="",las=1,
     main=expression(paste("(a) Pdf of date of net-zero emissions (",e[t]," = 0)",sep="")))
grid()
lines(2025+time_step_yrs*(1:(length(P_net_zero)-1)),diff(P_net_zero),
      type="l",lwd=2,xlab="Year",ylab="")

plot(values_tau, PDF[,1], type = "l", col = "white",
     xlab = expression(paste("Temperature (", tau[t], "), in °C", sep = "")),
     ylab = "", las = 1,
     main=expression(paste("(b) Pdf of temperature anomaly (",tau[t],")",sep="")))
grid()
for(i in 1:length(all_h)){
  lines(values_tau,PDF[,i],lty=i,lwd=2)
}

legend("topright",legend=all_h,
       lty=c(1,2,3),
       lwd=2,
       pt.cex = 1,
       col= "black",
       seg.len = 3)
```



The next chunk computes the social cost of carbon in the affine climate-finance model, first with and then without the risk-premium component.

```{r compute_SCC, fig.width=8, fig.height=4}
# Purpose: Compute the social cost of carbon and separate the risk-premium component.

compute_SCC <- function(model,HH=200,
                        indic_compute_riskpremium = FALSE,
                        eps = 10^(-6), # to compute dG
                        max_x = 1000000,
                        dx_statio = 1000,
                        min_dx = 1e-06,
                        nb_x1 = 1000){
  # This function computes the social of carbon (SCC) in the context of a small
  # Integrated Assessment Model (IAM) featuring a form of stochastic tipping point.
  
  W <- t(matrix(c(model$et,
                  model$S/log(2)*log(model$Et/model$E0),
                  model$tau_star),3,2))
  # Use two values (i.e., two rows) to compute dU/dE
  
  # Perturb emissions to approximate the marginal effect on utility.
  epsE <- 1 # Delta E_t, in GtCO2
  W[2,2] <- model$S/log(2)*log((model$Et + epsE)/model$E0)
  
  a <- (model$gamma - 1) * matrix(c(0,model$beta,-model$beta),ncol=1)
  c <- (model$gamma - 1) * matrix(c(0,model$alpha,0),ncol=1)
  # Compute expected discounted marginal utility under the affine model.
  F_affine <- compute_F_Shadow_affine(W=W,psi_IAM,
                                      psi.parameterization=model,
                                      ell_bar=0,
                                      b=0,a=a,c=c,chi=1,
                                      H=HH,eps = eps, # to compute dG
                                      max_x = max_x,dx_statio = dx_statio,
                                      min_dx = min_dx,nb_x1 = nb_x1)
  f_bond <- F_affine$F - log(model$delta) - (1 - model$gamma)*model$mu_c
  aux    <- exp(t(apply(-cbind(0,f_bond),1,cumsum)))
  
  Ct <- model$Ct*time_step_yrs # $tr, 2024
  U <- Ct^(1 - model$gamma)/(1 - model$gamma) * apply(aux,1,sum)
  dU <- (U[2] - U[1])/epsE
  
  SCC <- -Ct^(model$gamma) * dU / 10^9 # /10^9 because eps is in GtCO2
  
  if(indic_compute_riskpremium){
    # Compute the SDF discount factors used to isolate the risk-premium component.
    a <- model$gamma * matrix(c(0,model$beta,-model$beta),ncol=1)
    c <- model$gamma * matrix(c(0,model$alpha,0),ncol=1)
    F_affine_SDF <- compute_F_Shadow_affine(W=W,psi_IAM,
                                            psi.parameterization=model,
                                            ell_bar=0,
                                            b=0,a=a,c=c,chi=1,
                                            H=HH,eps = eps, # to compute dG
                                            max_x = max_x,dx_statio = dx_statio,
                                            min_dx = min_dx,nb_x1 = nb_x1)
    f_bond <- F_affine_SDF$F - log(model$delta) + model$gamma*model$mu_c
    DF <- exp(t(apply(-cbind(0,f_bond),1,cumsum))) # Discount factors
    DF <- DF[1,]
    
    # Compute the consumption response to the emissions perturbation.
    a <- matrix(c(0,model$beta,-model$beta),ncol=1)
    c <- - matrix(c(0,model$alpha,0),ncol=1)
    F_affine_C <- compute_F_Shadow_affine(W=W,psi_IAM,
                                          psi.parameterization=model,
                                          ell_bar=0,
                                          b=0,a=a,c=c,chi=-1,
                                          H=HH,eps = eps, # to compute dG
                                          max_x = max_x,dx_statio = dx_statio,
                                          min_dx = min_dx,nb_x1 = nb_x1)
    f_bond <- F_affine_C$F - model$mu_c
    C    <- Ct * exp(t(apply(-cbind(0,f_bond),1,cumsum)))
    dc_h   <- (C[2,] - C[1,])/epsE
  }else{
    DF   <- NaN
    dc_h <- NaN
    F_affine_SDF <- NaN
    F_affine_C   <- NaN
  }
  
  SCC_noRP <- sum(-dc_h * DF)/10^9
  
  return(list(SCC=SCC,SCC_noRP=SCC_noRP,
              U = U,
              F_affine=F_affine,
              F_affine_SDF=F_affine_SDF,
              F_affine_C=F_affine_C,
              DF=DF,C=C,dc_h=dc_h))
}

res_SCC <- compute_SCC(model,HH=HH,indic_compute_riskpremium=TRUE)
SCC <- res_SCC$SCC

# Report the SCC with and without the risk-premium component.
scc_summary <- data.frame(
  estimate = c("With risk adjustment", "Without risk adjustment"),
  value = round(c(SCC, res_SCC$SCC_noRP), 2)
)
knitr::kable(scc_summary, col.names = c("Estimate", "SCC"))

```

The following table records the calibration used in the social-cost-of-carbon exercise.

```{r make_table_SCC, fig.width=8, fig.height=4}
# Purpose: The following table records the calibration used in the social-cost-of-carbon
# exercise

# Load required package
library(xtable)

# ---- Parameters to exclude ----
exclude_params <- c("rho","alpha","beta","n_w","h","tau_star")  # add any names you want to exclude

# Filter model list
model_filtered <- model[!names(model) %in% exclude_params]

# ---- Mapping between R names and LaTeX names ----
names_R <- c("mu_c","gamma","et","E0","Et","delta","mu_e","rho_e","rho","S",
             "alpha","beta","tau_star","rho_star","stdv_tau_star","Ct")

names_Latex <- c(
  "$\\mu_c$",       # Mean consumption growth rate
  "$\\gamma$",      # Risk aversion
  "$e_t$",          # Carbon growth rate
  "$\\mathcal{E}_0$", # Pre-industrial carbon stock
  "$\\mathcal{E}_t$", # Current carbon stock
  "$\\delta$",      # Discount factor
  "$\\mu_e$",       # Mean carbon growth rate
  "$\\rho_e$",      # Persistence of carbon growth
  "$\\rho$",        # Persistence of carbon stock
  "$S$",            # ECS
  "$\\alpha$",      # Damage slope
  "$\\beta$",       # High-damage slope
  "$\\tau_*$",      # Tipping threshold
  "$\\rho_*$",      # Persistence of threshold
  "$\\sigma(\\tau_t^*)$", # Volatility of threshold
  "$C_t$"           # Global consumption
)

# Create a lookup table
latex_map <- setNames(names_Latex, names_R)

# ---- Convert to data frame ----
param_df <- data.frame(
  Parameter = names(model_filtered),
  Value = unlist(model_filtered)
)

# Replace R names by LaTeX names
param_df$Parameter <- latex_map[param_df$Parameter]

# ---- Format numbers ----
param_df$Value <- formatC(param_df$Value, format = "g", digits = 5)

# ---- Optional descriptions ----
descriptions <- c(
  "$\\mu_c$" = "Mean consumption growth rate (5 years)",
  "$\\gamma$" = "Risk aversion coefficient",
  "$e_t$" = "Initial carbon growth rate (5 years)",
  "$\\mathcal{E}_0$" = "Pre-industrial carbon stock (GtCO$_2$)",
  "$\\mathcal{E}_t$" = "Current carbon stock (GtCO$_2$)",
  "$\\delta$" = "Time discount factor (5 years time step)",
  "$\\mu_e$" = "Scale of carbon growth rate shocks",
  "$\\rho_e$" = "Persistence of carbon growth rate",
  "$S$" = "Equilibrium climate sensitivity (°C per doubling)",
  "$\\alpha$" = "Baseline damage slope (change in 5-year consumpotion growth per °C)",
  "$\\beta$" = "High-damage slope (change in 5-year consumpotion growth per °C",
  "$\\tau_*$" = "Expected tipping threshold (°C)",
  "$\\rho_*$" = "Persistence of tipping threshold",
  "$\\sigma(\\tau_t^*)$" = paste0(
    "Unconditional volatility of tipping threshold ",
    "($\\sqrt{\\mathbb{V}ar(\\tau_t^*)}$, in °C)"
  ),
  "$C_t$" = "Global consumption (USD)"
)

param_df$Description <- descriptions[param_df$Parameter]

# ---- Generate LaTeX table ----
latex_table <- c(
  "\\begin{table}[ht]",
  "\\centering",
  "\\caption{This table reports the parameterization of the model used in Example~\\ref{exm.SCC}.}",
  "\\label{tab:model_parameters}",
  "\\begin{tabularx}{\\textwidth}{l r r X}",
  "\\hline",
  "Parameter & Value && Description \\\\",
  "\\hline",
  apply(param_df, 1, function(row) {
    paste(row['Parameter'], '&', row['Value'], '&&', row['Description'], '\\\\')
  }),
  "\\hline",
  "\\end{tabularx}",
  "% ---- Add explanatory note below ----",
  "\\vspace{0.3em}",
  paste0(
    "\\parbox{1.0\\textwidth}{\\footnotesize\\textit{Notes}: ",
    "Calibration values are based on standard IAM assumptions for ",
    "consumption growth and equilibrium climate sensitivity. Carbon stock ",
    "and temperature parameters are consistent with historical data. The ",
    "unconditional volatility of the threshold satisfies ",
    "$\\sigma(\\tau_t^*) = \\sqrt{1-\\rho_*^2}\\sigma_*$.}"
  ),
  "\\end{table}"
)

# ---- Save to file ----
output_dir <- "tables"
output_file <- file.path(output_dir, "table_param_SCC.txt")
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
writeLines(latex_table, con = output_file)

```


This figure shows how the social cost of carbon responds to the average tipping threshold and to the curvature of damages.

```{r fig_SCC, fig.width=7, fig.height=4}
# Purpose: This figure shows how the social cost of carbon responds to the average tipping
# threshold and to the curvature of damages

# Define the baseline damage slopes and threshold grid.
alpha_baseline <- .001 * time_step_yrs * 1

model$alpha <- .001 * time_step_yrs * 1
model$beta  <- .001 * time_step_yrs * 2
model$tau_star <- 2.5

all_alpha <- c(.5*alpha_baseline,alpha_baseline,alpha_baseline)
all_beta  <- c(2*alpha_baseline,2*alpha_baseline,4*alpha_baseline)

all_tau_star <- rev(seq(2.25,4.25,by=.5))
SCC_matrix <- matrix(NaN,length(all_alpha),length(all_tau_star))

# Recompute the SCC over the damage and threshold scenarios.
for(i in 1:length(all_alpha)){
  for(j in 1:length(all_tau_star)){
    model$alpha <- all_alpha[i]
    model$beta  <- all_beta[i]
    model$tau_star <- all_tau_star[j]
    res_SCC <- compute_SCC(model,HH=HH,
                           indic_compute_riskpremium=TRUE,
                           eps = 10^(-6), # to compute dG
                           max_x = 10000,
                           dx_statio = 200,
                           min_dx = 1e-05,
                           nb_x1 = 500)
    SCC_matrix[i,j] <- res_SCC$SCC
  }
}

# Plot the SCC response curves.
par(mfrow=c(1,1))
par(plt=c(.15,.95,.2,.95))
plot(all_tau_star, 0 * all_tau_star, ylim = c(0, max(SCC_matrix)),
     col = "white",
     xlab = expression(paste("Average temperature threshold (",
                             tau['*'], "), in °C", sep = "")),
     ylab = "Social Cost of Carbon, in USD/tCO2")
grid()
for(i in 1:length(all_alpha)){
  spline_points <- spline(all_tau_star, SCC_matrix[i,], n = 50)
  lines(spline_points$x,spline_points$y,lwd=2,lty=i)
  # points(all_tau_star, SCC_matrix[i,],col="red")
}

legend_labels <- mapply(function(a, b) {
  bquote(alpha == .(100*a/time_step_yrs)*"%, " ~ beta == .(100*b/time_step_yrs)*"%")
}, all_alpha, all_beta, SIMPLIFY = FALSE)

# Add legend
legend("topright", legend = legend_labels, lty = 1:3, lwd = 2)

```


The final block checks the analytical formula against simulated paths of the climate-economy system.

```{r SCC_check_formula, results='hide'}
# Purpose: Check the analytical formula against simulated climate-economy paths.
# Check formulas using simulations.

nb.replic <- 40000
TT <- 200
set.seed(123)

# Compute the analytical benchmark.
res_SCC <- compute_SCC(model,HH=TT,indic_compute_riskpremium=TRUE,
                       eps = 10^(-5), # to compute dG
                       max_x = 100000,
                       dx_statio =100,
                       min_dx = 1e-8,
                       nb_x1 = 10000)

model$sigma_star <- model$stdv_tau_star*sqrt(1-model$rho_star^2)

# Simulate the tipping threshold and the carbon stock.
res <- 
  simul_GaussianVAR(list(mu=model$tau_star*(1-model$rho_star),Phi = matrix(model$rho_star),
                  Sigma12 = matrix(model$sigma_star)),
             nb.sim=TT,nb.replic = nb.replic)
sim_tau_star <- matrix(res,nrow=dim(res)[1])

sim_e <- simul_ARG(nb.sim = TT,
                   mu=model$mu_e,
                   alpha=0,
                   nu = 0,
                   rho = model$rho_e,
                   w0 = model$et,
                   nb.replic=nb.replic)
sim_logE    <- log(model$Et) + apply(sim_e,2,cumsum)
epsE <- 1
sim_logE_dE <- log(model$Et + epsE) + apply(sim_e,2,cumsum)
sim_tau    <- model$S/log(2)*(sim_logE - log(model$E0))
sim_tau_dE <- model$S/log(2)*(sim_logE_dE - log(model$E0))

sim_dc <- model$mu_c - model$alpha*sim_tau - model$beta*pmax(0,sim_tau - sim_tau_star)
sim_dc_demeaned <- - model$alpha*sim_tau - model$beta*pmax(0,sim_tau - sim_tau_star)

sim_dc_dE <- model$mu_c - model$alpha*sim_tau_dE - model$beta*pmax(0,sim_tau_dE - sim_tau_star)

# Compare simulated transform moments with the affine formula. The left panel
# checks the baseline discounted transform; the right panel checks the small
# change generated by the one-GtCO2 emissions perturbation used to compute SCC.
par(mfrow=c(1,2))
par(mar=c(4.2,4.4,3.2,1.2))

sim_exp_sum_dc    <- exp(apply(log(model$delta)+(1-model$gamma)*sim_dc,2,cumsum))
sim_exp_sum_dc_dE <- exp(apply(log(model$delta)+(1-model$gamma)*sim_dc_dE,2,cumsum))
E_exp_sum_dc_simul <- apply(sim_exp_sum_dc,1,mean)
E_exp_sum_dc_dE_simul <- apply(sim_exp_sum_dc_dE,1,mean)
f <- res_SCC$F_affine$F - log(model$delta) - (1-model$gamma)*model$mu_c
E_exp_sum_dc_formula <- exp(t(apply(-f,1,cumsum)))
plot(E_exp_sum_dc_simul,type="l",lwd=4,col="grey70",
     xlab="Horizon",ylab="Baseline transform moment",las=1,
     main="(a) Baseline transform")
lines(E_exp_sum_dc_formula[1,],col="black",lty=2,lwd=2)
grid()
legend("topright",c("Monte Carlo","Affine formula"),
       col=c("grey70","black"),lty=c(1,2),lwd=c(4,2),bty="n")

plot(E_exp_sum_dc_dE_simul-E_exp_sum_dc_simul,type="l",lwd=2,col="grey",
     ylim=c(min(E_exp_sum_dc_formula[2,]-E_exp_sum_dc_formula[1,]),
            max(E_exp_sum_dc_formula[2,]-E_exp_sum_dc_formula[1,])),
     xlab="Horizon",ylab="Effect of one additional GtCO2",las=1,
     main="(b) Marginal-emissions effect")
lines(E_exp_sum_dc_formula[2,]-E_exp_sum_dc_formula[1,],
      col="black",lty=2,lwd=2)
grid()

SCC_simul <- model$Ct*time_step_yrs/(model$gamma-1)*
  (sum(E_exp_sum_dc_dE_simul) - sum(E_exp_sum_dc_simul))/epsE/10^9

```

The two curves are nearly indistinguishable in both panels. The affine calculation gives an SCC of `r round(res_SCC$SCC, 1)` USD/tCO2, compared with `r round(SCC_simul, 1)` USD/tCO2 from the Monte Carlo calculation.

## Monte Carlo assessment of the affine approximation used in the SCC example

The SCC calculation above uses `compute_F_Shadow_affine()` to handle the kink in damages at the stochastic temperature threshold. The following simpler Gaussian bond-pricing benchmark isolates the accuracy of this nonlinear affine approximation. It compares the generic affine method, two Gaussian shadow-rate approximations, and a model that ignores the lower bound with a direct Monte Carlo calculation.

```{r fig_SR1_check, fig.width=9, fig.height=4}
# Purpose: Benchmark the nonlinear affine approximation used in the SCC example
# against direct Monte Carlo integration in a transparent Gaussian setting.

validation_horizon <- 30
validation_lower_bound <- 0
validation_initial_state <- 0.03
validation_replicates <- 100000

# Use a self-contained Gaussian shadow-rate model so that this diagnostic does
# not depend on objects created elsewhere in the chapter.
validation_rho <- 0.96
validation_model <- list(
  mu = matrix(0.03 * (1 - validation_rho), 1, 1),
  Phi = matrix(validation_rho, 1, 1),
  Sigma = (1 - validation_rho^2) * 0.02^2 * matrix(1),
  n_w = 1
)
validation_model$Sigma12 <- matrix(sqrt(validation_model$Sigma))
validation_state <- matrix(validation_initial_state, 1, 1)

# Direct Monte Carlo bond yields provide the numerical benchmark.
set.seed(1234)
validation_simulated_state <- simul_GaussianVAR(
  validation_model,
  nb.sim = validation_horizon,
  x0 = validation_initial_state,
  nb.replic = validation_replicates
)
validation_simulated_state <- matrix(
  validation_simulated_state,
  validation_horizon,
  validation_replicates
)
validation_simulated_rate <- pmax(
  validation_simulated_state,
  validation_lower_bound
)
validation_discount_factor <- exp(
  -apply(validation_simulated_rate, 2, cumsum)
)
validation_yield_mc <- -log(rowMeans(validation_discount_factor)) /
  (1:validation_horizon)

# Compute the two Gaussian approximations and the generic affine approximation
# used by the SCC application.
validation_WX <- compute_F_Shadow_Gaussian(
  validation_state,
  psi.parameterization = validation_model,
  ell_bar = validation_lower_bound,
  b = 0,
  a = matrix(1),
  c = matrix(0),
  H = validation_horizon - 1,
  indic_WuXia = TRUE
)
validation_PR <- compute_F_Shadow_Gaussian(
  validation_state,
  psi.parameterization = validation_model,
  ell_bar = validation_lower_bound,
  b = 0,
  a = matrix(1),
  c = matrix(0),
  H = validation_horizon - 1,
  indic_WuXia = FALSE
)
validation_affine <- compute_F_Shadow_affine(
  validation_state,
  psi_GaussianVAR,
  psi.parameterization = validation_model,
  ell_bar = validation_lower_bound,
  b = 0,
  a = matrix(1),
  c = matrix(0),
  H = validation_horizon - 1
)

validation_current_rate <- pmax(
  validation_lower_bound,
  validation_initial_state
)
validation_yield_WX <- cumsum(
  c(validation_current_rate, validation_WX$F)
) / (1:validation_horizon)
validation_yield_PR <- cumsum(
  c(validation_current_rate, validation_PR$F)
) / (1:validation_horizon)
validation_yield_affine <- cumsum(
  c(validation_current_rate, validation_affine$F)
) / (1:validation_horizon)

# The unconstrained Gaussian model shows what is lost by ignoring the bound.
validation_no_bound <- compute_AB_classical(
  xi0 = 0,
  xi1 = matrix(1),
  H = validation_horizon,
  psi = psi_GaussianVAR,
  psi.parameterization = validation_model
)
validation_yield_no_bound <- c(validation_no_bound$b) +
  validation_initial_state * c(validation_no_bound$a)

validation_yields <- cbind(
  "Monte Carlo" = validation_yield_mc,
  "Wu-Xia" = validation_yield_WX,
  "Alternative Gaussian" = validation_yield_PR,
  "Generic affine" = validation_yield_affine,
  "No lower bound" = validation_yield_no_bound
)
validation_errors_bp <- 10000 * sweep(
  validation_yields[, -1, drop = FALSE],
  1,
  validation_yield_mc,
  "-"
)
validation_max_error_bp <- apply(abs(validation_errors_bp), 2, max)

# Panel (a) compares yield curves; Panel (b) makes the small approximation
# errors visible by expressing them in basis points relative to Monte Carlo.
validation_maturities <- 1:validation_horizon
validation_colours <- c("black", "grey25", "grey45", "grey60", "grey75")
validation_lty <- c(1, 2, 3, 4, 5)
validation_lwd <- c(3, 2, 2, 2, 2)

par(mfrow = c(1, 2), mar = c(4.2, 4.4, 3.2, 1.2))
matplot(
  validation_maturities,
  100 * validation_yields,
  type = "l",
  col = validation_colours,
  lty = validation_lty,
  lwd = validation_lwd,
  xlab = "Maturity",
  ylab = "Yield, percent",
  main = "(a) Yield curves",
  las = 1
)
grid()
legend(
  "bottomleft",
  colnames(validation_yields),
  col = validation_colours,
  lty = validation_lty,
  lwd = validation_lwd,
  bty = "n",
  cex = 0.78
)

matplot(
  validation_maturities,
  validation_errors_bp,
  type = "l",
  col = validation_colours[-1],
  lty = validation_lty[-1],
  lwd = validation_lwd[-1],
  xlab = "Maturity",
  ylab = "Error relative to Monte Carlo, bp",
  main = "(b) Approximation errors",
  las = 1
)
abline(h = 0, col = "grey80")
grid()
legend(
  "bottomleft",
  colnames(validation_errors_bp),
  col = validation_colours[-1],
  lty = validation_lty[-1],
  lwd = validation_lwd[-1],
  bty = "n",
  cex = 0.78
)
```

In this benchmark, the maximum absolute yield discrepancies are `r round(validation_max_error_bp["Wu-Xia"], 2)` basis points for the Wu-Xia approximation, `r round(validation_max_error_bp["Alternative Gaussian"], 2)` basis points for the alternative Gaussian approximation, and `r round(validation_max_error_bp["Generic affine"], 2)` basis points for the generic affine approximation used in the SCC calculation. Ignoring the lower bound produces the largest discrepancy (`r round(validation_max_error_bp["No lower bound"], 2)` basis points).
