# Forward, futures, dividends, commodity pricing, and convenience yields

## Persistent seasonal factors can generate recurring convenience-yield patterns

This example illustrates how persistent seasonal factors can generate recurring convenience-yield patterns.

```{r fig_seasonal, fig.width=7, fig.height=3}
# Purpose: This example illustrates how persistent seasonal factors can generate recurring
# convenience-yield patterns
library(DTAM)

ell <- 12
nb_j <- trunc(ell/2)

# Build the trigonometric transition matrix for seasonal cycles.
rho <- rep(.995,nb_j)
# sigma_xi <- rep(.01,nb_j)
sigma_xi <- c(.01,.01,rep(0,nb_j-2))

Theta  <- matrix(0,2*nb_j,2*nb_j)
for(j in 1:nb_j){
  lambda_j <- 2*pi*j/ell
  Theta[(2*(j-1)+1):(2*j),(2*(j-1)+1):(2*j)] <- rho[j] * matrix(c(+cos(lambda_j),
                                                                  -sin(lambda_j),
                                                                  +sin(lambda_j),
                                                                  +cos(lambda_j)),2,2)
}
Omega <- diag(sigma_xi %x% c(1,1))

nb.sim <- 100*ell
start_set <- 92*ell

set.seed(123)

# Simulate the seasonal state vector.
model <- list(Phi   = Theta,
              mu    = matrix(0,2*nb_j,1),
              Sigma12 = Omega)

x0 <- matrix(rep(1,2*nb_j),ncol=1)
x0 <- rnorm(n=2*nb_j)

simulated_gamma <- simul_GaussianVAR(model,x0 = x0,nb.sim = nb.sim)
gamma <- simulated_gamma %*% (matrix(1,nb_j,1) %x% matrix(c(1,0),2,1))

# Plot the final part of the simulated seasonal component.
par(mfrow=c(1,1))
par(plt=c(.15,.95,.15,.95))
plot(gamma[(start_set+1):nb.sim],col="white",
     type="l",las=1,
     xlab="",ylab="")
abline(h=0,col="grey")
lines(gamma[(start_set+1):nb.sim],lwd=2)
grid()
```


## Gaussian state-space model for commodity futures

This example fits a Gaussian state-space model to commodity futures and compares fitted and observed curves.
The futures panel contains NYMEX natural-gas futures settlement prices. The rows are monthly observation dates from January 2010 to January 2025, and the columns are contract-specific settlement-price series for monthly maturities from the February 2010 contract to the December 2030 contract. The data were extracted through licensed [LSEG Workspace](https://www.lseg.com/en/data-analytics/products/workspace/data-and-content) and [Datastream](https://www.lseg.com/en/data-analytics/products/datastream-macroeconomic-analysis) queries: the Workspace/Refinitiv sheet uses RIC-style futures tickers such as `NGG25`, while the Datastream sheet uses mnemonics such as `NNG0210` and series names of the form `NYM-NATURAL GAS FEB 2010 - SETT. PRICE`.^[The underlying settlement-price observations are not distributed with DTAM or with the public version of this Rbookdown project. A public blank extraction workbook, with the observation dates, contract headers, and contract metadata but no price observations, is provided here: [data/Futures_extraction_template.xlsx](data/Futures_extraction_template.xlsx).]

```{r fig_futures_KF, fig.width=8, fig.height=8}
# Purpose: This example fits a Gaussian state-space model to commodity futures and compares
# fitted and observed curves
library(DTAM)
library(stringr)

private_futures_file <- "private_data/Futures.rda"
precomputed_figure <- if (knitr::is_latex_output()) {
  "_main_files/figure-latex/fig_futures_KF-1.pdf"
} else {
  "figure/fig_futures_KF-1.png"
}

if (!file.exists(private_futures_file) && file.exists(precomputed_figure)) {
  knitr::include_graphics(precomputed_figure)
} else {
  if (!file.exists(private_futures_file)) {
    stop("Private futures data not found: ", private_futures_file, call. = FALSE)
  }
  load(private_futures_file)

  # Prepare the panel of futures prices by pricing date and time to maturity.

  tickers <- names(Futures)[2:dim(Futures)[2]]
  nb_maturities <- length(tickers)

month_letters <- c("JAN","FEB","MAR","APR","MAY","JUN",
                   "JUL","AUG","SEP","OCT","NOV","DEC")

months_letter_in_ticker <- str_sub(tickers,17,19)
months <- apply(matrix(1:length(months_letter_in_ticker),ncol=1),1,
                function(i){which(months_letter_in_ticker[i]==month_letters)})
years <- as.numeric(str_sub(tickers,21,24))
tickers_maturities <- as.Date(paste(years,"-",months,"-01",sep=""))

nb_dates <- 60
N <- dim(Futures)[1]
list_smpl_dates <- sort(Futures$Date[(N-nb_dates+1):N])

max_maturity <- 60 # in months

data <- matrix(NaN,nb_dates,max_maturity)
tempo <- NULL
for(count_date in 1:nb_dates){
  ddate <- list_smpl_dates[count_date]
  indic_date <- which(Futures$Date==ddate)
  for(iii in 1:nb_maturities){
    p <- Futures[indic_date,1+iii]
    # Compute maturity, in months.
    diff_years <- as.numeric(format(as.Date(tickers_maturities[iii]), "%Y")) -
      as.numeric(format(ddate, "%Y"))
    diff_month <- as.numeric(format(as.Date(tickers_maturities[iii]), "%m")) -
      as.numeric(format(ddate, "%m"))
    maturity_in_months <- 12*diff_years + diff_month
    
    if((maturity_in_months >= 1)&(maturity_in_months <= max_maturity)){
      data[count_date,maturity_in_months] <- p
    }
  }
}
Y <- log(data)


# Build the Gaussian state-space representation used in the Kalman filter.

make_model <- function(phi,rho,sigma_xi,sigma_mu,sigma_z){
  
  Omega    <- diag(sigma_xi %x% c(1,1))
  
  Theta  <- matrix(0,2*nb_j,2*nb_j)
  for(j in 1:nb_j){
    lambda_j <- 2*pi*j/ell
    Theta[(2*(j-1)+1):(2*j),(2*(j-1)+1):(2*j)] <- rho[j] * matrix(c(+cos(lambda_j),
                                                                    -sin(lambda_j),
                                                                    +sin(lambda_j),
                                                                    +cos(lambda_j)),2,2)}
  
  Phi <- matrix(0,2+nb_j*2,2+nb_j*2)
  Phi[3:(2+nb_j*2),3:(2+nb_j*2)] <- Theta
  Phi[1,1] <- 1
  Phi[2,2] <- phi
  
  Sigma <- matrix(0,2+nb_j*2,2+nb_j*2)
  Sigma[1,1] <- sigma_mu
  Sigma[2,2] <- sigma_z
  Sigma[3:(2+nb_j*2),3:(2+nb_j*2)] <- Omega
  
  # Pricing formulas:
  psi.parameterization <- list(
    mu = matrix(0,2+2*nb_j,1),
    Phi = Phi,
    Sigma12 = Sigma,
    Sigma = Sigma %*% t(Sigma))
  
  return(psi.parameterization)
}

estim_model <- function(param){
  
  phi <- exp(param[1])/(1+exp(param[1]))
  rho <- exp(param[2:(1+nb_j)])/(1+exp(param[2:(1+nb_j)]))
  # sigma_xi <- param[(1+nb_j+1):(1+2*nb_j)]
  # sigma_mu <- param[2+2*nb_j]
  # sigma_z  <- param[3+2*nb_j]
  
  # Pricing formulas:
  model <- make_model(phi,rho,sigma_xi,sigma_mu,sigma_z)
  
  # Use the reverse-order multi-horizon transform to price futures across maturities.
  h = max_maturity # maximum maturity
  u1 <- matrix(c(1,1,rep(c(1,0),nb_j)),ncol=1)
  u2 <- 0 * u1
  AB <- reverse_MHLT(psi_GaussianVAR,u1,u2,H=h,psi.parameterization=model)
  
  # Define matrices needed in the Kalman_filter procedures:
  nu_t <- matrix(0,dim(Y)[1],2+2*nb_j) # intercept in transition equation
  H <- model$Phi # autoregressive matrix in transition equation
  G <- t(matrix(AB$A,2+2*nb_j,max_maturity)) # matrix of loadings
  mu_t <- t(matrix(AB$B,max_maturity,nb_dates))
  N <- model$Sigma12 # volatility matrix in transition equation
  M <- diag(rep(.1,max_maturity))  # matrix of std dev of measurement errors
  Sigma_0 <- 2*diag(2+2*nb_j) # unconditional variance of w
  rho_0 <- rep(0,2+2*nb_j) # unconditional mean of w
  filter.res   <- Kalman_filter(Y,nu_t,H,N,mu_t,G,M,Sigma_0,rho_0)
  # smoother.res <- Kalman_smoother(Y,nu_t,H,N,mu_t,G,M,Sigma_0,rho_0)
  
  return(filter.res)
}

loglik <- function(param){
  # print(param)
  filter.res <- estim_model(param)
  return(-filter.res$loglik)
}


# Set model specifications:
ell <- 12
nb_j <- trunc(ell/2)

# Model parameterizaton:
rho      <- rep(.995,nb_j)
phi      <- .98 # autocorrelation of z
sigma_xi <- rep(0.01,nb_j)
sigma_mu <- .03
sigma_z  <- .10

param <- c(log(phi/(1-phi)),log(rho/(1-rho)),sigma_xi,sigma_mu,sigma_z)

# res.optim <- optim(par=param,
#                    fn=log-likelihood,
#                    method="Nelder-Mead",
#                    control=list(trace=FALSE,maxit=50),hessian=FALSE)
# param <- res.optim$par

filter.res <- estim_model(param)

Y_fitted <- filter.res$fitted.obs


# Produce the chart comparing fitted and observed futures curves.

indic_selected_dates <-  c(9,33,60)
selected_dates <- list_smpl_dates[indic_selected_dates]

cols <- gray.colors(n=length(indic_selected_dates), start = .0, end = .7)

nf <- layout(
  matrix(c(1,2), nrow=2, ncol=1, byrow=TRUE),
  widths=c(1,1),
  heights=c(1,2)
)

par(plt=c(.1,.95,.2,.8))
# par(mfrow=c(2,1))

# sigma_mu <- sd(diff(filter.res$r[,1]))
# sigma_z  <- sd(diff(filter.res$r[,2]))
# plot(list_smpl_dates,filter.res$r[,1],type="l")
# lines(list_smpl_dates,filter.res$r[,2],type="b")

gamma <- filter.res$r[,3:(2+2*nb_j)] %*% matrix(rep(c(1,0),nb_j),ncol=1)
plot(list_smpl_dates,gamma,type="l",lwd=2,las=1,
     xlab="",ylab="",
     main=expression(paste("(a) Estimated seasonality component ",gamma[t],sep="")))
abline(h=0,col="black")
grid()

plot(tickers_maturities,rep(0,length(tickers_maturities)),
     xaxt="n",yaxt="n",las=1,
     xlab = "maturity",ylab = "",
     ylim=c(-.5 + min(exp(Y[indic_selected_dates,]),na.rm = TRUE),
            .0 + max(exp(Y[indic_selected_dates,]),na.rm = TRUE)),
     xlim=c(as.Date("2020-06-01"),as.Date("2030-03-01")),
     main=expression(paste("(b) Term structure of futures prices ",F[t*','*T],sep="")),
     col="white")

xlab <- seq(as.Date("2020-01-01"), by="+6 month", length.out=60)
axis.Date(1, at=xlab, format="%m.%y",las=2)
for(i in 1:length(xlab)){
  abline(v=xlab[i],col="grey",lty=3)
}
ylab <- seq(trunc(exp(min(Y[indic_selected_dates,],na.rm = TRUE))),
            trunc(exp(max(Y[indic_selected_dates,],na.rm = TRUE))+1),by=.5)
axis(2,at=ylab,labels=ylab,las=1)
for(i in 1:length(ylab)){
  abline(h=ylab[i],col="grey",lty=3)
}

count_date <- 0
for(ddate in selected_dates){
  count_date <- count_date + 1
  ddate <- as.Date(ddate)
  indic_date <- which(ddate == list_smpl_dates)
  
  # x-axis
  x_axis <- seq(ddate, by="+1 month", length.out=max_maturity+1)
  x_axis <- x_axis[2:length(x_axis)]
  
  # y-axis
  y_axis_fitted <- exp(Y_fitted[indic_date,])*complete.cases(Y[indic_date,])
  y_axis_data   <- exp(Y[indic_date,])
  
  points(x_axis,y_axis_fitted,pch=3,col=cols[count_date],lwd=2)
  lines(x_axis,y_axis_data,col=cols[count_date],lwd=2)
}

legend("topleft",
       title="Pricing date:",
       legend=selected_dates,
       lwd=c(2),lty=c(1),col=cols,seg.len = 2)

legend("bottomright",
       legend=c("Data","Model"),
       lwd=c(2),lty=c(1,NaN),col="black",pch=c(NaN,3),seg.len = 2)

}

```

*Source: LSEG Datastream via LSEG Workspace.*
