# Consumption Capital Asset Pricing Model {#CCAPM}

## Risk aversion and intertemporal substitution in the CCAPM

This example contrasts the roles of risk aversion and intertemporal substitution in the CCAPM.

```{r fig_RRAIESCCAPM0, fig.width=8, fig.height=4}
# Purpose: This example contrasts the roles of risk aversion and intertemporal substitution in
# the CCAPM

# Define the two-state consumption lottery.
C.l <- .8
C.h <- 1.2
cc <- c(C.l,C.h)

par(mfrow=c(1,2))
par(plt=c(.15,.95,.25,.85))

C <- seq(.7,1.5,by=.01)

# Plot certainty equivalents for moderate risk aversion.
gamma <- 2
E.u <- mean(cc^(1-gamma)/(1-gamma))
c.CE <- (E.u*(1-gamma))^(1/(1-gamma))
U <- C^(1-gamma)/(1-gamma)
plot(C,U,type="l",lwd=2,las=1,
     xlab="Consumption",ylab="Utility",
     main=expression(paste("(a) ",gamma," = 2",sep="")))
points(cc,cc^(1-gamma)/(1-gamma),pch=19,col="black",cex=1.5)
lines(cc,cc^(1-gamma)/(1-gamma),col="black")
lines(c(c.CE,mean(cc)),c(E.u,E.u),lty=1,col="darkgrey",lwd=1)
lines(c(c.CE,c.CE),c(-1000,E.u),lty=1,col="darkgrey",lwd=1)
lines(c(mean(cc),mean(cc)),c(-1000,1/(1-gamma)),lty=1,col="darkgrey",lwd=1)
points(c(1,1),c(mean(cc^(1-gamma)/(1-gamma)),1/(1-gamma)),col="darkgrey",pch=19)
lines(c(1,1),c(mean(cc^(1-gamma)/(1-gamma)),1/(1-gamma)),col="darkgrey")
rp <- - mean(cc^(1-gamma)/(1-gamma)) + 1/(1-gamma)

# Annotate the certainty equivalent and the risk-premium gap.
text(x = 1.3,y=1/(1-gamma) - .2,
     labels = paste("Risk premium = ",toString(round(mean(cc)-c.CE,2)),sep=""),
     col="black")
lines(c(mean(c(c.CE,1)),1.2),c(C[1]^(1-gamma)/(1-gamma),1/(1-gamma) - .22),col="black")

# Repeat the same construction with higher risk aversion.
gamma <- 10
E.u <- mean(cc^(1-gamma)/(1-gamma))
c.CE <- (E.u*(1-gamma))^(1/(1-gamma))
U <- C^(1-gamma)/(1-gamma)

plot(C,U,type="l",lwd=2,las=1,
     xlab="Consumption",ylab="Utility",
     main=expression(paste("(b) ",gamma," = 10",sep="")))
points(cc,cc^(1-gamma)/(1-gamma),pch=19,col="black",cex=1.5)
lines(cc,cc^(1-gamma)/(1-gamma),col="black")
lines(c(c.CE,mean(cc)),c(E.u,E.u),lty=1,col="darkgrey",lwd=1)
lines(c(c.CE,c.CE),c(-1000,E.u),lty=1,col="darkgrey",lwd=1)
lines(c(mean(cc),mean(cc)),c(-1000,1/(1-gamma)),lty=1,col="darkgrey",lwd=1)
points(c(1,1),c(mean(cc^(1-gamma)/(1-gamma)),1/(1-gamma)),col="darkgrey",pch=19)
lines(c(1,1),c(mean(cc^(1-gamma)/(1-gamma)),1/(1-gamma)),col="darkgrey")
rp <- - mean(cc^(1-gamma)/(1-gamma)) + 1/(1-gamma)
text(x = 1.3,y=1/(1-gamma) - 1,
     labels = paste("Risk premium = ",toString(round(mean(cc)-c.CE,2)),sep=""),
     col="black")
lines(c(mean(c(c.CE,1)),1.2),c(C[1]^(1-gamma)/(1-gamma),1/(1-gamma) - 1.05),col="black")
```

## Asset-pricing implications of risk-aversion and EIS calibrations

This example visualizes asset-pricing implications of different risk-aversion and EIS calibrations.

```{r fig_RRAIESCCAPM, fig.width=7, fig.height=3.5}
# Purpose: This example visualizes asset-pricing implications of different risk-aversion and EIS
# calibrations

# Define the two-state consumption lottery.
C.l <- .8
C.h <- 1.2
cc <- c(C.l,C.h)

par(mfrow=c(1,1))
par(plt=c(.15,.95,.25,.95))

C <- seq(.7,1.5,by=.01)

# Plot the certainty equivalent and the implied risk premium.
gamma <- 5
E.u <- mean(cc^(1-gamma)/(1-gamma))
c.CE <- (E.u*(1-gamma))^(1/(1-gamma))
U <- C^(1-gamma)/(1-gamma)
plot(C,U,type="l",lwd=2,las=1,
     xlab="Consumption",ylab="Utility")
grid()
points(cc,cc^(1-gamma)/(1-gamma),pch=19,col="black",cex=1.5)
lines(cc,cc^(1-gamma)/(1-gamma),col="black")
lines(c(c.CE,mean(cc)),c(E.u,E.u),lty=1,col="darkgrey",lwd=1)
lines(c(c.CE,c.CE),c(-1000,E.u),lty=1,col="darkgrey",lwd=1)
lines(c(mean(cc),mean(cc)),c(-1000,1/(1-gamma)),lty=1,col="darkgrey",lwd=1)
points(c(1,1),c(mean(cc^(1-gamma)/(1-gamma)),1/(1-gamma)),col="darkgrey",pch=19)
lines(c(1,1),c(mean(cc^(1-gamma)/(1-gamma)),1/(1-gamma)),col="darkgrey")
rp <- - mean(cc^(1-gamma)/(1-gamma)) + 1/(1-gamma)
text(x = 1.3,y=1/(1-gamma) - .2,
     labels = paste("Risk premium = ",toString(round(mean(cc)-c.CE,2)),sep=""),
     col="black")
lines(c(mean(c(c.CE,1)),1.2),c(C[3]^(1-gamma)/(1-gamma),1/(1-gamma) - .24),col="black")
text(x=C.l,y=min(U),labels = expression(C[L]),pos=2)
text(x=C.h,y=min(U),labels = expression(C[H]),pos=2)
text(x=c.CE,y=min(U),labels = expression(tilde(C)),pos=2)
text(x=1,y=min(U),labels = expression(bar(C)),pos=2)

# Mark the risk-premium distance on the consumption axis.
arrows(c.CE,C[3]^(1-gamma)/(1-gamma),1,C[3]^(1-gamma)/(1-gamma),lwd=2,
       angle = 20,code = 3,length = 0.1)

```

The following codes makes a table that illustrates the equity premium puzzle.

This table summarizes the long-run joint behavior of macro-financial variables used in the CCAPM discussion.

```{r table_JST}
# Purpose: This table summarizes the long-run joint behavior of macro-financial variables used
# in the CCAPM discussion

library(DTAM)

lag_exp_infl <- 5 # order of moving average process used to capture inflation expectations

matrix_periods <- rbind(c(1950,2019),
                        c(1980,2019))

list.of.JST.ctries <- levels(as.factor(JSTdataset$iso))
# remove countries with no equity return data:
list.of.JST.ctries <- list.of.JST.ctries[!(list.of.JST.ctries %in% c("CAN","IRL"))]

matrix_corresp_ctries <- rbind(c("AUS","BEL","CAN","CHE","DEU","DNK","ESP",
                                 "FIN","FRA","GBR","IRL","ITA","JPN","NLD","NOR",
                                 "PRT","SWE","USA"),
                               c("Australia","Belgium","Canada","Switzerland","Germany",
                                 "Denmark","Spain","Finland","France","U.K.",
                                 "Ireland","Italy","Japan","Netherlands","Norway",
                                 "Portugal","Sweden","U.S.A."))

# outputs: Sample-period,
#          Average excess return, sd(return), sd(dc), Cor(dc,er), Cov(dc,er),
#          Sharpe ratio, RRA1, RRA2
matrix_result <- array(NaN,c(length(list.of.JST.ctries),9,dim(matrix_periods)[1]))

DATA <- JSTdataset[,c("year","iso","cpi","stir","ltrate","eq_tr",
                      "rconsbarro")]

for(period in 1:dim(matrix_periods)[1]){
  first.year <- matrix_periods[period,1]
  last.year  <- matrix_periods[period,2]
  
  count_ctry <- 0
  for(ctry in list.of.JST.ctries){
    count_ctry <-   count_ctry + 1
    DATA.reduced <- subset(DATA,iso==ctry)
    T <- dim(DATA.reduced)[1]
    DATA.reduced$dc <- NaN
    DATA.reduced$dc <- 
      c(NaN,log(DATA.reduced$rconsbarro[2:T]/DATA.reduced$rconsbarro[1:(T-1)]))
    DATA.reduced$eq_tr
    DATA.reduced$infl <-
      c(NaN,log(DATA.reduced$cpi[2:T]/DATA.reduced$cpi[1:(T-1)]))
    DATA.reduced$exp_infl <- 0
    for(i in 1:lag_exp_infl){
      DATA.reduced$exp_infl <- DATA.reduced$exp_infl + 1/lag_exp_infl*
        c(rep(NaN,i),DATA.reduced$infl[1:(T-i)])
    }
    DATA.reduced$rstir <- DATA.reduced$stir/100 - DATA.reduced$exp_infl
    DATA.reduced$req_tr <- DATA.reduced$eq_tr - DATA.reduced$infl
    DATA.reduced <- DATA.reduced[which(DATA.reduced$year==first.year):
                                   which(DATA.reduced$year==last.year),]
    
    sample <- range(DATA.reduced$year[complete.cases(DATA.reduced)])
    
    DATA.reduced$Excess_return <- DATA.reduced$req_tr - DATA.reduced$rstir
    
    mean_er <- mean(DATA.reduced$Excess_return,na.rm=TRUE)
    sd_er   <- sd(DATA.reduced$Excess_return,na.rm=TRUE)
    sd_dc   <- sd(DATA.reduced$dc,na.rm=TRUE)
    cor_er_dc <- cor(DATA.reduced$dc,DATA.reduced$req_tr,use="complete.obs")
    cov_er_dc <- cov(DATA.reduced$dc,DATA.reduced$req_tr,use="complete.obs")
    SR <- mean_er/sd_er
    mean_Rf <- mean(DATA.reduced$rstir,na.rm=TRUE)
    RRA1 <- mean_er/(1 + mean_Rf)/cov_er_dc
    RRA2 <- mean_er/(1 + mean_Rf)/sd_er/sd_dc
    
    matrix_result[count_ctry,,period] <- 
      c(paste(sample[1],"-",sample[2],sep=""),
        round(100*mean_Rf,2),
        round(100*mean_er,2),round(100*sd_er,2),
        round(100*sd_dc,2),
        round(cor_er_dc,2),
        round(SR,2),round(RRA1,1),round(RRA2,1))
  }
}


make_entry <- function(x,format.nb){
  output <- paste("$",sprintf(paste("%.",format.nb,"f",sep=""),x),"$",sep="")
  return(output)
}

format.nb <- 2

# Build LaTeX table
latex.table <- paste0(
  "Country&Sample&$R_f$&$\\overline{er}$&$\\sigma(er)$",
  "&$\\sigma(\\Delta c)$&$\\rho(\\Delta c,er)$&$SR$",
  "&$RRA1$&$RRA2$\\\\"
)
for(period in 1:dim(matrix_periods)[1]){
  latex.table <- rbind(latex.table,
                       "\\hline",
                       paste("&\\multicolumn{9}{l}{\\textbf{",LETTERS[period],". Sample: ",
                             matrix_periods[period,1],"-",matrix_periods[period,2],"}}\\\\",sep=""),
                       "\\hline")
  count_ctry <- 0
  for(ctry in list.of.JST.ctries){
    count_ctry <- count_ctry + 1
    this.line <- matrix_corresp_ctries[2,which(ctry==matrix_corresp_ctries[1,])]
    for(j in 1:dim(matrix_result)[2]){
      if(j==1){
        this.line <- paste(this.line,"&",
                           matrix_result[count_ctry,j,period],sep="")
      }else{
        num_value <- as.numeric(matrix_result[count_ctry,j,period])
        if((num_value<0)&(j>7)){
          this.line <- paste(this.line,"&$ < 0$",sep="")
        }else{
          this.line <- paste(this.line,"&",make_entry(num_value,format.nb),sep="")
        }
      }
    }
    latex.table <- rbind(latex.table,
                         paste(this.line,"\\\\",sep=""))
  }
}
latex.table <- rbind(latex.table,
                     "\\hline")

latex.file <- paste("tables/table_EquityPuzzle.txt",sep="")
write(latex.table, file = latex.file)

```

The following code illustrates the link between IES and smoothness of the optimal consumption path.

## Smoothing preferences affect consumption and pricing dynamics

This example shows how smoothing preferences affect consumption and pricing dynamics.

```{r fig_IESsmooth, fig.width=7.5, fig.height=3.5}
# Purpose: This example shows how smoothing preferences affect consumption and pricing dynamics

par(mfrow=c(1,1))
par(plt=c(.15,.95,.25,.95))

# Set the two-period saving problem.
W <- 1
delta <- 1
R <- .05

# Compute and plot optimal consumption paths for several IES values.
gamma <- 1.5
C1 <- W * (1 + R) / ((1/delta * 1/(1+R))^(-1/gamma) + (1+R))
C2 <- (W - C1) * (1 + R)
par(mgp = c(3, 0.5, 0))

# Draw the baseline path before overlaying alternative IES cases.
plot(c(C1,C2),pch=19,xlab="Period",ylab="Optimal consumption path",ylim=c(.50,.525),
     xaxt="n",xlim=c(.75,2.25),lwd=2,cex=1.3,las=1)
grid()
axis(side=1,at=c(1,2),labels=c("t = 0","t = 1"))
lines(c(C1,C2),lwd=2)
gamma <- 5
C1 <- W * (1 + R) / ((1/delta * 1/(1+R))^(-1/gamma) + (1+R))
C2 <- (W - C1) * (1 + R)
points(c(C1,C2),pch=15,col="darkgrey",cex=1.3)
lines(c(C1,C2),col="darkgrey",lwd=2)
gamma <- 50
C1 <- W * (1 + R) / ((1/delta * 1/(1+R))^(-1/gamma) + (1+R))
C2 <- (W - C1) * (1 + R)
points(c(C1,C2),pch=17,col="grey",cex=1.3)
lines(c(C1,C2),col="grey",lwd=2)

legend("topleft",
       c(expression(paste(gamma," = 1.5"," (IES = 2/3)",sep="")),
         expression(paste(gamma," = 5"," (IES = 1/5)",sep="")),
         expression(paste(gamma," = 50"," (IES = 1/50)",sep=""))
       ),
       pt.bg = c("black","darkgrey","grey"),
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2,2), # line width
       pch=c(19,15,17),
       pt.cex = 1.3,
       col= c("black","darkgrey","grey"), # gives the legend lines the correct color and width
       seg.len = 4
)
```


The following code shows the time-varying conditional correlation between consumption and inflation.

## Empirical correlations between key macro-financial variables

This example reports empirical correlations between key macro-financial variables.

```{r fig_fredCorrel, fig.width=7, fig.height=6}
# Purpose: This example reports empirical correlations between key macro-financial variables
library(DTAM)
# Build consumption-growth and inflation series at a two-year horizon.
vec_dates <- as.Date(Data_Macro_US_monthly$date)
CONSO <- Data_Macro_US_monthly$RCONS
CPI   <- Data_Macro_US_monthly$CPIAUCSL

h <- 24
T <- length(CPI)
conso     <- 100*(12/h)*log(CONSO[(h+1):T]/CONSO[1:(T-h)])
inflation <- 100*(12/h)*log(CPI[(h+1):T]/CPI[1:(T-h)])

# Compute rolling correlations.
correla <- NaN*inflation
rwindow <- 12*7
for(t in (rwindow+1):(T-h)){
  correla[t] <- cor(conso[(t-rwindow+1):t],inflation[(t-rwindow+1):t])
}

par(mfrow=c(2,1))
par(plt=c(0.15,0.88,0.2,0.84))
layout(matrix(c(1,2)), heights=c(2, 2))

# Plot consumption growth and inflation.
plot(vec_dates[(h+1):T],conso,type='l',
     ylim=c(-8,15),col="black",lty=1,lwd=2,cex.axis=.8,cex=.8,las=1,
     ylab="in percent",xlab="",main="(a) Inflation and consumption growth",cex.main=.9)
grid()
abline(h=0,col="black",lty=3)
lines(vec_dates[(h+1):T],inflation,col="darkgrey",lwd=2)

legend("bottomleft",
       c("Consumption growth","Inflation rate"),
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2,2), # line width
       col=c("black","darkgrey"), # gives the legend lines the correct color and width
       cex=.8# size of the text
)
# Plot the rolling correlation together with term-premium estimates.
plot(vec_dates[(h+1):T],100*correla,col="black",type="l",
     lwd=3,ylim=c(-130,130),lty=2,
     ylab="Correlation, in percent",xlab="",cex.axis=.8,cex=.8,las=1,
     main="(b) Correlation between consumption growth and inflation",cex.main=.9)
grid()
abline(h=0,col="black",lty=3)
text(vec_dates[330],-15,labels="Supply shocks prevail",cex=.9)
text(vec_dates[330],+15,labels="Demand shocks prevail",cex=.9)

values.TP <- seq(-450,450,by=150)
location.ticks <- -values.TP/4.5
axis(4,at=-location.ticks,
     labels = paste(-values.TP,"",sep=""),
     col="darkgrey",col.axis="darkgrey",las=1,cex.axis=.9)
mtext("Term premium, in basis points", side=4, line=3,col="darkgrey")

lines(ACMTermPremium$date,-100/4.5*ACMTermPremium$ACMTP10,
      col="darkgrey",lwd=2)
lines(Data_Macro_US_monthly$date,
      -100/4.5*Data_Macro_US_monthly$THREEFYTP10,col="black",lwd=2,lty=1)

legend("topleft",
       title="Term premiums (rhs)",
       c("ACM","KW"),
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2,2), # line width
       seg.len = 3,
       col=c("darkgrey","black"), # gives the legend lines the correct color and width
       cex=.8# size of the text
)
legend("topright",
       c("Consumption-inflation correlation"),
       lty=c(2), # gives the legend appropriate symbols (lines)
       lwd=c(3), # line width
       seg.len = 3,
       col=c("black"), # gives the legend lines the correct color and width
       cex=.8# size of the text
)


```

The following code creates a figure that displays a simluated path for the short-term rate, and shows the term structure of interest rates at different point in times.

## Term structures under a CCAPM-based pricing kernel

This example simulates term structures implied by a CCAPM-based pricing kernel.

```{r fig_simulTS_CCAPM, fig.width=8, fig.height=4}
# Purpose: This example simulates term structures implied by a CCAPM-based pricing kernel
library(DTAM)
set.seed(123)

phi <- 0.8
mu <- 0.005
sigma.bar <- .01 # unconditional std dev of growth
sigma2  <- sigma.bar^2 * (1 - phi^2)
sigma.c <- sqrt(sigma2)
gamma   <- 10
delta   <- 0.99
# Determine specification of real short-term rate:
eta.0 <- -log(delta) + mu * gamma * (1 - phi) - 
  gamma^2 * sigma.c^2 / 2
eta.1 <- gamma * phi
# Specify the model:
psi.parameterization <- list(
  mu = matrix(mu * (1-phi)),
  Phi = matrix(phi),
  Sigma = matrix(sigma.c^2))
# Use reverse-order multi-horizon LT to price ZC bonds:
h = 10 # maximum maturity
u2 <- matrix(-gamma)
u1 <- matrix(-gamma)
AB <- reverse_MHLT(psi_GaussianVAR,u1,u2,H=h,psi.parameterization)
a.h <- -c(AB$A)/(1:h)
b.h <- - log(delta) + -c(AB$B)/(1:h)
# Under EH:
u2 <- matrix(- eta.1)
u1 <- matrix(0)
AB <- reverse_MHLT(psi_GaussianVAR,u1,u2,H=h,psi.parameterization)
a.h.EH <- -c(AB$A - eta.1)/(1:h)
b.h.EH <- eta.0 - c(AB$B)/(1:h)
# Simulate growth:
model <- list(mu=mu*(1-phi),
              Phi=as.matrix(phi),
              Sigma12=as.matrix(sigma.c))
T <- 100
x.sim <- simul_GaussianVAR(model,nb.sim=T)
r.t   <- (eta.0 + eta.1 * x.sim)

# Prepare plots:
par(mfrow=c(1,1))
par(plt=c(.15,.98,.1,.95))

plot(100*r.t,type="l",xlab="",ylab="interest rate, in percent",
     ylim=100*c(min(r.t),
                max(r.t)+.07),las=1,
     lwd=2,col="grey")
grid()
abline(h=0,lty=3)
for(t in seq(1,T,by=10)){
  lines(t+(1:h)-1,100*(b.h    + a.h    * x.sim[t]),col="black",lwd=2)
  lines(t+(1:h)-1,100*(b.h.EH + a.h.EH * x.sim[t]),col="black",lwd=2,lty=2)}
legend("topleft",
       c("Short-term rate","Term structure of real yields",
         "Term structure of real yields (without risk premiums)"),
       lwd=c(2),lty=c(1,1,2),col=c("grey","black","black"),seg.len = 4)
```

The following code produces a figure illustrating the real term premium.

## Decomposes the real term premium in the baseline CCAPM setting

This example decomposes the real term premium in the baseline CCAPM setting.

```{r fig_RTP, fig.width=8, fig.height=4}
# Purpose: This example decomposes the real term premium in the baseline CCAPM setting
library(DTAM)

# Load and merge nominal, real, and survey data.
data("SPF", package = "DTAM")
data("Data_Macro_US_quarterly", package = "DTAM")
data("YC_LW", package = "DTAM")

data <- merge(YC_LW,SPF,by="date")
data <- merge(data,DAT_GSW,by="date")

par(mfrow=c(1,2))
par(plt=c(.2,.95,.1,.85))

# Plot the slope of the real yield curve.
data_reduced <- subset(data,select=c(date,TIPSY10,TIPSY02))
data_reduced <- data_reduced[complete.cases(data_reduced),]
plot(data_reduced$date,data_reduced$TIPSY10 - data_reduced$TIPSY02,type="l",
     lwd=2,las=1,ylab="in percent",xlab="",
     main="(a) Slope of TIPS yield curve (2y-10y)",cex.main=.9)
grid()
abline(h=0,lty=3)

# Computation of model-free real term premium:
data_reduced <- subset(data,select=c(date,TIPSY10,yld_10y,CPI10,BILL10))
data_reduced <- data_reduced[complete.cases(data_reduced),]

BEIR10 <- data_reduced$yld_10y - data_reduced$TIPSY10
IRP10  <- BEIR10 - data_reduced$CPI10
NTP10  <- data_reduced$yld_10y - data_reduced$BILL10
RTP10  <- NTP10 - IRP10
plot(data_reduced$date,RTP10,type="b",
     lwd=2,las=1,ylab="in percent",xlab="",
     main="(b) Survey-based 10y real term premium",cex.main=.9)
grid()
abline(h=0,lty=3)
```


The following code produces a figure showing average real and inflation yield curves.

## Model-implied yield-curve moments with CCAPM restrictions

This example compares model-implied yield-curve moments with CCAPM restrictions.

```{r fig_TSMCCAPM, fig.width=9, fig.height=4}
# Purpose: This example compares model-implied yield-curve moments with CCAPM restrictions
library(DTAM)

values_sigma.pi.c <- c(0,-.03)
title_plot <- c(
  expression(paste("(a) ",sigma[pi*','*c]," = 0",sep="")),
  expression(paste("(a) ",sigma[pi*','*c]," = -0.03",sep=""))
)

par(plt=c(.2,.95,.25,.85))
par(mfrow=c(1,2))

pi_bar <- .02
phi_pi <- .8
sigma.pi   <- .01

count_plot <- 0
for(sigma.pi.c in values_sigma.pi.c){
  count_plot <- count_plot + 1
  # Determine specification of nominal short-term rate:
  eta.0.i <- -log(delta) + mu * gamma * (1 - phi) + pi_bar * 
    (1 - phi_pi) - (gamma*sigma.c + sigma.pi.c)^2/2 - sigma.pi^2/2
  eta.1.i <- matrix(c(gamma * phi,phi_pi),2,1)
  # Specify the model:
  psi.parameterization <- list(
    mu = matrix(c(mu*(1-phi),pi_bar*(1-phi_pi)),2,1),
    Phi = diag(c(phi,phi_pi)),
    Sigma = matrix(c(sigma.c^2,sigma.c*sigma.pi.c,sigma.c*
                       sigma.pi.c,sigma.pi^2+sigma.pi.c^2),2,2))
  # Use reverse-order multi-horizon LT to price ZC bonds:
  u2 <- matrix(c(-gamma,-1),2,1);u1 <- matrix(c(-gamma,-1),2,1)
  AB <- reverse_MHLT(psi_GaussianVAR,u1,u2,H=h,psi.parameterization)
  a.h.i <- -matrix(AB$A,2,h)/t(matrix((1:h),h,2))
  b.h.i <- -log(delta) + -c(AB$B)/(1:h)
  # Under EH:
  u2 <- matrix(- eta.1.i);u1 <- matrix(0,2,1)
  AB <- reverse_MHLT(psi_GaussianVAR,u1,u2,H=h,psi.parameterization)
  a.h.i.EH <- -(matrix(AB$A,2,h)-
                  matrix(eta.1.i,2,h))/t(matrix((1:h),h,2))
  b.h.i.EH <- eta.0.i - c(AB$B)/(1:h)
  # Average yield curves:
  avg.i.h    <- b.h.i    + matrix(c(mu,pi_bar),1,2) %*% a.h.i
  avg.i.h.EH <- b.h.i.EH + matrix(c(mu,pi_bar),1,2) %*% a.h.i.EH
  
  plot(100*c(avg.i.h),type="l",lwd=2,col="darkgrey",lty=1,
       ylim=100*c(.01,max(avg.i.h)+.01),
       main=title_plot[count_plot],
       xlab="maturity",ylab="yield, in percent",las=1);grid()
  lines(100*c(avg.i.h.EH),type="l",lwd=2,col="darkgrey",lty=2)
  lines(100*(b.h+a.h*mu),lwd=2,col="black")
  lines(100*(b.h.EH+a.h.EH*mu),lwd=2,col="black",lty=2)
  if(count_plot==1){
    legend("bottomleft",title="Type of yield:",
           c("Nominal yields",
             "Real yields"),
           lwd=c(2),lty=c(1,1),col=c("darkgrey","black"),seg.len = 3)
  }
  if(count_plot==2){
    legend("bottomleft",title="With or w/o term premiums:",
           c("with",
             "without (i.e., under EH)"),
           lwd=c(2),lty=c(1,2),col=c("black"),seg.len = 3)
  }
}
```


## Separates trend and cycle contributions to the real term premium

This example separates trend and cycle contributions to the real term premium.

```{r fig_RTP_TrendCycle, fig.width=9, fig.height=4}
# Purpose: This example separates trend and cycle contributions to the real term premium

mu_g    <- .02
rho_g   <- .9
rho_z   <- .8
sigma_g <- .002
sigma_z <- .02
delta   <- 1
gamma   <- 2

h <- 8

# Specify the model:
psi.parameterization <- list(
  mu = matrix(c(mu_g*(1-rho_g),0,0),3,1),
  Phi = matrix(c(rho_g,0,0,0,rho_z,1,0,0,0),3,3),
  Sigma = diag(c(sigma_g^2,sigma_z^2,0)))

all_sigma_g <- c(0,.005,.01)
all_lty     <- c(1,2,3)
all_col     <- c("black","black","black")
names_legend <- c(expression(paste(sigma[g],' = ',0,sep="")),
                  expression(paste(sigma[g],' = ',0.005,sep="")),
                  expression(paste(sigma[g],' = ',0.01,sep="")))

par(plt=c(.15,.95,.25,.95))
for(i in 1:length(all_sigma_g)){
  
  psi.parameterization$Sigma[1,1] <- all_sigma_g[i]^2

  # Use reverse-order multi-horizon LT to price ZC bonds:
  u1 <- matrix(c(-gamma,-gamma,gamma),3,1)
  AB <- reverse_MHLT(psi_GaussianVAR,u1,H=h,
                     psi.parameterization=psi.parameterization)
  a.h.r <- -matrix(AB$A,3,h)/t(matrix((1:h),h,3))
  b.h.r <- - log(delta) + -c(AB$B)/(1:h)
  # Average yield curves:
  yds    <- 100 * (b.h.r    + matrix(c(mu_g,0,0),1,3) %*% a.h.r)
  
  if(i==1){
    plot(1:h,yds,type="l",col=all_col[i],lty=all_lty[i],lwd=2,
         xlab="maturity",ylab="real interest rate (in percent)",las=1,
         ylim=c(3.6,4.1))
    grid()
  }else{
    lines(1:h,yds,col=all_col[i],lty=all_lty[i],lwd=2)
  }
}

abline(h=0,col="grey",lty=3)

legend("bottomleft",
       names_legend,
       lty=all_lty, # gives the legend appropriate symbols (lines)       
       lwd=2, # line width
       col=all_col, # gives the legend lines the correct color and width
       bg="white",
       seg.len = 4
)



```
