# Credit and liquidity risks {#CreditLiquidity}

## Standard framework

This example compares standard structural and reduced-form views of credit spreads. The empirical illustration combines a maturity-matched 20-year corporate credit spread with a rescaled business-loan loss indicator. The spread is the difference between the U.S. Treasury's 20-year High Quality Market (HQM) corporate spot rate and its 20-year Treasury Nominal Coupon-Issue (TNC) spot rate; the source workbooks and methodology are available from the [HQM](https://home.treasury.gov/data/treasury-coupon-issues-and-corporate-bond-yield-curve/corporate-bond-yield-curve) and [TNC](https://home.treasury.gov/data/treasury-coupon-issues-and-corporate-bond-yield-curves/treasury-coupon-issues) yield-curve pages. The credit-loss indicator uses the Federal Reserve Board's [business-loan charge-off rate](https://www.federalreserve.gov/releases/chargeoff/), rescaled so that its average expected loss corresponds to high-quality U.S. corporate bonds [@AmatoRemolona2003].^[Amato and Remolona [-@AmatoRemolona2003, Table 1] report expected losses of 0.61, 2.70, and 7.32 basis points for AAA-, AA-, and A-rated bonds in their longest maturity bucket; their equal-weight average is 3.54 basis points. Let $c_t$ denote the charge-off rate averaged over the quarters of year $t$, and let $\bar c$ be its 1986--2024 average. We set $\mathrm{EL}_t=(3.54\text{ basis points})c_t/\bar c$, so that the sample average of $\mathrm{EL}_t$ is exactly 3.54 basis points. We then set the physical default-intensity indicator to $\lambda_t=\mathrm{EL}_t/(1-\zeta)$, where $\zeta$ is the recovery rate assumed below. The companion workbook records the source observations and each transformation.]

```{r fig_standardCreditFedLoss, fig.width=5.5, fig.height=7}
# Purpose: This example compares standard structural and reduced-form views of credit spreads
library(DTAM)

public_spread_file <- "data/Treasury_HQM_TNC_20Y.xlsx"
credit_loss_file <- "data/FederalReserve_HQM_CreditLossProxy.xlsx"

if (!file.exists(public_spread_file)) {
  stop("Treasury spread data not found: ", public_spread_file, call. = FALSE)
}
if (!file.exists(credit_loss_file)) {
  stop("Credit-loss indicator not found: ", credit_loss_file, call. = FALSE)
}

spread_data <- readxl::read_excel(public_spread_file,
                                  sheet = "Annual 20Y spread",
                                  skip = 6)
spread_data <- data.frame(
  year = suppressWarnings(as.numeric(spread_data[[1]])),
  HQM_yld = suppressWarnings(as.numeric(spread_data[[2]])),
  TNC_yld = suppressWarnings(as.numeric(spread_data[[3]])),
  HQM_TNC_spd = suppressWarnings(as.numeric(spread_data[[4]]))
)
spread_data <- subset(spread_data,
                      year >= 1986 & year <= 2024 &
                        complete.cases(spread_data))

loss_data <- readxl::read_excel(credit_loss_file,
                                sheet = "Annual calibrated",
                                skip = 6)
loss_data <- data.frame(
  year = suppressWarnings(as.numeric(loss_data[[1]])),
  chargeoff_rate = suppressWarnings(as.numeric(loss_data[[2]])),
  expected_loss_bp = suppressWarnings(as.numeric(loss_data[[5]])),
  default_intensity_pct = suppressWarnings(as.numeric(loss_data[[6]]))
)
loss_data <- subset(loss_data,
                    year >= 1986 & year <= 2024 &
                      complete.cases(loss_data))

credit_data <- merge(spread_data, loss_data, by = "year", all = FALSE)
stopifnot(nrow(credit_data) == 39,
          !anyDuplicated(credit_data$year),
          all(credit_data$HQM_TNC_spd > 0),
          all(credit_data$default_intensity_pct >= 0))

RR <- .6
lambda       <- credit_data$default_intensity_pct/100
spd_HQM_TNC  <- credit_data$HQM_TNC_spd/100
stopifnot(max(abs(lambda*(1-RR) - credit_data$expected_loss_bp/10000)) < 1e-10)

# The observed spread and model bond both have a 20-year maturity.
H <- 20

make_models <- function(param){
  # Specify two-factor VARG model:
  mu_x       <- exp(param[1])
  mu_lambda  <- exp(param[2])
  rho_x      <- exp(param[3])/(1+exp(param[3]))
  rho_lambda <- exp(param[4])/(1+exp(param[4]))
  nu         <- exp(param[5])
  # prices of risk:
  alpha_x      <- exp(param[6])
  alpha_lambda <- exp(param[7])
  alpha_SDF <- matrix(c(alpha_x,
                        alpha_lambda),ncol=1)
  
  alpha = matrix(0,2,1)
  mu    = matrix(c(mu_x,mu_lambda),2,1)
  nu    = matrix(c(nu,0),2,1)
  beta  = matrix(c(rho_x/mu_x,(1-rho_lambda)/mu_lambda,0,rho_lambda/mu_lambda),2,2)
  
  model <- list(alpha=alpha,mu=mu,nu=nu,beta=beta)
  
  muQ    <- mu / (1 - mu * alpha_SDF)
  alphaQ <- alpha * muQ/mu
  nuQ    <- nu
  betaQ  <- beta * (matrix(muQ/mu,ncol=1) %*% matrix(1,1,dim(beta)[2]))
  
  modelQ <- list(alpha=alphaQ,mu=muQ,nu=nuQ,beta=betaQ)
  
  return(list(model = model,
              modelQ = modelQ,
              mu_x=mu_x,mu_lambda=mu_lambda,
              rho_x=rho_x,rho_lambda=rho_lambda,
              nu=nu,alpha_x=alpha_x,alpha_lambda=alpha_lambda))
}

loss4optim <- function(param){
  
  Models <- make_models(param)
  model  <- Models$model
  modelQ <- Models$modelQ
  
  xi0 <- 0
  xi1 <- matrix(0,length(model$alpha),1)
  kappa0 <- 0
  kappa1 <- matrix(c(0,1-RR),ncol=1)
  
  bond_prices <- compute_AB_classical(xi0=xi0, xi1=xi1,
                                      kappa0=kappa0, kappa1=kappa1,
                                      H=H, # maximum maturity
                                      psi=psi_VARG,psi.parameterization=modelQ)
  
  e <- 2 # e = 1 is for the risk-free yield
  b_spd <- bond_prices$b[1,e,H]
  a_spd <- matrix(bond_prices$a[,e,H],ncol=1)
  
  # Compute x to have a perfect fit of spreads:
  x <- (spd_HQM_TNC - b_spd - a_spd[2]*lambda)/a_spd[1]
  
  # Moments:
  mean_Y <- compute_uncondmean_VARG(model)
  Var_Y  <- compute_uncondvar_VARG(model)
  
  # Penalize negative x values:
  penalty_neg_x <- 10000000000*sum(abs(x[x<0]))
  
  # Penalty for distance between empirical and model-implied means:
  penalty_mean_distance_lambda <- 1000000000*abs(mean_Y[2] - mean(lambda))
  
  # penalty for average spread:
  avg_spd <- b_spd + sum(a_spd*mean_Y)
  penalty_mean_spd <- 1000000000*abs(avg_spd - mean(spd_HQM_TNC))
  
  # penalty for stdv of spreads:
  stdv_spd <- sqrt(t(a_spd) %*% Var_Y %*% a_spd)
  penalty_var_spd <- 10000*abs(stdv_spd - sd(spd_HQM_TNC))
  
  # penalty for stdv of lambda:
  a_lambda <- matrix(c(0,1),ncol=1)
  stdv_lambda <- sqrt(t(a_lambda) %*% Var_Y %*% a_lambda)
  penalty_var_lambda <- 10000*abs(stdv_lambda - sd(lambda))
  
  Phi_Y <- model$beta * matrix(model$mu,2,2)
  Var_Y_Y_1 <- rbind(cbind(Var_Y,Phi_Y%*%Var_Y),
                     cbind(Var_Y%*%t(Phi_Y),Var_Y))
  
  # penalty for autocorrelation of lambda:
  T <- length(spd_HQM_TNC)
  Autocor_lambda <- cor(lambda[2:T],lambda[1:(T-1)])
  a1 <- rbind(a_lambda,0*a_lambda)
  a2 <- rbind(0*a_lambda,a_lambda)
  cov_lambda_lambda_1 <- t(a1) %*% Var_Y_Y_1 %*% a2
  AC_lambda_model <- cov_lambda_lambda_1/stdv_lambda^2
  penalty_AC_lambda <- 1000*abs(AC_lambda_model - Autocor_lambda)
  
  # penalty for autocorrelation of spd:
  Autocor_spd <- cor(spd_HQM_TNC[2:T],spd_HQM_TNC[1:(T-1)])
  a1 <- rbind(a_spd,0*a_spd)
  a2 <- rbind(0*a_spd,a_spd)
  cov_spd_spd_1 <- t(a1) %*% Var_Y_Y_1 %*% a2
  AC_spd_model <- cov_spd_spd_1/stdv_spd^2
  penalty_AC_spd <- 1000*abs(AC_spd_model - Autocor_spd)
  
  # Compute loss:
  loss <- penalty_neg_x + 
    penalty_mean_distance_lambda +
    penalty_mean_spd + penalty_var_spd + penalty_var_lambda +
    penalty_AC_lambda + penalty_AC_spd
  
  return(loss)
}

# Initial values:
rho_x      <- .91
rho_lambda <- .2
mu_x       <- .0005
mu_lambda  <- .0025
nu         <- 2.7
alpha_x    <- 50
alpha_lambda <- 0.05
param <- log(c(mu_x,mu_lambda,rho_x,rho_lambda,nu,alpha_x,alpha_lambda))
param[3] <- log(rho_x/(1-rho_x))
param[4] <- log(rho_lambda/(1-rho_lambda))

for(i in 1:10){
  res.optim <- optim(par=param,
                     fn=loss4optim,
                     method="Nelder-Mead",
                     control=list(trace=FALSE,maxit=2000),hessian=FALSE)
}

param  <- res.optim$par
Models <- make_models(param)
model  <- Models$model
modelQ <- Models$modelQ


# Compute x_t ------------------------------------------------------------------
xi0 <- 0
xi1 <- matrix(0,length(model$alpha),1)
kappa0 <- 0
kappa1 <- matrix(c(0,1-RR),ncol=1)
bond_prices <- compute_AB_classical(xi0=xi0, xi1=xi1,
                                    kappa0=kappa0, kappa1=kappa1,
                                    H=H, # maximum maturity
                                    psi=psi_VARG,psi.parameterization=modelQ)
k <- 2 # k=1 is for the risk-free yield
b_spd <- bond_prices$b[1,k,H]
a_spd <- matrix(bond_prices$a[,k,H],ncol=1)
x <- (spd_HQM_TNC - b_spd - a_spd[2]*lambda)/a_spd[1]
Y <- cbind(x,lambda)

# Calculate Physical probabilities of default:
kappa1_lambda <- kappa1/(1-RR)
surviv_proba_P <- compute_AB_classical(xi0=xi0, xi1=xi1,
                                       kappa0=kappa0, kappa1=kappa1_lambda,
                                       H=H, # maximum maturity
                                       psi=psi_VARG,psi.parameterization=model)
surviv_proba_Q <- compute_AB_classical(xi0=xi0, xi1=xi1,
                                       kappa0=kappa0, kappa1=kappa1_lambda,
                                       H=H, # maximum maturity
                                       psi=psi_VARG,psi.parameterization=modelQ)

e <- 2 #   e = 1 is for the risk-free entity
H1 <- 5
PD_P_H1 <- 1 - exp(surviv_proba_P$B[1,e,H1] + Y %*% surviv_proba_P$A[,e,H1])
PD_Q_H1 <- 1 - exp(surviv_proba_Q$B[1,e,H1] + Y %*% surviv_proba_Q$A[,e,H1])
H2 <- 10
PD_P_H2 <- 1 - exp(surviv_proba_P$B[1,e,H2] + Y %*% surviv_proba_P$A[,e,H2])
PD_Q_H2 <- 1 - exp(surviv_proba_Q$B[1,e,H2] + Y %*% surviv_proba_Q$A[,e,H2])

# nb_periods <- 200
# sim_Y <- simul_VARG(Models$model, nb_periods = nb_periods)
# plot(sim_Y[2,],type="l",lwd=2)
# lines(sim_Y[1,],col="red",lwd=2,lty=3)

par(mfrow=c(3,1))
par(plt=c(.1,.95,.2,.85))

plot(credit_data$year,credit_data$default_intensity_pct,
     ylim=c(0,1.2*max(c(credit_data$default_intensity_pct,
                        credit_data$HQM_TNC_spd))),col="white",
     xlab="",ylab="in percent",las=1,
     main=expression(paste("(a) Default-intensity indicator and 20-year credit spread",sep="")))
grid()
lines(credit_data$year,credit_data$default_intensity_pct,lwd=2,lty=2)
lines(credit_data$year,credit_data$HQM_TNC_spd,col="dark grey",type="b",lwd=2)

legend("topleft",
       legend=c(expression(paste("Default-intensity indicator",sep="")),
                expression(paste("20Y HQM-TNC spread",sep=""))),
       lty=c(2,NaN),
       pch=c(NaN,1),
       col=c("black","dark grey"),
       lwd = 2,bty = "n",seg.len = 3)

plot(credit_data$year,100*lambda,
     ylim=c(0,1.2*max(c(100*lambda,100*x))),col="white",
     xlab="",ylab="in percent",las=1,
     main=expression(paste("(b) Factors ",x[t]," and ",lambda[t],sep="")))
grid()
lines(credit_data$year,100*lambda,lwd=2,type="b",pch=3)
lines(credit_data$year,100*x,col="dark grey",lwd=2,type="b",pch=0)

legend("topleft",
       legend=c(expression(x[t]),expression(lambda[t])),
       lty=c(NaN,NaN),
       pch=c(0,3),
       col=c("dark grey","black"),
       lwd = 2,bty = "n")

plot(credit_data$year,100*lambda,
     ylim=c(0,120*max(PD_Q_H2)),col="white",
     xlab="",ylab="in percent",las=1,
     main=expression(paste("(c) Physical and risk-neutral probabilities of default (PD)",sep="")))
grid()
lines(credit_data$year,100*PD_P_H1,lwd=2,lty=1)
lines(credit_data$year,100*PD_Q_H1,lwd=3,lty=3,lend="round")
lines(credit_data$year,100*PD_P_H2,lwd=2,lty=1,col="dark grey")
lines(credit_data$year,100*PD_Q_H2,lwd=3,lty=3,col="dark grey",lend="round")

legend("topleft",
       legend=c("5-year physical PD",
                "5-year risk-neutral PD",
                "10-year physical PD",
                "10-year risk-neutral PD"),
       lty=c(1,3,1,3),
       col=c("black","black","dark grey","dark grey"),
       lwd = c(2,3,2,3),bty = "n",seg.len = 3)
```




## Credit ratings



This example illustrates bond pricing and migration dynamics in a ratings-based framework.

The spectral construction preserves unit row sums and the absorbing default state, but it does not impose non-negativity of every entry of the transition matrix. Under the present calibration, violations larger than $5\times10^{-5}$ have probability of about $3\times10^{-7}$ per period.

```{r fig_Ratings, fig.width=8, fig.height=4.5}
# Purpose: This example illustrates bond pricing and migration dynamics in a ratings-based
# framework
library(DTAM)

nb.sim <- 150

# Baseline matrix of transition probabilities:
Pi <- t(matrix(c(.95,.03,.02,.0,
                 .03,.94,.025,.005,
                 .05,.09,.85,.01,
                 0,0,0,1),4,4))
K <- dim(Pi)[1] # number of ratings
eig_res <- eigen(Pi)
V <- eig_res$vectors
V <- V[,K:1]
eig_values <- eig_res$values[K:1] # ensure 1 is the last eigenvalue.

# Specify model ----------------------------------------------------------------

#  Credit ratings:
model <- list()
model$xi0 <- .02
model$xi1 <- matrix(c(.01,0),ncol=1)
model$kappa0 <- matrix(-log(eig_values),nrow=1)
model$kappa1 <- rbind(0,.2*model$kappa0)
model$V <- V

# Specify y_t's dynamics:
mu <- matrix(0,2,1)
rho <- .9
Phi <- rho*diag(2)
Sigma <- (1-rho^2) * diag(2) # This ensures that Var(y_{i,t})=1.
model$psi.parameterization <- list(mu = mu, Phi = Phi, Sigma = Sigma)

# Simulate Y:
Y <- simul_GaussianVAR(model$psi.parameterization,
                nb.sim = nb.sim)


# Price bonds: -----------------------------------------------------------------
res_ratings <- compute_bond_price_Ratings(model,Y,H=10,psi=psi_GaussianVAR)


# Simulate ratings: ------------------------------------------------------------
# flag remains 1 as long as:
# - not all ratings visited
# - a default occur
# - short-lived rating (less than 2 periods)
# <=> max(tau) == K-1
flag <- 1
while(flag==1){
  tau1 <- simul_rating_migration(model,Y,tau_ini=1)
  if((max(tau1)==K-1)&( sum(abs(diff(tau1[1:(nb.sim-1)])*diff(tau1[2:nb.sim])))==0 )){
    flag <- 0
  }
}
flag <- 1
while(flag==1){
  tau2 <- simul_rating_migration(model,Y,tau_ini=1)
  if((max(tau2)==K-1)&( sum(abs(diff(tau2[1:(nb.sim-1)])*diff(tau2[2:nb.sim])))==0 )){
    flag <- 0
  }
}



# Prepare plot -----------------------------------------------------------------

h <- 5

spreads <- cbind(res_ratings$Risky_bond_spreads[,,h],10)
spds1 <- matrix(NaN,dim(Y)[1],1)
spds2 <- matrix(NaN,dim(Y)[1],1)
for(t in 1:dim(Y)[1]){
  spds1[t] <- spreads[t,tau1[t]]
  spds2[t] <- spreads[t,tau2[t]]
}

par(mfrow=c(1,1))
par(plt=c(.1,.95,.2,.95))
plot(10000*res_ratings$Risky_bond_spreads[,1,h],type="l",col="grey",lty=1,lwd=3,
     ylim=10000*c(0,1.3*max(res_ratings$Risky_bond_spreads[,,h])),
     las=1,xlab="time",ylab="spread (in basis points)")
lines(10000*res_ratings$Risky_bond_spreads[,2,h],type="l",col="grey",lty=2,lwd=3)
lines(10000*res_ratings$Risky_bond_spreads[,3,h],type="l",col="grey",lty=3,lwd=3)

lines(10000*spds1,lwd=3)
lines(10000*spds2,lwd=3,lty=3)

legend("topleft",
       legend=c(paste("Rating: ",LETTERS[1:(K-1)],sep="")),
       lty=c(1:(K-1)),
       col=c(rep("grey",K-1)),
       lwd = c(rep(3,K-1)),bty = "n")

legend("topright",
       legend=c("Specific entity 1",
                "Specific entity 2"),
       lty=c(1,3),
       col=c("black","black"),
       lwd = c(3,3),bty = "n")


```


## VARG

This example simulates default intensities and state variables in the VARG credit model.

```{r fig_VARGCredit_simul, fig.width=8, fig.height=6}
# Purpose: This example simulates default intensities and state variables in the VARG credit
# model

library(DTAM)

nb_periods <- 150

# Set the dimensions of the VARG credit system.
# Number of entities (hence dimension of delta):
n.e <- 2
# Number of elements in y:
n.y <- 2 # in the context of this example: two entity-specific factors
# Dimension of w:
n.w <- n.y + n.e
# (w_t = [y_t',delta_t']')


beta.delta.y <- .2
beta.delta   <- 1

# Calibrate the common state factors and the default indicators.
mu.y <- .004
nu.y <- .5
mu.delta <- 1

rho.y <- .8
beta.y <- rho.y/mu.y

alpha <- matrix(0,n.w,1)
mu    <- matrix(c(rep(mu.y,n.y),
                  rep(mu.delta,n.e)),n.w,1)
# mu[2] <- .75*mu[1]
nu    <- matrix(c(rep(nu.y,n.y),
                  rep(0,n.e)),n.w,1)
beta  <- matrix(0,n.w,n.w)
beta[1,1] <- beta.y
beta[2,2] <- beta.y
beta[3,1] <- beta.delta.y
beta[4,2] <- beta.delta.y
# beta[3,4] <- beta.delta
beta[4,3] <- beta.delta
model <- list(
  alpha = alpha,
  mu = mu,
  nu = nu,
  beta = beta
)

# Simulate state variables and default indicators.
w0 <- matrix(0,n.w,1) # Initial state variable

simul_w <- simul_VARG(model,
                      nb_periods=nb_periods,w0=w0)


# Plot the two entity-specific state factors.
par(mfrow=c(2,1))
par(plt=c(.1,.95,.15,.8))

plot(simul_w[1,],
     type="l",ylim=c(0,max(simul_w[1:2,])),lwd=2,
     xlab="",ylab="",
     main=expression(paste("(a) Simulated ",y[1*','*t]," and ",y[2*','*t],sep="")),
     las=1)
grid()
lines(simul_w[2,],col="dark grey",lwd=2)

legend("topleft",
       legend=c(
         expression(paste(y[1*','*t],sep="")),
         expression(paste(y[2*','*t],sep=""))),
       lty=c(1,1),
       col=c("black","dark grey"),
       lwd=c(2,2),bty = "n")

# Reset realized defaults before computing conditional default probabilities.
# Remove defaults:
simul_w[(n.y+1):n.w,] <- 0
# plot(simul_w[3,],type="l",
#      ylim=c(0,max(simul_w[3:4,])))
# lines(simul_w[4,],col="dark grey",lwd=2)

h <- 5
res_PD <- compute_proba_def_VARG(model,H=h,indic_delta1 = n.y+1,
                                 X=t(simul_w))

# Plot default probabilities implied by the simulated state variables.
plot(res_PD$PD[,1,h],
     type="l",ylim=c(0,max(simul_w[1:2,])),lwd=2,
     xlab="",ylab="",
     main=expression(paste("(b) Probability of default (",h,"=5)",sep="")),
     las=1)
grid()
lines(res_PD$PD[,2,h],col="dark grey",lwd=2)

legend("topleft",
       legend=c(
         expression(paste("e = 1",sep="")),
         expression(paste("e = 2",sep=""))),
       lty=c(1,1),
       col=c("black","dark grey"),
       lwd=c(2,2),bty = "n")

```

### CDS spreads vary with the state variables in the VARG credit model

This example shows how CDS spreads vary with the state variables in the VARG credit model.

```{r fig_VARGCredit_CDS, fig.width=8, fig.height=4}
# Purpose: This example shows how CDS spreads vary with the state variables in the VARG credit
# model

H <- 20

# Specification of the short-term rate:
model$xi0 <- 0
model$xi1 <- matrix(rep(0,n.w),ncol=1)

# Specification of the recovery rate:
# ( RR_t = exp(- a.0 - a.1'w_t) )
model$mu_R1 <- matrix(0,n.y+n.e,n.e) # one column = one entity
model$mu_R0 <- rep(100,n.e,1)

# SDF specification:
kappa <- 0.3
model$alpha_w <- matrix(c(rep(0,n.y),rep(kappa,n.e)),n.w,1)

# Determine risk-neutral dynamics:
modelQ <- make_VARG_Q(model)

# Determine values of state vector where CDS evaluated:
Ew <- compute_uncondmean_VARG(model)
w <- Ew
w[(n.y+1):n.w] <- 0 # no default
w <- cbind(w,w)
w[,2] <- 2*w[,1]


# CDS pricing ------------------------------------------------------------------
res_cds <- prices_CDS_RFV_VARG(modelQ,H=H,indic_delta1 = n.y+1,
                               X=t(w))
res_cds_P <- prices_CDS_RFV_VARG(model,H=H,indic_delta1 = n.y+1,
                                 X=t(w))
# Add FX -----------------------------------------------------------------------
model_FX <- model

# Specification of exchange rate:
# (increase in s = depreciation of domestic currency)
model_FX$mu_s1 <- matrix(0,n.y+n.e,1) # one column = one entity
model_FX$mu_s1[(n.y+1):n.w] <- .05
model_FX$mu_s0 <- 0

# Determine risk-neutral dynamics:
model_FXQ <- make_VARG_Q(model_FX)

res_cds_FX <- prices_CDS_RFV_VARG(model_FXQ,H=H,indic_delta1 = n.y+1,
                                  X=t(w))
res_cds_FXP <- prices_CDS_RFV_VARG(model_FX,H=H,indic_delta1 = n.y+1,
                                   X=t(w))
# ------------------------------------------------------------------------------


par(mfrow=c(1,2))
par(plt=c(.2,.95,.2,.85))

for(e in 1:2){
  if(e==1){
    maint <- expression(paste("(a) Entity e = 1",sep=""))
  }else{
    maint <- expression(paste("(b) Entity e = 2",sep=""))
  }
  plot(10000*res_cds_FX$CDS_spreads[1,e,],type="l",
       ylim=10000*c(0,1.3*max(res_cds_FX$CDS_spreads[1,e,])),lwd=2,
       xlab="maturity",ylab="CDS spread, in basis points",
       main=maint)
  grid()
  lines(10000*res_cds_FXP$CDS_spreads[1,e,],type="l",lwd=3,lty=3)
  
  lines(10000*res_cds$CDS_spreads[1,e,],type="l",lwd=2,lty=1,col="darkgrey")
  lines(10000*res_cds_P$CDS_spreads[1,e,],type="l",lwd=3,lty=3,col="darkgrey")
  
  if(e==1){
    legend("bottomright",
           # legend=c(
           #   expression(paste(y[i*','*t]," = E(",y[i*','*t],")",sep="")),
           #   expression(paste(y[i*','*t]," = ",2%*%E,"(",y[i*','*t],")",sep=""))
           # ),
           legend=c("CDS denom.: domestic currency",
                    "CDS denom.: foreign currency"),
           lty=c(1,1),
           col=c("dark grey","black"),
           lwd=c(2,2),bty = "n")
    
    legend("topright",
           c("Risk-neutral measure",
             "Physical measure"),
           lty=c(1,3),
           col=c("black","black"),
           lwd=c(2,3),bty = "n")
  }
}

# for(e in 1:2){
#   if(e==1){
#     maint <- expression(paste("(c) Entity e = 1 (foreign currency)",sep=""))
#   }else{
#     maint <- expression(paste("(d) Entity e = 2 (foreign currency)",sep=""))
#   }
#   plot(10000*res_cds_FX$CDS_spreads[1,e,],type="l",
#        ylim=10000*c(0,max(res_cds_FX$CDS_spreads[,e,])),lwd=2,
#        xlab="maturity",ylab="CDS spread, in basis points",
#        main=maint)
#   grid()
#   lines(10000*res_cds_FXP$CDS_spreads[1,e,],type="l",lwd=2,lty=3)



#   lines(10000*res_cds_FX$CDS_spreads[2,e,],type="l",lwd=2,lty=1,col="dark grey")
#   lines(10000*res_cds_FXP$CDS_spreads[2,e,],type="l",lwd=2,lty=3,col="dark grey")
# }


```



## Top-down approach

This example illustrates clustering and default-count dynamics in the top-down credit model.

```{r fig_TopDown_clustering, fig.width=8, fig.height=5}
# Purpose: This example illustrates clustering and default-count dynamics in the top-down credit
# model
library(DTAM)

J <- 2 # Number of segments

# Specification of y_t process:
rho <- .9
mu.y <- .002
beta.yy <- rho/mu.y
nu.y <- .25

kappa <- .2

beta_A <- 100
beta_B <- 10
zeta_A <- 0
zeta_B <- 0.45

model_no_cluster <- list(
  mu.y = mu.y,
  nu.y = nu.y,
  alpha.y = 0,
  beta.yy = beta.yy,
  beta.yn = matrix(0,1,J),
  alpha.n = rep(0,J),
  beta.nn = matrix(zeta_A,J,J),
  beta.ny = matrix(beta_A,J,1),
  n_w = J + 1,
  xi0 = 0,
  xi1 = matrix(c(0,rep(0,J)),J+1,1),
  alpha.m = matrix(c(0,kappa,0),J+1,1),
  I = rep(100,J)
)

nb_periods <- 200
sim_TopDown <- simul_TopDown(model_no_cluster,nb_periods=nb_periods)
W <- sim_TopDown$all_w

# Construct selection vectors:
e1 <- matrix(0,J+1,1)
e1[1+1] <- 1
e2 <- matrix(0,J+1,1)
e2[1+2] <- 1

H <- 5
max.N <- 20

for(indic_cluster in 0:1){
  if(indic_cluster==1){
    model_cluster <- model_no_cluster
    # Specification with clustering:
    model_cluster$beta.nn <- matrix(zeta_B,J,J)
    model_cluster$beta.ny <- matrix(beta_B,J,1)
    model <- model_cluster
  }else{
    model <- model_no_cluster
  }
  
  # Compute first and second moments of w_t:
  EV  <- compute_expect_variance(psi_w_TopDown,model,du=1e-5)
  EVQ <- compute_expect_variance(psiQ_w_TopDown,model,du=1e-5)
  
  # Situation where the value of y_t differs:
  # W <- cbind(EV$Ew,
  #            EV$Ew + 1*matrix(sqrt(diag(EV$Vw)),ncol=1))
  
  # Situation where the value of n_{2,t} differs:
  W <- cbind(EV$Ew,EV$Ew)
  
  W[2:(J+1),] <- 0 # set n to zeros.
  W[J+1,2] <- 2
  
  res <- compute_G(model,
                   W = W,
                   gamma = matrix(0,J+1,1),
                   v = e1,
                   b.matrix = t(matrix(.5+(0:max.N),max.N+1,H)),
                   H = H,
                   indic_Q = FALSE,
                   max_x = 500,
                   dx_statio = .2,
                   min_dx = 10^(-5), 
                   nb_x1 = 500)
  Probas <- cbind(res[,1,H],
                  res[,2:(max.N+1),H] - res[,1:(max.N),H])
  
  if(indic_cluster==0){
    Probas_no_cluster <- Probas
  }else{
    Probas_cluster <- Probas
  }
}

par(mfrow=c(1,dim(W)[2]))
par(plt=c(.2,.95,.17,.88))

for(i in 1:dim(W)[2]){
  if(i==1){
    main.t <- expression(paste("(a) ",n[1*','*t]," = 0 and ",n[2*','*t]," = 0",sep=""))
  }else{
    main.t <- expression(paste("(b) ",n[1*','*t]," = 0 and ",n[2*','*t]," = 2",sep=""))
  }
  plot(0:max.N,Probas_no_cluster[i,],type="b",pch=19,lwd=2,cex=1.2,
       ylim=c(0,1.2*max(Probas_no_cluster[i,],
                        Probas_cluster[i,])),
       main=main.t,
       xlab="Number of defaults",ylab="probability")
  lines(0:max.N,Probas_cluster[i,],type="b",pch=24,lwd=2,cex=1.2,
        bg="white",col="dark grey")
  grid()
  
  if(i==1){
    legend("topright",
           legend=c(
             "Model A (w/o clustering)",
             "Model B (w/ clustering)"),
           lty=c(NaN,NaN),
           pch = c(19,24),
           col=c("black","dark grey"),
           pt.bg = c("black","white"),
           lwd=c(2,2),bty = "n")
  }
}


```


### Prices CDS index spreads in the top-down model

This example prices CDS index spreads in the top-down model.

```{r fig_TopDown_CIS, fig.width=8, fig.height=4}
# Purpose: This example prices CDS index spreads in the top-down model
H <- 10

# Compute CDS spreads without default clustering.
CDS_no_cluster   <- compute_CDS_TopDown(model_no_cluster,
                                        H=H,
                                        W=W,
                                        RR=0)
CDS_P_no_cluster <- compute_CDS_TopDown(model_no_cluster,
                                        H=H,
                                        W=W,
                                        RR=0,
                                        indic_Q = FALSE)
# Compute CDS spreads with default clustering.
CDS_cluster   <- compute_CDS_TopDown(model_cluster,H=H,
                                     W = W,
                                     RR=0)
CDS_P_cluster <- compute_CDS_TopDown(model_cluster,H=H,
                                     W = W,
                                     RR=0,
                                     indic_Q = FALSE)

par(mfrow=c(1,2))
par(plt=c(.2,.95,.2,.85))

# Plot both models under two initial default-count states.
for(value_of_w in 1:2){
  if(value_of_w==1){
    main.t <- expression(paste("(a) ",n[1*','*t]," = 0 and ",n[2*','*t]," = 0",sep=""))
  }else{
    main.t <- expression(paste("(b) ",n[1*','*t]," = 0 and ",n[2*','*t]," = 2",sep=""))
  }
  
  plot(1:H,10000*CDS_cluster[value_of_w,1,],type="l",lwd=2,
       ylim=10000*c(0,1.4*max(CDS_cluster[,,],
                              CDS_no_cluster[,,])),
       main=main.t,
       xlab="maturity",ylab="CDS spread (in bps)",las=1)
  grid()
  
  points(1:H,10000*CDS_P_cluster[value_of_w,1,],lwd=2,cex=1,pch=3)
  lines(1:H,10000*CDS_cluster[value_of_w,2,],lwd=2,cex=1,lty=3)
  
# Overlay the no-clustering benchmark in grey.
  points(1:H,10000*CDS_P_no_cluster[value_of_w,1,],lwd=2,cex=1,pch=3,
         col="dark grey")
  lines(1:H,10000*CDS_no_cluster[value_of_w,2,],lwd=2,cex=1,lty=3,
        col="dark grey")
  lines(1:H,10000*CDS_no_cluster[value_of_w,1,],lwd=2,cex=1,lty=1,
        col="dark grey")
  
  if(value_of_w==1){
    legend("topleft",
           legend=c(
             "Segment 1",
             "Segment 2",
             "Under P (Segment 1 or 2)"),
           lty=c(1,3,NaN),
           pch=c(NaN,NaN,3),
           col=c("black"),
           lwd=c(2,2,2),bty = "n")
  }else{
    legend("topleft",
           legend=c(
             "Model A (w/o clustering)",
             "Model B (w/ clustering)"),
           lty=c(NaN),
           pch=c(15),
           col=c("black","dark grey"),
           lwd=c(2),bty = "n")
    
  }
}


```


### Tranche pricing across attachment points in the top-down framework

This example studies tranche pricing across attachment points in the top-down framework.

```{r fig_TopDown_Tranche, fig.width=8, fig.height=4}
# Purpose: This example studies tranche pricing across attachment points in the top-down
# framework

# Maximum maturity of tranche products:
H <- 10

# Segment:
j <- 1

# Under P or Q:
indic_Q <- TRUE

vector.of.a <- c(0,.03,.06)
tranche_no_cluster <- compute_tranche(model_no_cluster,
                                      W = W,
                                      RR = 0.5,
                                      indic_Q = indic_Q,
                                      j=j,H=H,
                                      vector.of.a,
                                      max_x = 500,
                                      dx_statio = .2,
                                      min_dx = 10^(-5), 
                                      nb_x1 = 500)
tranche_cluster <- compute_tranche(model_cluster,
                                   W = W,
                                   RR = 0.5,
                                   indic_Q = indic_Q,
                                   j=j,H=H,
                                   vector.of.a,
                                   max_x = 500,
                                   dx_statio = .2,
                                   min_dx = 10^(-5), 
                                   nb_x1 = 500)


par(mfrow=c(1,2))
par(plt=c(.2,.95,.2,.85))

for(value_of_w in 1:2){
  if(value_of_w==1){
    main.t <- expression(paste("(a) ",n[1*','*t]," = 0 and ",n[2*','*t]," = 0",sep=""))
  }else{
    main.t <- expression(paste("(b) ",n[1*','*t]," = 0 and ",n[2*','*t]," = 2",sep=""))
  }
  
  tranche <- 1
  plot(tranche_no_cluster$tranche_prices[value_of_w,tranche,],
       ylim=c(0,1.3*max(tranche_no_cluster$tranche_prices[,,],
                        tranche_cluster$tranche_prices[,,])),
       type="l",las=1,lwd=2,col="white",main=main.t,
       xlab="maturity",ylab="tranche price")
  grid()
  
  for(tranche in 1:2){
    lines(tranche_no_cluster$tranche_prices[value_of_w,tranche,],
          lty=tranche,col="black",lwd=2)
    lines(tranche_cluster$tranche_prices[value_of_w,tranche,],
          lty=tranche,col="dark grey",lwd=2)
  }
  
  if(value_of_w==1){
    legend("topleft",
           legend=c(
             "Equity tranche (0%-3%)",
             "Second tranche (3%-6%)",
             "",
             "Model A (w/o clustering)",
             "Model B (w/ clustering)"),
           lty=c(1,2,NaN,NaN,NaN),
           pch=c(NaN,NaN,NaN,15,15),
           col=c("black","black","white","black","dark grey"),
           lwd=c(2),bty = "n")
  }
}



```
