# Recursive preferences {#recursive}




## External habits amplify variation in marginal utility

This example illustrates how external habits amplify variation in marginal utility.

```{r fig_Habits1, fig.width=6, fig.height=3}
# Purpose: This example illustrates how external habits amplify variation in marginal utility

# Set the habit-preference parameters.
delta <- 1
lambda <- .49
rho <- .5
gamma <- 3

r.f <- c(rep(0.07,5),c(-.2,-.2,-.2))

# Simulate scenario A.
T <- length(r.f)
C <- 1
C.t <- C
X <- lambda * C /(1-rho)
X.t <- X
U <- delta * (C.t - X.t)^(1 - gamma)/(1 - gamma)
U.no.habit <- delta * (C.t)^(1 - gamma)/(1 - gamma)
for(t in 1:(T)){
  C.t <-  (1/(1+r.f[t]))^(-1/gamma) * (C.t - X.t) + rho * X.t + lambda * C.t
  C <- c(C,C.t)
  X.t <- rho * X.t + lambda * C.t
  X <- c(X,X.t)
  U <- U + delta * (C.t - X.t)^(1 - gamma)/(1 - gamma)
  U.no.habit <- U.no.habit + delta * (C.t)^(1 - gamma)/(1 - gamma)
}
U1 <- U
U1.no.habit <- U.no.habit
C1 <- C
X1 <- X
R.f.1 <- r.f

# Simulate scenario B.
r.f <- c(rep(0.025,8))

T <- length(r.f)
C <- 1
C.t <- C
X <- lambda * C /(1-rho)
X.t <- X
U <- delta * (C.t - X.t)^(1 - gamma)/(1 - gamma)
U.no.habit <- delta * (C.t)^(1 - gamma)/(1 - gamma)
for(t in 1:(T)){
  C.t <-  (1/(1+r.f[t]))^(-1/gamma) * (C.t - X.t) + rho * X.t + lambda * C.t
  C <- c(C,C.t)
  X.t <- rho * X.t + lambda * C.t
  X <- c(X,X.t)
  U <- U + delta * (C.t - X.t)^(1 - gamma)/(1 - gamma)
  U.no.habit <- U.no.habit + delta * (C.t)^(1 - gamma)/(1 - gamma)
}
U2 <- U
U2.no.habit <- U.no.habit
C2 <- C
X2 <- X
R.f.2 <- r.f

# Plot interest rates, consumption, and surplus consumption.
par(mfrow=c(1,3))
par(plt=c(.2,.9,.2,.85))

plot(R.f.1,col="black",lwd=2,pch=19,ylim=c(-.25,.2),las=1,
     main=expression(paste("(a) Short-term interest rate ",
                           tilde(R)[f*","*t], sep = "")),
     xlab="time",ylab="",type="b")
grid()
lines(R.f.2,col="darkgrey",pch=15,cex=1.2,type="b",lwd=2)

legend("topright",
       c("Scenario A","Scenario B"),
       lty=c(1,1),
       lwd=c(2,2),
       pch=c(19,15),
       pt.cex=c(1,1.2),
       col=c("black","darkgrey"),
       seg.len = 4
)

plot(C1,col="black",lwd=2,pch=19,ylim=c(.995,1.005),las=1,
     main=expression(paste("(b) Consumption ",C[t],sep="")),xlab="time",ylab="",
     type="b")
grid()
lines(C2,col="darkgrey",pch=15,cex=1.2,lwd=2,type="b")

plot((C1-X1)/C1,col="black",lwd=2,pch=19,ylim=c(0.018,0.022),las=1,
     main=expression(paste("(c) ",S[t]," = ",(C[t]-X[t])/C[t],sep="")),xlab="time",ylab="",type="b")
grid()
lines((C2-X2)/C2,col="darkgrey",pch=15,cex=1.2,lwd=2,type="b")

```



## Exponentially smoothed consumption growth enters recursive preferences

This example shows how exponentially smoothed consumption growth enters recursive preferences.

```{r fig_Dcexponentialsommothing, fig.width=8, fig.height=6}
# Purpose: This example shows how exponentially smoothed consumption growth enters recursive
# preferences
# Simulate a consumption-growth path.
mu     <- .015
sigma  <- .02
nb_sim <- 300
p <- .00
sigma_crisis <- .0
eta_crisis <- .1
dc <- mu + sigma * rnorm(nb_sim) + 
  (runif(nb_sim)<p)*(- eta_crisis + sigma_crisis*rnorm(nb_sim))
c <- cumsum(dc)
C <- exp(c)

# Specify surplus-consumption smoothing parameters.
all_rho   <- c(.6,.97)
all_alpha <- c(1,1)
all_alpha_tilde <- c(.2,.2)

start.date <- 100
end.date   <- nb_sim

par(mfrow=c(2,1))
par(plt=c(.15,.95,.15,.83))
plot(C[start.date:end.date],type="l",lwd=2,
     ylim=c(1 - max(all_alpha_tilde),max(C)),
     main=expression(paste("(a) Consumption, ",C[t],sep="")),
     las=1,xlab="time",ylab="",col="darkgrey")
grid()

# Construct habit-adjusted consumption for each smoothing parameter.
all_S <- NULL
for(i in 1:length(all_rho)){
  lambda_t <- mu
  lambda   <- lambda_t
  for(t in 2:nb_sim){
    lambda_t <- all_rho[i] * lambda_t + (1 - all_rho[i]) * c[t-1]
    lambda <- c(lambda,lambda_t)
  }
  S <- all_alpha_tilde[i] * exp(all_alpha[i] * (c - lambda))
  all_S <- cbind(all_S,S)
  X <- C * (1 - S)
  lines(X[start.date:end.date],lwd=2,lty=i)
}
legend("topleft",
       c(expression(C[t]),
         expression(paste(X[t]," for ",rho," = 0.60",sep="")),
         expression(paste(X[t]," for ",rho," = 0.97",sep=""))),
       lty=c(1,1:length(all_rho)),
       lwd=2, # line width
       col=c("darkgrey",rep("black",length(all_rho))),
       seg.len = 3
)

# Plot the surplus-consumption ratio.
plot(log(all_S[start.date:end.date,1]),type="l",lwd=2,
     main=expression(paste("(b) Consumption surplus, ",S[t],sep="")),
     las=1,xlab="time",ylab="",ylim=c(min(log(all_S[start.date:end.date,])),
                                      .25+max(log(all_S[start.date:end.date,]))))
grid()
lines(log(all_S[start.date:end.date,2]),lwd=2,lty=2)

legend("topleft",
       c(expression(paste(log(S[t])," for ",rho," = 0.60",sep="")),
         expression(paste(log(S[t])," for ",rho," = 0.97",sep=""))),
       lty=c(1:length(all_rho)),
       lwd=2, # line width
       col=c(rep("black",length(all_rho))),
       seg.len = 3
)

# Compute state-space objects associated with the smoothing specification.
for(i in 1:length(all_rho)){
  rho   <- all_rho[i]
  alpha <- all_alpha[i]
  Phi <- matrix(c(0,1-rho,0,rho),2,2)
  Sigma <- matrix(c(sigma,0),2,1)
  V <- matrix(solve(diag(4)-Phi %x% Phi)%*%c(Sigma%*%t(Sigma)),2,2)
  
  vec_gamma_dc         <- matrix(c(-gamma,0),2,1)
  vec_gamma_dc_dlambda <- matrix(c(-gamma*(1+alpha),gamma*alpha),2,1)
}
```



## The fixed-point resolution behind Epstein-Zin preferences

This example illustrates the fixed-point resolution behind Epstein-Zin preferences.

```{r fig_EZresolution, fig.width=6, fig.height=3}
# Purpose: This example illustrates the fixed-point resolution behind Epstein-Zin preferences

# Compare the fixed-point implication across elasticity parameters.
par(mfrow=c(1,3))
par(plt=c(.2,.95,.2,.8))

for(rho in c(.5,.9,1.5)){
  
  all.gamma.values <- seq(.1001,6,by=.1)
  all.C.minus.D.values <- NULL
  
  # Compute the utility difference over risk-aversion values.
  for(gamma in all.gamma.values){
    c.h <- 1.2
    c.l <- .8
    beta <- .5
    
    U.2.h <- (1 - beta)^(1/(1-rho)) * c.h
    U.2.l <- (1 - beta)^(1/(1-rho)) * c.l
    
    U.1.C.hh <- ((1-beta)*c.h^(1-rho) + beta * U.2.h^(1-rho))^(1/(1-rho))
    U.1.C.hl <- ((1-beta)*c.h^(1-rho) + beta * U.2.l^(1-rho))^(1/(1-rho))
    U.1.C.lh <- ((1-beta)*c.l^(1-rho) + beta * U.2.h^(1-rho))^(1/(1-rho))
    U.1.C.ll <- ((1-beta)*c.l^(1-rho) + beta * U.2.l^(1-rho))^(1/(1-rho))
    
    U.0.C <- (1 - beta)^(1/(1-rho)) * (.25 * U.1.C.hh^(1-gamma) +.25 * U.1.C.hl^(1-gamma) +
                                         .25 * U.1.C.lh^(1-gamma) +.25 * U.1.C.ll^(1-gamma)
    )^(1/(1-gamma))
    
    E.U2gamma <- .5 * U.2.h^(1 - gamma) + .5 * U.2.l^(1 - gamma)
    
    U.1.D.h <- ((1-beta)*c.h^(1-rho) + beta * E.U2gamma^((1-rho)/(1-gamma)))^(1/(1-rho))
    U.1.D.l <- ((1-beta)*c.l^(1-rho) + beta * E.U2gamma^((1-rho)/(1-gamma)))^(1/(1-rho))
    
    U.0.D <- (1 - beta)^(1/(1-rho)) * (.5 * U.1.D.h^(1-gamma) +.5 * U.1.D.l^(1-gamma))^(1/(1-gamma))
    
    all.C.minus.D.values <- c(all.C.minus.D.values,U.0.C-U.0.D)
  }
  
  # Plot each panel for one value of rho.
  if(rho==.5){
    plot(all.gamma.values,all.C.minus.D.values,type="l",lwd=2,las=0,
         xlab=expression(paste("RRA coefficient (",gamma,")",sep="")),ylab="",
         main=expression(paste("(a) ",rho,"=0.50",sep="")),cex.lab=1.1
    )
  }
  if(rho==.9){
    plot(all.gamma.values,all.C.minus.D.values,type="l",lwd=2,las=0,
         xlab=expression(paste("RRA coefficient (",gamma,")",sep="")),ylab="",
         main=expression(paste("(b) ",rho,"=0.90",sep="")),cex.lab=1.1
    )
  }
  if(rho==1.5){
    plot(all.gamma.values,all.C.minus.D.values,type="l",lwd=2,las=0,
         xlab=expression(paste("RRA coefficient (",gamma,")",sep="")),ylab="",
         main=expression(paste("(c) ",rho,"=1.50",sep="")),cex.lab=1.1
    )
  }
  abline(v=rho,col="darkgrey")
  abline(h=0,col="black",lty=2)
}

```

## Epstein-Zin pricing implications vary with preference parameters

This example shows how Epstein-Zin pricing implications vary with preference parameters.

```{r fig_EZresolution2, fig.width=7, fig.height=3.5}
# Purpose: This example shows how Epstein-Zin pricing implications vary with preference
# parameters

beta <- .9
gamma <- seq(0.5001,10,by=.1)

par(mfrow=c(1,2))
par(plt=c(.15,.9,.25,.85))

omega <- .05
E.m.0 <- - log(0.5 * exp(beta*(1-gamma)*omega) + 0.5 * exp(-beta*(1-gamma)*omega))
P <- .5 * exp(E.m.0 - beta*(1-gamma)*omega)
P.check <- .5 * exp(E.m.0 + beta*(1-gamma)*omega) +
  .5 * exp(E.m.0 - beta*(1-gamma)*omega)
# Should be equal to 1.
plot(gamma,P,type="l",lwd=2,cex.main=1.1,las=1,
     xlab=expression(paste("RRA coefficient (",gamma,")",sep="")),ylab="",
     main=expression(paste("(a) ",omega," = 0.05",sep="")))
abline(v=1,col="darkgrey")
abline(h=0.5,col="black",lty=2)

omega <- .2
E.m.0 <- - log(0.5 * exp(beta*(1-gamma)*omega) + 0.5 * exp(-beta*(1-gamma)*omega))
P <- .5 * exp(E.m.0 - beta*(1-gamma)*omega)
P.check <- .5 * exp(E.m.0 + beta*(1-gamma)*omega) +
  .5 * exp(E.m.0 - beta*(1-gamma)*omega)
# Should be equal to 1.
plot(gamma,P,type="l",lwd=2,cex.main=1.1,las=1,
     xlab=expression(paste("RRA coefficient (",gamma,")",sep="")),ylab="",
     main=expression(paste("(b) ",omega," = 0.2",sep="")))
abline(v=1,col="darkgrey")
abline(h=0.5,col="black",lty=2)

```


<!-- ```{r fig_SPF10yrStock, fig.width=7, fig.height=4} -->
<!-- library(DTAM) -->
<!-- library(lubridate) -->

<!-- # SPF data: -->
<!-- DATA_SPF <- data.frame(Date=SPF$date,RGDP10=SPF$GDP10) -->
<!-- DATA_SPF$Date <- DATA_SPF$Date %m+% months(2) # because convention that quarterly dates use first month of quarter -->
<!-- # Shiller data (stock prices): -->
<!-- DATA_Shiller <- data.frame(Date=Shiller$Date,Ps=Shiller$RealTotReturnPrice) -->
<!-- # GDP data: -->
<!-- DATA_GDP <- data.frame(Date=Data_Macro_US_quarterly$date, -->
<!--                        RGDP=Data_Macro_US_quarterly$GDPC1) -->

<!-- DATA <- merge(DATA_SPF,DATA_Shiller,by="Date",all=TRUE) -->
<!-- DATA <- merge(DATA_GDP,DATA,by="Date",all=TRUE) -->

<!-- DATA <- DATA[complete.cases(DATA),] -->
<!-- T <- dim(DATA)[1] -->
<!-- DATA$chgestock <- NaN -->
<!-- DATA$chgestock[2:T] <- log(DATA$Ps[2:T]/DATA$Ps[1:(T-1)]) -->
<!-- DATA$chgeGDP10 <- NaN -->
<!-- DATA$chgeGDP10[2:T] <- DATA$RGDP10[2:T] - DATA$RGDP10[1:(T-1)] -->
<!-- DATA$chgeGDP <- NaN -->
<!-- DATA$chgeGDP[2:T] <- log(DATA$RGDP[2:T]/DATA$RGDP[1:(T-1)]) -->

<!-- #X <- cbind(DATA$chgestock,DATA$chgeGDP) -->
<!-- X <- cbind(DATA$chgestock) -->
<!-- Y <- DATA$chgeGDP10 -->
<!-- XY <- cbind(Y,X,as.numeric(format(DATA$Date,"%Y"))) -->
<!-- XY <- XY[complete.cases(XY),] -->
<!-- X <- XY[,2:(1+dim(X)[2])] -->
<!-- Y <- XY[,1] -->
<!-- model <- lm(Y ~ X) -->
<!-- coefficients <- coefficients(model) -->
<!-- x_interval <- seq(min(X), max(X), length.out = 100) -->
<!-- y_lower <- coefficients[1] + coefficients[2] * x_interval - (1.96 * qt(0.975, df = df.residual(model))) * sqrt(coefficients[2]^2 + (x_interval - coefficients[1])^2) -->
<!-- y_upper <- coefficients[1] + coefficients[2] * x_interval + (1.96 * qt(0.975, df = df.residual(model))) * sqrt(coefficients[2]^2 + (x_interval - coefficients[1])^2) -->
<!-- par(plt=c(.15,.95,.2,.95)) -->
<!-- plot(X, Y, xlab = "stock return", ylab = "change in 10-yr GDP expectation", main = "",ylim=c(-.5,.5),pch=19,las=1) -->
<!-- abline(lm(Y~X), col = "darkgrey") -->
<!-- grid() -->
<!-- text(X+.05,Y,XY[,3],col="light grey") -->
<!-- #polygon(c(x_interval, rev(x_interval)), c(y_lower, rev(y_upper)), col = "lightgrey", border = NA) -->

<!-- eq <- lm(chgestock~chgeGDP+chgeGDP10,data=DATA) -->
<!-- ``` -->


The following code produces a figure that shows the maximum Sharpe ratios that can be achieved with affine Epstein-Zin models.


## The Bansal-Yaron long-run-risk environment and its state variables

This example introduces the Bansal-Yaron long-run-risk environment and its state variables.

```{r fig_BY1, fig.width=7, fig.height=4}
# Purpose: This example introduces the Bansal-Yaron long-run-risk environment and its state
# variables
library(DTAM)

# ------------------------------------------------------------------------------
# Bansal and Yaron (2004): Homoskedastic case
# (using psi_GaussianVAR() for Laplace transform)
# ------------------------------------------------------------------------------

# Parameterization from Bansal and Yaron (2004):
EIS   <- 1.5
gamma <- 10
delta <- .998
mu    <- .0015
mu_d  <- .0015
phi   <- .979
beta  <- 3
sigma <- 0.0078
varphi_e <- .044
varphi_d <- 4.5

model <- list(
  mu = matrix(c(0,mu,mu_d),3,1),
  Phi = matrix(c(phi,1,beta,rep(0,6)),3,3),
  Sigma = diag(sigma^2*c(varphi_e,1,varphi_d)^2),
  rho = 1/EIS,
  phi = phi,
  gamma = gamma,
  delta = delta,
  sigma = sigma,
  mu_c0 = 0,
  mu_c1 = c(0,1,0),
  n_w = 3
)

model_solved <- solve_EZ_SDF(model,psi_GaussianVAR,z_bar_ini=5,
                             nb_loop_z_bar = 20,nb_loop_mu_z1 = 20)
# Check formula (compare with analytical BY formula):
A1 <- (1 - 1/EIS)/(1 - model_solved$kappa1*model$phi)

all_EIS <- c(.5,.9999,1.5,NaN)
all_col <- c("black","black","black","darkgrey")
all_lty <- c(1,2,3,1)
all_gamma <- c(seq(.50001,3,length.out=10),4,6,8,10)

all_maxSR <- matrix(0,length(all_EIS),length(all_gamma))

count_EIS <- 0
for(EIS in all_EIS){
  count_EIS <- count_EIS + 1
  count_gamma <- 0
  model$rho <- 1/EIS
  for(gamma in all_gamma){
    count_gamma <- count_gamma + 1
    model$gamma <- gamma
    if(is.na(EIS)){
      model$rho <- model$gamma
    }
    res <- solve_EZ_SDF(model,psi_GaussianVAR,z_bar_ini=5,
                        nb_loop_z_bar = 20,nb_loop_mu_z1 = 20)
    psi_w_2alpha <- psi_GaussianVAR(2*res$alpha,model)
    psi_w_alpha  <- psi_GaussianVAR(res$alpha,model)
    w <- res$Ew
    maxSR <- sqrt(exp(t(psi_w_2alpha$a - 2*psi_w_alpha$a)%*%w+
                        psi_w_2alpha$b - 2*psi_w_alpha$b)-1)
    all_maxSR[count_EIS,count_gamma] <- maxSR*sqrt(12)
  }
}

par(plt=c(.15,.95,.2,.99))
plot(all_gamma,(all_maxSR[1,]),ylim=c(0,.75),type="l",
     lwd=2,las=1,col=all_col[1],lty=all_lty[1],
     xlab=expression(paste("coefficient of RRA (",gamma,")",sep="")),
     ylab="maximum Sharpe ratio (annualized)")
grid()
for(i in 2:length(all_EIS)){
  lines(all_gamma,(all_maxSR[i,]),lwd=2,
        col=all_col[i],lty=all_lty[i])
}
legend("topleft",
       c(expression(paste("EIS = 0.5",sep="")),
         expression(paste("EIS = 1",sep="")),
         expression(paste("EIS = 1.5",sep="")),
         expression(paste("EIS = 1/",gamma," (CCAPM case)",sep=""))),
       lwd=2,
       lty=all_lty,
       col=all_col,
       seg.len = 3
)

# for(EIS in all_EIS){
#   abline(v=1/EIS,lty=3)
# }

```


Bansal and Yaron (2004) as a special case of Bansal and Shaliastovich (2012):

## Yields, returns, and premia in the Bansal-Yaron model

This example studies yields, returns, and premia in the Bansal-Yaron model.

```{r fig_BY2, fig.width=7, fig.height=7}
# Purpose: This example studies yields, returns, and premia in the Bansal-Yaron model
library(DTAM)

# Parameterization from Bansal and Yaron (2004):
EIS   <- 1.5
gamma <- 10
delta <- .998
mu    <- .0015
mu_d  <- .0015
phi   <- .979
beta  <- 3
sigma <- 0.0078
varphi_e <- .044
varphi_d <- 4.5
nu1      <- .987
sigma_w  <- .23*10^(-5)
# sigma_w  <- 0

freq <- 12 # monthly frequency

# Definition of theta, using Bansal and Yaron notations:
theta <- (1 - gamma)/(1 - 1/EIS)

Phi <- matrix(0,3,3)
Phi[1,1] <- phi
Phi[2,1] <- 1
Phi[3,1] <- beta

model <- list(
  mu_c0   = mu,
  mu_c1   = matrix(c(0,1,0, # x weightings (x,x_1,eta,u)
                     0), # z weightings (sigma_t)
                   ncol=1),
  mu_d0   = mu,
  mu_d1   = matrix(c(0,0,1, # x weightings (x,x_1,eta,u)
                     0), # z weightings (sigma_t)
                   ncol=1),
  Pi      = Phi,
  # unconditional means of x's conditional variances:
  chi.0   = c(varphi_e^2*sigma^2,sigma^2,varphi_d^2*sigma^2), 
  # loadings of x's conditional variances on z:
  chi.1   = matrix(c(varphi_e^2,1,varphi_d^2),nrow=1),
  nu      = as.matrix(nu1),
  sigma.w = c(sigma_w),
  gamma   = gamma,
  rho     = 1/EIS,
  delta   = delta,
  n_w = 3+1)


model_solved <- solve_EZ_SDF(model,psi_BansalShaliastovich,z_bar_ini=5,
                             nb_loop_z_bar = 10,nb_loop_mu_z1 = 400)
# Check with formulas reported in Bansal and Yaron:
# A1 <- (1 - 1/EIS)/(1 - model_solved$kappa1*phi)
# A2 <- .5*((theta - theta/EIS)^2 + (theta*model_solved$mu_z1[1]*model_solved$kappa1*varphi_e)^2)/
#   (theta*(1 - model_solved$kappa1*nu1))
# model_solved$mu_z1
# c(A1,A2)

all_EIS <- c(.5,.9999,1.5,NaN)
all_col <- c("black","black","black","darkgrey")
all_lty <- c(1,2,3,1)
all_gamma <- c(seq(.50001,4,length.out=20),6,8,10)
# all_gamma <- seq(.80001,4,length.out=40)

all_maxSR      <- matrix(0,length(all_EIS),length(all_gamma))
all_Erf        <- matrix(0,length(all_EIS),length(all_gamma))
all_Ers        <- matrix(0,length(all_EIS),length(all_gamma))
all_std_dev_pd <- matrix(0,length(all_EIS),length(all_gamma))

count_EIS <- 0
for(EIS in all_EIS){
  count_EIS <- count_EIS + 1
  count_gamma <- 0
  model$rho <- 1/EIS
  for(gamma in all_gamma){
    count_gamma <- count_gamma + 1
    model$gamma <- gamma
    if(is.na(EIS)){
      model$rho <- model$gamma
    }
    model_solved <- solve_EZ_SDF(model,psi_BansalShaliastovich,
                                 z_bar_ini=5,
                                 nb_loop_z_bar = 10,nb_loop_mu_z1 = 200)
    
    # Compute maximum Sharpe ratio:
    psi_w_2alpha <- psi_BansalShaliastovich(2*model_solved$alpha,model_solved)
    psi_w_alpha  <- psi_BansalShaliastovich(model_solved$alpha,model_solved)
    Ew <- model_solved$Ew
    maxSR <- sqrt(exp(t(psi_w_2alpha$a - 2*psi_w_alpha$a)%*%Ew+
                        psi_w_2alpha$b - 2*psi_w_alpha$b)-1)
    all_maxSR[count_EIS,count_gamma] <- maxSR*sqrt(freq)
    
    # Compute average risk-free rate:
    Erf <- model_solved$eta0 + t(model_solved$eta1)%*%Ew
    all_Erf[count_EIS,count_gamma] <- freq*Erf
    
    # Solve for stock prices:
    model_solved <- solve_EZ_stock_return(model_solved,psi_BansalShaliastovich,
                                          nb_loop_z_bar = 10,nb_loop_mu_z1 = 200)
    # Compute expected excess return:
    psi_mu_rs2 <- psi_BansalShaliastovich(model_solved$mu_rs2,
                                          model_solved)
    Ers <- model_solved$mu_rs0 + t(model_solved$mu_rs1)%*%Ew + 
      psi_mu_rs2$b + t(psi_mu_rs2$a) %*%Ew
    all_Ers[count_EIS,count_gamma] <- freq*Ers
    
    # Compute variance of p-d:
    Var_pd <- t(model_solved$mu_s1) %*% model_solved$Vw %*% model_solved$mu_s1
    all_std_dev_pd[count_EIS,count_gamma] <- sqrt(Var_pd)
  }
}

par(mfrow=c(2,2))
par(plt=c(.25,.98,.25,.85))

# Average stock excess return:
all_xs <- all_Ers - all_Erf
plot(all_gamma,100*all_xs[1,],ylim=c(0,7),type="l",
     lwd=2,las=1,col=all_col[1],lty=all_lty[1],
     xlab=expression(paste("coefficient of RRA (",gamma,")",sep="")),
     ylab="in percent, annualized",
     main=expression(paste("(a) Expected excess return",sep="")),cex.main=1.1)
grid()
for(i in 2:length(all_EIS)){
  lines(all_gamma,100*all_xs[i,],lwd=2,
        col=all_col[i],lty=all_lty[i])
}
# Plot points where the curve 1/EIS crosses the other ones:
interpolation <- spline(all_gamma, all_xs[4,], xout = 1/all_EIS[1:3])
points(1/all_EIS[1:3],interpolation$y,pch=19)

legend("topleft",
       c(expression(paste("EIS = 0.5",sep="")),
         expression(paste("EIS = 1",sep="")),
         expression(paste("EIS = 1.5",sep="")),
         expression(paste("EIS = 1/",gamma," (CCAPM case)",sep=""))),
       lwd=2,
       lty=all_lty,
       col=all_col,
       seg.len = 3
)

# Maximum Sharpe ratios:
plot(all_gamma,(all_maxSR[1,]),ylim=c(0,.75),type="l",
     lwd=2,las=1,col=all_col[1],lty=all_lty[1],
     xlab=expression(paste("coefficient of RRA (",gamma,")",sep="")),
     ylab="annualized",
     main=expression(paste("(b) Maximum Sharpe ratio",sep="")),cex.main=1.1)
grid()
for(i in 2:length(all_EIS)){
  lines(all_gamma,(all_maxSR[i,]),lwd=2,
        col=all_col[i],lty=all_lty[i])
}
# Plot points where the curve 1/EIS crosses the other ones:
interpolation <- spline(all_gamma, all_maxSR[4,], xout = 1/all_EIS[1:3])
points(1/all_EIS[1:3],interpolation$y,pch=19)

# Risk-free rate:
plot(all_gamma,(all_Erf[1,]),ylim=c(0,.2),type="l",
     lwd=2,las=1,col=all_col[1],lty=all_lty[1],
     xlab=expression(paste("coefficient of RRA (",gamma,")",sep="")),
     ylab="in percent, annualized",
     main=expression(paste("(c) Risk-free rate",sep="")),cex.main=1.1)
grid()
for(i in 2:length(all_EIS)){
  lines(all_gamma,(all_Erf[i,]),lwd=2,
        col=all_col[i],lty=all_lty[i])
}
# Plot points where the curve 1/EIS crosses the other ones:
interpolation <- spline(all_gamma, all_Erf[4,], xout = 1/all_EIS[1:3])
points(1/all_EIS[1:3],interpolation$y,pch=19)

# Volatility of p-d:
plot(all_gamma,(all_std_dev_pd[1,]),ylim=c(0,.3),type="l",
     lwd=2,las=1,col=all_col[1],lty=all_lty[1],
     xlab=expression(paste("coefficient of RRA (",gamma,")",sep="")),
     ylab="",
     main=expression(paste("(d) Volatility of ",log(P[s*','*t]/D[t]),sep="")),cex.main=1.1)
grid()
for(i in 2:length(all_EIS)){
  lines(all_gamma,(all_std_dev_pd[i,]),lwd=2,
        col=all_col[i],lty=all_lty[i])
}
# Plot points where the curve 1/EIS crosses the other ones:
interpolation <- spline(all_gamma, all_std_dev_pd[4,], xout = 1/all_EIS[1:3])
points(1/all_EIS[1:3],interpolation$y,pch=19)


# Computation of mean and variance of P/D:
T <- 500
psi_PD_inf <- reverse_MHLT(psi_BansalShaliastovich,
                           u1 = model_solved$mu_s1,
                           u2 = 0*model_solved$mu_s1,
                           H  = T, model_solved)
E_PD <- exp(model_solved$mu_s0 + psi_PD_inf$B[T])

psi_PD2_inf <- reverse_MHLT(psi_BansalShaliastovich,
                            u1 = 2*model_solved$mu_s1,
                            u2 = 0*model_solved$mu_s1,
                            H  = T, model_solved)
Var_PD <- exp(2*model_solved$mu_s0 + psi_PD2_inf$B[T]) -
  exp(2*model_solved$mu_s0 + 2*psi_PD_inf$B[T])
Std_dev_PD <- sqrt(Var_PD)
# Compare with Gaussian approximation:
Var_pd <- t(model_solved$mu_s1) %*% model_solved$Vw %*% model_solved$mu_s1
E_PD_approx <- exp(model_solved$mu_s0 + Var_pd/2)
```



## The Bansal-Shaliastovich specification for bond and currency risk premia

This example illustrates the Bansal-Shaliastovich specification for bond and currency risk premia.

```{r fig_BS1, fig.width=7, fig.height=4.5}
# Purpose: This example illustrates the Bansal-Shaliastovich specification for bond and currency
# risk premia
library(DTAM)

max_maturity_in_years <- 10

# Parameterization of Bansal & Shaliastovich (2012):
rho.c    <- 0.810
rho.cpi  <- -0.047
rho.pi   <- .988
mu.c     <- .0049
mu.pi    <- .009
sigma.c  <- 4.6e-03
sigma.pi <- 5.5e-03
delta    <- .994
gamma    <- 20.9
psi      <- 1.81
nu.c     <- 0.994
nu.pi    <- 0.979
sigma.xc  <- 1.09e-03
sigma.xpi <- 1.11e-03
sigma.wc  <- 1.85e-07
sigma.wpi <- 1.81e-07

Phi <- matrix(0,4,4)
Phi[1:2,1:2] <- c(rho.c,0,rho.cpi,rho.pi)

model <- list(
  mu_c0   = mu.c,
  mu_c1   = matrix(c(1,0,0,0, # x weightings
                     0,0), # z weightings
                   ncol=1),
  Pi      = Phi,
  # uncondit. means of x's conditional variances:
  chi.0   = c(sigma.xc^2,sigma.xpi^2,sigma.c^2,sigma.pi^2), 
  # loadings of x's conditional variances on z:
  chi.1   = cbind(diag(2),0*diag(2)),
  nu      = diag(c(nu.c,nu.pi)),
  sigma.w = c(sigma.wc,sigma.wpi),
  gamma   = gamma,
  rho     = 1/psi,
  delta   = delta,
  n_w = 2+2+2)

freq <- 4 # quarterly frequency

model_solved <- solve_EZ_SDF(model,psi_BansalShaliastovich,z_bar_ini=5,
                             nb_loop_z_bar = 10,nb_loop_mu_z1 = 500)

# Q log Laplace transform:
psiQ <- function(u,model_solved){
  psi1 <- psi_BansalShaliastovich(u + model_solved$alpha, model_solved)
  psi2 <- psi_BansalShaliastovich(    model_solved$alpha, model_solved)
  return(list(a = psi1$a - psi2$a,
              b = psi1$b - psi2$b))
}

# Specification of real yields:
psi <- reverse_MHLT(psiQ,
                    u1 = 0*model_solved$eta1, u2 = - model_solved$eta1,
                    H=max_maturity_in_years*freq, model_solved)
Br <- - (1:(max_maturity_in_years*freq))*c(model_solved$eta0) + c(psi$B)
Ar <- - model_solved$eta1 %*% matrix(1,1,max_maturity_in_years*freq) + 
  matrix(psi$A,model_solved$n_w,
         max_maturity_in_years*freq)
br <- - Br/(1:(max_maturity_in_years*freq))
ar <- - Ar/t(matrix(1:(max_maturity_in_years*freq),max_maturity_in_years*freq,
                    model_solved$n_w))

# Specification of nominal yields:
mu_pi1   = matrix(c(0,1,0,0, # x weightings
                    0,0), # z weightings
                  ncol=1)
psi <- reverse_MHLT(psiQ,
                    u1 = - mu_pi1, u2 = - model_solved$eta1 - mu_pi1,
                    H=max_maturity_in_years*freq, model_solved)
Bn <- - (1:(max_maturity_in_years*freq))*(mu.pi + c(model_solved$eta0)) + c(psi$B)
An <- - model_solved$eta1 %*% matrix(1,1,max_maturity_in_years*freq) + 
  matrix(psi$A,model_solved$n_w,
         max_maturity_in_years*freq)
bn <- - Bn/(1:(max_maturity_in_years*freq))
an <- - An/t(matrix(1:(max_maturity_in_years*freq),max_maturity_in_years*freq,
                    model_solved$n_w))

avg_real_yields <- freq*(br + t(ar) %*% model_solved$Ew)
avg_nomi_yields <- freq*(bn + t(an) %*% model_solved$Ew)

# Compute standard deviations of yields:
loadings_r <- freq * t(ar)
loadings_n <- freq * t(an)
vec_1 <- matrix(1,1,model_solved$n_w)
aux_r <- (loadings_r %x% vec_1)*(vec_1 %x% loadings_r) * 
  (matrix(1,max_maturity_in_years*freq,1) %*% matrix(model_solved$Vw,nrow=1))
aux_n <- (loadings_n %x% vec_1)*(vec_1 %x% loadings_n) * 
  (matrix(1,max_maturity_in_years*freq,1) %*% matrix(model_solved$Vw,nrow=1))
std_dev_r <- sqrt(aux_r %*% matrix(1,model_solved$n_w^2,1))
std_dev_n <- sqrt(aux_n %*% matrix(1,model_solved$n_w^2,1))

# Consider alternative values of covariance between consumption and inflation
#  and recompute nominal prices.
other_avg_nomi_yields <- NULL
all_rho.cpi  <- c(0,0.047)
for(rho.cpi in all_rho.cpi){
  model$Pi[1,2] <- rho.cpi
  
  model_solved <- solve_EZ_SDF(model,psi_BansalShaliastovich,z_bar_ini=5,
                               nb_loop_z_bar = 10,nb_loop_mu_z1 = 500)
  psi <- reverse_MHLT(psiQ,
                      u1 = - mu_pi1, u2 = - model_solved$eta1 - mu_pi1,
                      H=max_maturity_in_years*freq, model_solved)
  Bn <- - (1:(max_maturity_in_years*freq))*(mu.pi + c(model_solved$eta0)) + c(psi$B)
  An <- - model_solved$eta1 %*% matrix(1,1,max_maturity_in_years*freq) + 
    matrix(psi$A,model_solved$n_w,
           max_maturity_in_years*freq)
  bn <- - Bn/(1:(max_maturity_in_years*freq))
  an <- - An/t(matrix(1:(max_maturity_in_years*freq),max_maturity_in_years*freq,
                      model_solved$n_w))
  other_avg_nomi_yields <- cbind(other_avg_nomi_yields,
                                 freq*(bn + t(an) %*% model_solved$Ew))
}


# Prepare plot

par(plt=c(.12,1,.2,.95))
x_axis <- 1:(max_maturity_in_years*freq)/freq
plot(x_axis,100*avg_real_yields,lwd=2,ylim=c(0,14),type="l",
     col="white",
     las=1,xlab="maturity, in years",ylab="yield, in percent")

lower_bound <- 100*(avg_real_yields - std_dev_r)
upper_bound <- 100*(avg_real_yields + std_dev_r)
polygon(c(x_axis,rev(x_axis)),
        c(lower_bound,rev(upper_bound)),col="#66666633",border=NaN)
lower_bound <- 100*(avg_nomi_yields - std_dev_n)
upper_bound <- 100*(avg_nomi_yields + std_dev_n)
polygon(c(x_axis,rev(x_axis)),
        c(lower_bound,rev(upper_bound)),col="#66666633",border=NaN)
grid()
lines(x_axis,100*avg_real_yields,lwd=2,lty=1)
lines(x_axis,100*avg_nomi_yields,lwd=2,lty=2)

for(i in 1:length(all_rho.cpi)){
  lines(x_axis,100*other_avg_nomi_yields[,i],lwd=3,lty=i+1,
        col="darkgrey")
}

legend("topleft",
       c("Real yields",
         expression(paste("Nominal yields, ",rho[c*','*pi]," = -0.047",sep="")),
         expression(paste("Nominal yields, ",rho[c*','*pi]," = 0",sep="")),
         expression(paste("Nominal yields, ",rho[c*','*pi]," = +0.047",sep=""))),
       lwd=c(2,2,3,3), # line width
       lty=c(1,2,2,3),
       col=c("black","black","darkgrey","darkgrey"),
       seg.len = 3
)

```
