# Preferred-habitat model {#PreferredHabitat}


## The preferred-habitat specification and the maturity-demand function

This example introduces the preferred-habitat specification and the maturity-demand function.

```{r fig_VV_specif_Alpha, fig.width=8, fig.height=3.5}
# Purpose: This example introduces the preferred-habitat specification and the maturity-demand
# function
library(DTAM)

n <- 2 # number of factors
H <- 30 # maximum maturity
# Calibration: (see Table 1 in Vayanos-Vila):
kappa_r <- .125
sigma_r <- .0146
kappa_beta <- .053
a.theta <- 3155
a.alpha <- 35.3
delta_alpha <- .297
delta_theta <- .307
alpha    <- 5.21
r_bar    <- 4.8
a.theta0 <- 289

# Specif of short-rate:
a1 <- matrix(c(1,rep(0,n-1)),ncol=1) # The first factor is r_t
b1 <- r_bar/100 # mean of the short-term rate

# Specify factors' dynamics:
mu    <- matrix(0,n,1)
Phi   <- matrix(0,n,n)
Phi[1:2,1:2] <- diag(c(1-kappa_r,1-kappa_beta))
Sigma        <- sigma_r^2*diag(n) # since sigma_beta = sigma_r

# Specify supply of bonds:
a      <- a.alpha/alpha
theta  <- a.theta/a
theta0 <- a.theta0/a

Beta      <- matrix(theta0 * (exp(-delta_alpha*(1:H)) -
                                exp(-delta_theta*(1:H))),H,1)
Alpha      <- matrix(0,H,n)
Alpha[,2]  <- theta * (exp(-delta_alpha*(1:H)) -
                         exp(-delta_theta*(1:H)))
xi <- alpha * matrix(1:H * exp(-delta_alpha*(1:H)),H,1)
model_VV <- list(gamma = a,mu = mu,Phi = Phi,Sigma = Sigma,
                 alpha = Alpha,beta = Beta,xi = xi,a1 = a1,b1 = b1)

par(mfrow=c(1,2))
par(plt=c(.2,.98,.25,.85))
plot(2:H,
     las=1,
     Alpha[2:H,2],type="l",lwd=2,
     xlab=expression(paste("maturity (",h,")",sep="")),ylab=expression(alpha[h*','*2]),
     main=expression(paste("(a) w/o first maturity (",alpha[1*','*2],")",sep=""))
)
grid()
abline(h=0,col="black",lty=3)

# Solve the model to get the right specification of alpha_1 and beta_1:
RES <- solve_PH_TSM(model_VV)
Alpha <- RES$alpha
Beta  <- RES$beta
plot(1:H,
     las=1,
     Alpha[1:H,2],type="l",lwd=2,
     xlab=expression(paste("maturity (",h,")",sep="")),
     ylab=expression(alpha[h*','*2]),
     main=expression(paste("(b) w/ first maturity (",alpha[1*','*2],")",sep=""))
)
grid()
abline(h=0,col="black",lty=3)
```

## Factor loadings in the preferred-habitat model

This example shows the factor loadings generated by the preferred-habitat term-structure model.

```{r fig_VV_factor_loadings, fig.width=8, fig.height=3.5}
# Purpose: This example shows the factor loadings generated by the preferred-habitat
# term-structure model
# Single yield curve:
RES <- solve_PH_TSM(model_VV)
par(mfrow=c(1,n))
par(plt=c(.2,.98,.25,.85))
for(i in 1:n){
  if(i==1){
    main = expression(paste("(a) Loading on short-term rate",sep=""))
    ylab = expression(-a[h*','*1]/h)
  }else{
    main = expression(paste("(b) Loading on supply factor",sep=""))
    ylab = expression(-a[h*','*2]/h)
  }
  plot(RES$a[,i],type="l",
       las=1,lwd=2,
       xlab=expression(paste("maturity (",h,")",sep="")),
       ylab=ylab,
       main=main)
  grid()
}
```

## How bond supply shocks affect the yield curve under preferred habitat

This example studies how bond supply shocks affect the yield curve under preferred habitat.

```{r fig_VV_effect_supply, fig.width=8, fig.height=6}
# Purpose: This example studies how bond supply shocks affect the yield curve under preferred
# habitat
# Shock:
shock <- matrix(0,n,1)
shock[2] <- sigma_r # shock on the net supply factor
# baseline yield curve:
avg.y <- RES$b # average yield curve
# yield curve after shock:
shock.y <- RES$b + RES$a %*% shock
# Quantities:
avg.z   <- Beta - xi * avg.y
shock.z <- Beta + Alpha %*% shock - xi * shock.y

legendtext <- c("w/o supply shock",
                "w/ supply shock")

par(mfrow=c(2,1))
par(plt=c(.15,.95,.2,.8))

plot(100*avg.y,type="l",
     ylim=c(4,8),
     lwd=2,las=1,
     xlab=expression(paste("maturity (",h,")",sep="")),
     ylab=expression(paste("yield, in percent",sep="")),
     main=expression(paste("(a) Yield curve impact",sep="")))
grid()
lines(100*shock.y,lty=2,lwd=2,col="dark grey")
legend("bottomright",
       legendtext,
       lty=c(1,2), # gives the legend appropriate symbols (lines)
       lwd=c(2,2), # line width
       col= c("black","dark grey"),
       # pch=c(19,15,17),
       # pt.cex = 1,
       # pt.bg = c("black","dark grey"),
       seg.len = 2
)

# plot(avg.z,ylim=c(min(avg.z,shock.z),1.1*max(avg.z,shock.z)),pch=19,
#      xlab="maturity",ylab="Fraction of portfolio",las=1)
# grid()
# abline(h=0,col="black",lty=3)
# points(shock.z,lty=2,col="red",lwd=2)

# Create the barplot
barplot(100*rbind(c(avg.z), c(shock.z)), beside = TRUE,
        names.arg = 1:H, col = c("black", "dark grey"), las = 1,
        legend.text = legendtext, args.legend = list(x = "bottomright",bg="white"),
        main=expression(paste("(b) Holdings impact",sep="")),
        xlab=expression(paste("maturity (",h,")",sep="")),
        ylab=expression(paste("Fraction, in percent",sep="")))
grid()
```

## The excess-return implications of the preferred-habitat model

This example reports the excess-return implications of the preferred-habitat model.

```{r fig_VV_Excess_returns, fig.width=8, fig.height=3.5}
# Purpose: This example reports the excess-return implications of the preferred-habitat model
a_h <- RES$A
a_h_1 <- rbind(0,a_h[1:(H-1),])
b_h <- RES$B
b_h_1 <- matrix(rbind(0,matrix(b_h[1:(H-1)])),ncol=1)

# Conditional variance of returns:
cond_var <- ((a_h_1 %x% matrix(1,1,n)) * (matrix(1,1,n) %x% a_h_1)) %*% c(Sigma)
# Unconditional mean of the conditional continuously compounded excess return:
avg_xs_returns <- b_h_1 - b_h - b1 + 0.5*cond_var
# Maturities with a non-degenerate one-period holding return:
risky_maturities <- which(cond_var > .Machine$double.eps)

# Compute unconditional standard deviation of excess return:--------------------
# Compute matrix of variance of W_t = [w_t',w_{t-1}']':
Var_w <- matrix(solve(diag(4) - Phi %x% Phi) %*% c(Sigma),2,2)
Var_w_w_1 <- cbind(rbind(Var_w,Var_w %*% t(Phi)),
                   rbind(Phi %*% Var_w, Var_w))
# Loadings of excess return on W_t:
a_h_1_a_h <- cbind(a_h_1,- a_h - matrix(1,H,1) %*% t(a1))
# Unconditional variance of excess return:
cond_var_xs <- 
  ((a_h_1_a_h %x% matrix(1,1,2*n)) * (matrix(1,1,2*n) %x% a_h_1_a_h)) %*%
  c(Var_w_w_1)
# Unconditional variance of the conditional expected excess return:
aux <- a_h_1 %*% Phi - a_h - matrix(1,H,1) %*% t(a1)
var_expected_xs <-
  ((aux %x% matrix(1,1,n)) * (matrix(1,1,n) %x% aux)) %*% c(Var_w)

# Central 90% stationary range of the conditional expected excess return:------
q05_normal <- qnorm(.05)
q95_normal <- qnorm(.95)
lower_bound_xs <- avg_xs_returns + q05_normal * sqrt(var_expected_xs)
upper_bound_xs <- avg_xs_returns + q95_normal * sqrt(var_expected_xs)

# Exact conditional Sharpe ratio of the arithmetic holding return. If
# x_{t,h}=log(E_t[G_{t+1,h}])-r_t and
# v_h=Var_t(log(G_{t+1,h})), conditional lognormality gives
# SR_{t,h}=(1-exp(-x_{t,h}))/sqrt(exp(v_h)-1). Its absolute value is the
# asset-specific object bounded by the Hansen--Jagannathan maximum Sharpe ratio.
# Because x_{t,h} is Gaussian under the stationary distribution, the
# unconditional mean and quantiles can be computed analytically.
SR <- lower_bound_SR <- upper_bound_SR <- rep(NA_real_,H)
SR_denominator <- sqrt(exp(cond_var[risky_maturities])-1)
SR[risky_maturities] <-
  (1-exp(-avg_xs_returns[risky_maturities] +
           0.5*var_expected_xs[risky_maturities]))/SR_denominator
q05_xs <- avg_xs_returns[risky_maturities] +
  q05_normal*sqrt(var_expected_xs[risky_maturities])
q95_xs <- avg_xs_returns[risky_maturities] +
  q95_normal*sqrt(var_expected_xs[risky_maturities])
lower_bound_SR[risky_maturities] <- (1-exp(-q05_xs))/SR_denominator
upper_bound_SR[risky_maturities] <- (1-exp(-q95_xs))/SR_denominator
# ------------------------------------------------------------------------------

par(mfrow=c(1,2))
par(plt=c(.23,.98,.23,.85))
plot(100*avg_xs_returns,col="white",las=1,
     xlab=expression(paste("maturity (",h,")",sep="")),
     ylab=expression(paste("Excess return, in percentage points",sep="")),
     ylim=range(100*c(lower_bound_xs,upper_bound_xs), finite=TRUE),
     main=expression(paste("(a) Expected excess return",sep="")))
grid()
polygon(c(1:H,H:1),100*c(lower_bound_xs,rev(upper_bound_xs)),
        col="#CCCCCCBB",border = NaN)
lines(100*avg_xs_returns,lwd=2)

plot(risky_maturities,100*SR[risky_maturities],col="white",las=1,
     ylim=range(100*c(lower_bound_SR,upper_bound_SR), finite=TRUE),
     xlab=expression(paste("maturity (",h,")",sep="")),
     ylab=expression(paste("Sharpe ratio, in percent",sep="")),
     main=expression(paste("(b) Conditional Sharpe ratio",sep="")))
grid()
polygon(c(risky_maturities,rev(risky_maturities)),
        100*c(lower_bound_SR[risky_maturities],rev(upper_bound_SR[risky_maturities])),
        col="#CCCCCCBB",border = NaN)
lines(risky_maturities,100*SR[risky_maturities],lwd=2)

# Compute average maximum Sharpe ratio
k <- 20
quantiles <- seq(.05,.95,length.out=20)
q <- matrix(qnorm(quantiles),nrow=1)
values.w <- rbind(q %x% matrix(1,1,k),
                  matrix(1,1,k) %x% q)
values.lambda <- RES$lambda0 %*% matrix(1,1,k^2) + RES$lambda1 %*% values.w

lambda.Sigma.lambda <- ((t(values.lambda) %x% matrix(1,1,n)) *
                          (matrix(1,1,n) %x% t(values.lambda))) %*% c(Sigma)
E_SR <- mean(sqrt(exp(lambda.Sigma.lambda)-1))
```

## The preferred-habitat setup to exchange-rate implications

This example extends the preferred-habitat setup to exchange-rate implications.

```{r fig_VV_plus_FX}
# Purpose: This example extends the preferred-habitat setup to exchange-rate implications
# Example where there is a foreign yield curve.
mu_2yc    <- matrix(0,3,1)
Phi_2yc   <- diag(c(diag(Phi),0.3))
# Phi_2yc   <- diag(c(diag(Phi),Phi[1,1]))
Sigma_2yc <- diag(c(diag(Sigma),5*Sigma[1,1]))
rho <- .5
Sigma_2yc[1,3] <- rho*Sigma[1,1]
Sigma_2yc[3,1] <- rho*Sigma[1,1]
Sigma_2yc[2,2] <- 1*Sigma[2,2]

model <- list(gamma = a,
              mu = mu_2yc,Phi = Phi_2yc,Sigma = Sigma_2yc,
              alpha = cbind(.5*Alpha,0),
              beta = .5*Beta,xi = .5*xi,
              alpha_star = cbind(0.5*Alpha,0),
              beta_star = .5*Beta,xi_star = .5*xi,
              # b1_star = b1, a1_star = matrix(c(0,0,1),n+1,1),
              mu_e0 = 0, mu_e1 = matrix(c(0,0,1),n+1,1),
              a1 = matrix(c(1,0,0),n+1,1),b1 = b1)
RES_2yc <- solve_PH_TSM(model)

# Compute beta's:
mu_e1 <- RES_2yc$mu_e1

Var_w <- matrix(solve(diag((n+1)^2) - Phi_2yc %x% Phi_2yc)%*%c(Sigma_2yc),n+1,n+1)

sum_Phi_h <- 0
all_beta <- NULL
Phi_h <- diag(n+1)
for(h in 1:H){
  a_h      <- matrix(RES_2yc$a[h,],ncol=1)
  a_h_star <- matrix(RES_2yc$a_star[h,],ncol=1)
  Phi_h <- Phi_h %*% Phi_2yc
  sum_Phi_h <- sum_Phi_h + Phi_h
  Cov <- t(mu_e1) %*% (sum_Phi_h) %*% Var_w %*% (a_h - a_h_star)
  Var_x <- t(a_h - a_h_star) %*% Var_w %*% (a_h - a_h_star)
  beta_h <- Cov/Var_x/h
  all_beta <- c(all_beta,beta_h)
}
plot(all_beta)
```

## Nominal and real term structures in the preferred-habitat framework

This example compares nominal and real term structures in the preferred-habitat framework.

```{r fig_nom_real1, fig.width=8, fig.height=3.5}
# Purpose: This example compares nominal and real term structures in the preferred-habitat
# framework

# Model parameterization:
rho_r  <- .8
rho_pi <- .8
sigma_r  <- .01 
sigma_pi <- sigma_r
a_r   <- 0.1
a_pi  <- 0.3
Phi <- matrix(c(rho_r,-a_r,a_pi,rho_pi),2,2)
Sigma_12 <- diag(c(sigma_r,sigma_pi))
Sigma <- Sigma_12 %*% t(Sigma_12)

H <- 20

# Compute unconditional variances of state variables:
Var_w <- matrix(solve(diag((2)^2) - Phi %x% Phi)%*%c(Sigma),2,2)
stdv_r  <- sqrt(Var_w[1,1])
stdv_pi <- sqrt(Var_w[2,2])

# Elasticity of net supply of Inflation-linked bonds (ILBs) to inflation:
alpha_pi <- -0.3/H/stdv_pi

# Average share of ILBs:
share_ILB <- .2

avg_r  <- .015 # average real short-term rate
avg_pi <- .02 # average inflation rate

# Check IRFs:
# shock <- c(1,0)
# x <- Sigma_12 %*% shock
# all_x <- matrix(x,ncol=1)
# for(h in 1:20){
#   x <- Phi %*% x
#   all_x <- cbind(all_x,x)
# }
# par(mfrow=c(1,2))
# plot(all_x[1,])
# plot(all_x[2,])

mu <- matrix(0,2,1)

# Create matrices defining supply functions:
Alpha      <- matrix(0,H,2)
Alpha[,2]  <- alpha_pi
Alpha_star <- matrix(0,H,2)
Beta      <- matrix(share_ILB/H,H,1)
Beta_star <- matrix((1-share_ILB)/H,H,1)
xi <- matrix(0,H,1)

model <- list(gamma = a,
              mu = mu,Phi = Phi,Sigma = Sigma,
              alpha = Alpha,
              beta = Beta,xi = xi,
              alpha_star = Alpha_star,
              beta_star = Beta_star,xi_star = xi,
              # b1_star = b1, a1_star = matrix(c(0,0,1),n+1,1),
              mu_e0 = - avg_pi, mu_e1 = matrix(c(0,-1),2,1),
              a1 = matrix(c(1,0),2,1),b1 = avg_r)
RES_2yc <- solve_PH_TSM(model) # solve model

# Plot average real and nominal yield curves:
par(mfrow=c(1,2))
par(plt=c(.2,.98,.25,.85))
plot(100*RES_2yc$b,type="l",col="black",lwd=2,las=1,
     ylim=c(1,7),
     xlab="maturity (in years)",ylab="yield (in percent)",
     main=expression("(a) Average yield curves"),cex.main=1.1)
grid()
lines(100*RES_2yc$b_star,col="black",lwd=2,lty=2)
legend("topleft",
       c("Nominal","Real"),
       lty=c(2,1), # gives the legend appropriate symbols (lines)
       lwd=c(2,2), # line width
       col= c("black"),
       seg.len = 2)

# Compute average inflation swap rates -----------------------------------------
# Specify the model for it to be used in 'reverse_MHLT':
psi.parameterization.P <- list(
  mu = mu,
  Phi = Phi,
  Sigma = Sigma)
psi.parameterization.Q <- list(
  mu = RES_2yc$muQ,
  Phi = RES_2yc$PhiQ,
  Sigma = Sigma)
# Use reverse-order multi-horizon LT to price ZC bonds:
AB.P <- reverse_MHLT(psi_GaussianVAR,u1=matrix(c(0,+1),2,1),u2=NaN,
                     H=H,psi.parameterization.P)
a.h.P <- - t(matrix(AB.P$a,2,H))
b.h.P <- - model$mu_e0 - matrix(AB.P$b,ncol=1)
AB.Q <- reverse_MHLT(psi_GaussianVAR,u1=matrix(c(0,+1),2,1),u2=NaN,
                     H=H,psi.parameterization.Q)
a.h.Q <- - t(matrix(AB.Q$a,2,H))
b.h.Q <- - model$mu_e0 - matrix(AB.Q$b,ncol=1)

plot(100*(RES_2yc$b_star-RES_2yc$b),type="l",col="black",lwd=2,las=1,
     ylim=c(1.5,4),
     xlab="maturity (in years)",ylab="inflation rate (in percent)",
     main=expression("(b) Average inflation expectations"),cex.main=1.1)
grid()
points(100*b.h.Q,col="black",lwd=2,pch=3)
lines(100*b.h.P,col="dark grey",lwd=2)
legend("topleft",
       c("BEIR","Inflation swap","Expected inflation"),
       lty=c(1,NaN,1), # gives the legend appropriate symbols (lines)
       lwd=c(2,2,2), # line width
       pch=c(NaN,3,NaN),
       col= c("black","black","dark grey"),
       seg.len = 2)

```

## How preferred habitat shapes inflation compensation and term premia

This example studies how preferred habitat shapes inflation compensation and term premia.

```{r fig_nom_real2, fig.width=8, fig.height=4}
# Purpose: This example studies how preferred habitat shapes inflation compensation and term
# premia

# Solve the model after shutting down elastic ILB supply.
model_no_elast <- list(gamma = a,
                       mu = mu,Phi = Phi,Sigma = Sigma,
                       alpha = 0*Alpha,
                       beta = Beta,xi = xi,
                       alpha_star = Alpha_star,
                       beta_star = Beta_star,xi_star = xi,
                       # b1_star = b1, a1_star = matrix(c(0,0,1),n+1,1),
                       mu_e0 = - avg_pi, mu_e1 = matrix(c(0,-1),2,1),
                       a1 = matrix(c(1,0),2,1),b1 = avg_r)
RES_2yc_no_elast <- solve_PH_TSM(model_no_elast) # solve model

psi.parameterization.Q_no_elast <- list(mu = RES_2yc_no_elast$muQ,
                                        Phi = RES_2yc_no_elast$PhiQ,
                                        Sigma = Sigma)


# Compare real-yield loadings with and without elastic ILB supply.
par(mfrow=c(1,2))
par(plt=c(.2,.98,.2,.85))
plot(RES_2yc$a[,2],ylim=c(0,1.3),type="l",col="black",lwd=2,
     main=expression("(a) Real yields"),
     xlab="maturity (in years)",
     ylab="loadings",
     cex.main=1.1,
     las=1)
grid()
lines(RES_2yc$a[,1],col="dark grey",lwd=2)
lines(RES_2yc_no_elast$a[,2],col="black",lty=2,lwd=2)
lines(RES_2yc_no_elast$a[,1],col="dark grey",lty=2,lwd=2)

legend("topright",
       title="Loadings on...",
       c(expression(paste(r[t],sep="")),
         expression(paste(pi[t],sep=""))),
       lty=c(NaN,NaN), # gives the legend appropriate symbols (lines)
       lwd=c(2,2), # line width
       pch=c(15,15),pt.cex = 1.3,
       col= c("dark grey","black"),
       seg.len = 2)

# Compare nominal-yield loadings with and without elastic ILB supply.
plot(RES_2yc$a_star[,2],ylim=c(0,1.3),type="l",col="black",lwd=2,
     main=expression("(b) Nominal yields"),
     xlab="maturity (in years)",
     ylab="loadings",
     cex.main=1.1,
     las=1)
grid()
lines(RES_2yc$a_star[,1],col="dark grey",lwd=2)
lines(RES_2yc_no_elast$a_star[,2],col="black",lty=2,lwd=2)
lines(RES_2yc_no_elast$a_star[,1],col="dark grey",lty=2,lwd=2)

legend("topright",
       title="Model with ...",
       c(expression(paste("elastic supply of ILBs (",alpha[pi]," < 0)",sep="")),
         expression(paste("inelastic supply of ILBs (",alpha[pi]," = 0)",sep=""))),
       lty=c(1,2), # gives the legend appropriate symbols (lines)
       lwd=c(2,2), # line width
       pch=c(NaN,NaN),pt.cex = 1.3,
       col= c("black","black"),
       seg.len = 2)


```


## Additional nominal-real yield-curve implications of habitat demand

This example highlights additional nominal-real yield-curve implications of habitat demand.

```{r fig_nom_real3, fig.width=8, fig.height=3.5}
# Purpose: This example highlights additional nominal-real yield-curve implications of habitat
# demand

# Compute factor loadings under physical and risk-neutral dynamics.

for(type_yield in c("real","nominal")){
  for(type_dynamics in c("P","Q","Q_no_elast")){
    if(type_dynamics=="P"){
      psi.parameterization <- psi.parameterization.P
    }else if(type_dynamics=="Q"){
      psi.parameterization <- psi.parameterization.Q
    }else{
      psi.parameterization <- psi.parameterization.Q_no_elast
    }
    
    aux.loadings <- 
      reverse_MHLT(psi_GaussianVAR,
                   u1=matrix(c(+0,ifelse(type_yield=="nominal",-1,0)),2,1),
                   u2=matrix(c(-1,ifelse(type_yield=="nominal",-1,0)),2,1),
                   H=H,
                   psi.parameterization)
    # Multiply by exp(-r_t):
    Loadings <- matrix(c(-1,0),2,H)
    Loadings[,1:H] <- Loadings[,1:H] + aux.loadings$A[,,1:H]
    
    eval(parse(text = gsub(" ","",paste("Loadings_",type_yield,"_",type_dynamics,
                                        " <- Loadings",
                                        sep=""))))
  }
}

# Convert loadings into term-premium loadings.
Loadings_real_TP    <- Loadings_real_Q -    Loadings_real_P
Loadings_nominal_TP <- Loadings_nominal_Q - Loadings_nominal_P

Loadings_real_TP_no_elast    <- Loadings_real_Q_no_elast -    Loadings_real_P
Loadings_nominal_TP_no_elast <- Loadings_nominal_Q_no_elast - Loadings_nominal_P

# Plot real and nominal term-premium loadings.
par(mfrow=c(1,2))

plot(Loadings_real_TP[1,],type="l",lwd=2,col="black",ylim=c(-.5,1),
     main=expression("(a) Real term premium"))
grid()
lines(Loadings_real_TP[2,],lwd=2,col="dark grey")
lines(Loadings_real_TP_no_elast[1,],lwd=2,col="black",lty=3)
lines(Loadings_real_TP_no_elast[2,],lwd=2,col="dark grey",lty=3)

plot(Loadings_nominal_TP[1,],type="l",lwd=2,col="black",ylim=c(-.5,1),
     main=expression("(b) Nominal term premium"))
grid()
lines(Loadings_nominal_TP[2,],lwd=2,col="dark grey")
lines(Loadings_nominal_TP_no_elast[1,],lwd=2,col="black",lty=3)
lines(Loadings_nominal_TP_no_elast[2,],lwd=2,col="dark grey",lty=3)

```

## Preferred-habitat bond demand to stock-market valuation effects

This example connects preferred-habitat bond demand to stock-market valuation effects.

```{r fig_stocks1, fig.width=8, fig.height=3.5}
# Purpose: This example connects preferred-habitat bond demand to stock-market valuation effects
# Check model with stocks

# ---- Different models: ----
# 1. Perpetuity
# 2. Dividend medium vol (Baseline)
# 3. Dividend high vol (HV)
# 4. Search for yields SFY
# 5. omega
# 6. More stock in economy

rho_k <- .5

mu_k  <- 0.02

omega <- .05
sd_k_med  <- .08
sd_k_high <- .13
alpha_r <- +1
share_stocks <- .25
share_stocks_high <- .5

all_omega <- c(0,0,0,0,omega,0)
all_stdev_k <- c(0,
                 sd_k_med,
                 sd_k_high,
                 sd_k_med,sd_k_med,sd_k_med)
all_alpha_r <- c(0,0,0,alpha_r,0,0)
all_share_stocks <- c(rep(share_stocks,5),
                      share_stocks_high)
all_mu_k  <- c(0,rep(mu_k,length(all_omega)-1))

mu_K1 <- matrix(c(0,0,1),ncol=1)

beta_K   <- 0
alpha_K  <- matrix(c(0,0,0),ncol=1)
xi_K     <- 0

# Output table (relevant results stored in "all_results"):
all_results <- matrix(0,16,length(all_omega))
rownames(all_results) <- c("$\\omega$",
                           "$\\mu_k$",
                           "$\\rho_k$",
                           "$\\sigma_k$",
                           "$\\alpha_{K,r}$",
                           "Share of stocks ($\\beta_K$)",
                           "$\\mathbb{E}(R_{K,t,t+1}-R^f_{t,t+1})$",
                           "$\\sigma(R_{K,t,t+1})$",
                           "$\\mathbb{E}(SR_{K,t})$",
                           "$\\mathbb{E}(r_{t,10})$",
                           "$\\sigma(r_{t,10})$",
                           "$\\mathbb{E}(R_{10,t,t+1}-R^f_{t,t+1})$",
                           "$\\sigma(R_{10,t,t+1})$",
                           "$\\mathbb{E}(SR_{10,t})$",
                           "$\\mathbb{C}or(R_{K,t,t+1},r_t)$",
                           "$\\mathbb{C}or(R_{10,t,t+1},r_t)$")
# "$\\mathbb{C}or(\\mathbb{E}_t[R_{K,t,t+1}],r_t)$")
all_results[1,] <- all_omega
all_results[2,] <- all_mu_k
all_results[3,] <- rho_k
all_results[4,] <- all_stdev_k
all_results[5,] <- all_alpha_r
all_results[6,] <- all_share_stocks

# Formatting of output table:
indic_hline <- c(6,9,14) # put "\hline" after that one.
multip_factors <- c(1,100,1,100,1,100,100,100,100,100,100,100,100,100,100,100)
suffix <- c("","\\%","","\\%","","\\%","\\%","\\%","\\%","\\%","\\%","\\%","\\%","\\%","\\%","\\%")
all_nb <- c(2,0,1,0,1,0,2,2,0,2,2,2,2,1,1,1)

for(i in 1:length(all_omega)){
  Alpha <- (1 - all_share_stocks[i]) * cbind(model_VV$alpha,0)
  Beta  <- (1 - all_share_stocks[i]) * model_VV$beta
  xi    <- (1 - all_share_stocks[i]) * model_VV$xi
  
  alpha_K[,1] <- all_alpha_r[i]
  
  mu    <- rbind(model_VV$mu,0)
  Phi   <- diag(c(diag(model_VV$Phi),rho_k))
  Sigma <- diag(c(diag(model_VV$Sigma),all_stdev_k[i]^2))
  Ew    <- solve(diag(3) - Phi) %*% mu
  # Unconditional variance matrix of w_t:
  Var_w <- matrix(solve(diag(9) - Phi %x% Phi) %*% c(Sigma),3,3)
  # Unconditional variance matrix of [w_{t+1}',w_t']':
  Var_w_w_1 <- cbind(
    rbind(Var_w,Var_w %*% t(Phi)),
    rbind(Phi %*% Var_w,Var_w)
  )
  
  model <- list(gamma = a,mu = mu,
                Phi = Phi,
                Sigma = Sigma,
                alpha = Alpha,
                beta = Beta,
                xi = xi,
                a1 = rbind(model_VV$a1,0),b1 = model_VV$b1,
                omega = all_omega[i], mu_K0=all_mu_k[i], mu_K1=mu_K1,
                beta_K=all_share_stocks[i], alpha_K=alpha_K, xi_K=xi_K)
  H <- dim(model$alpha)[1]
  
  RES <- solve_PH_TSM(model,max_iter = 200)
  
  # Exact stationary moments implied by the affine Gaussian approximation.
  # R^f_{t,t+1}=exp(r_t)-1 is the arithmetic risk-free return.
  mean_log_Rf <- model$b1 + c(t(model$a1) %*% Ew)
  var_log_Rf <- c(t(model$a1) %*% Var_w %*% model$a1)
  avg_gross_Rf <- exp(mean_log_Rf + 0.5*var_log_Rf)

  # The package returns both the arithmetic-return intercept C_K used by the
  # equilibrium approximation and C_K_log, the intercept of log(1+R_K).
  mean_log_RK <- c(RES$C_K_log + RES$Theta_K %*% model$mu +
                     (RES$D_K + RES$Theta_K %*% model$Phi) %*% Ew)
  loadings_RK <- matrix(c(RES$Theta_K,RES$D_K),ncol=1)
  var_log_RK <- c(t(loadings_RK) %*% Var_w_w_1 %*% loadings_RK)
  avg_gross_RK <- exp(mean_log_RK + 0.5*var_log_RK)
  sd_RK <- sqrt((exp(var_log_RK)-1)*exp(2*mean_log_RK+var_log_RK))
  all_results["$\\mathbb{E}(R_{K,t,t+1}-R^f_{t,t+1})$",i] <-
    avg_gross_RK - avg_gross_Rf
  all_results["$\\sigma(R_{K,t,t+1})$",i] <- sd_RK

  # Mean of the exact conditional Sharpe ratio of the arithmetic stock return.
  # If x_t=log(E_t[1+R_K])-r_t and v_K=Var_t(log(1+R_K)), then
  # SR_t=(1-exp(-x_t))/sqrt(exp(v_K)-1). Here x_t is Gaussian, so its
  # stationary mean can be evaluated analytically.
  v_K <- c(RES$Theta_K %*% Sigma %*% t(RES$Theta_K))
  loadings_xK <- RES$D_K + RES$Theta_K %*% model$Phi - t(model$a1)
  mean_xK <- c(RES$C_K_log + RES$Theta_K %*% model$mu + 0.5*v_K - model$b1 +
                 loadings_xK %*% Ew)
  var_xK <- c(loadings_xK %*% Var_w %*% t(loadings_xK))
  avg_SR_K <- (1-exp(-mean_xK+0.5*var_xK))/sqrt(exp(v_K)-1)
  all_results["$\\mathbb{E}(SR_{K,t})$",i] <- avg_SR_K
  
  # Average 10-year yield:
  avg_r10 <- RES$b[10]
  all_results["$\\mathbb{E}(r_{t,10})$",i] <- avg_r10
  
  # Volatility of 10-year yield:
  loadings_r10 <- matrix(RES$a[10,],ncol=1)
  sd_r10 <- sqrt(t(loadings_r10) %*% Var_w %*% loadings_r10)
  all_results["$\\sigma(r_{t,10})$",i] <- sd_r10
  
  # Bond-return coefficients and conditional variance:
  a_h <- RES$A
  a_h_1 <- rbind(0,a_h[1:(H-1),])
  b_h <- RES$B
  b_h_1 <- matrix(rbind(0,matrix(b_h[1:(H-1)])),ncol=1)
  cond_var <- ((a_h_1 %x% matrix(1,1,3)) * (matrix(1,1,3) %x% a_h_1)) %*% c(Sigma)

  # Exact unconditional arithmetic-return moments.
  loadings_log_R <- a_h_1 %*% model$Phi - a_h
  mean_log_R <- b_h_1 - b_h + a_h_1 %*% model$mu + loadings_log_R %*% Ew
  loadings_R10 <- matrix(c(a_h_1[10,],-a_h[10,]),ncol=1)
  var_log_R10 <- c(t(loadings_R10) %*% Var_w_w_1 %*% loadings_R10)
  avg_gross_R10 <- exp(mean_log_R[10] + 0.5*var_log_R10)
  sd_R10 <- sqrt((exp(var_log_R10)-1)*
                   exp(2*mean_log_R[10]+var_log_R10))
  all_results["$\\mathbb{E}(R_{10,t,t+1}-R^f_{t,t+1})$",i] <-
    avg_gross_R10 - avg_gross_Rf
  all_results["$\\sigma(R_{10,t,t+1})$",i] <- sd_R10

  # Mean of the exact conditional arithmetic-return Sharpe ratio.
  loadings_x_bond <- loadings_log_R - matrix(1,H,1) %*% t(model$a1)
  mean_x_bond <- b_h_1 - b_h + a_h_1 %*% model$mu + 0.5*cond_var - model$b1 +
    loadings_x_bond %*% Ew
  var_x_bond <-
    ((loadings_x_bond %x% matrix(1,1,3)) *
       (matrix(1,1,3) %x% loadings_x_bond)) %*% c(Var_w)
  avg_SR_bond <- rep(NA_real_,H)
  risky_maturities <- which(cond_var > .Machine$double.eps)
  avg_SR_bond[risky_maturities] <-
    (1-exp(-mean_x_bond[risky_maturities]+0.5*var_x_bond[risky_maturities]))/
    sqrt(exp(cond_var[risky_maturities])-1)
  all_results["$\\mathbb{E}(SR_{10,t})$",i] <- avg_SR_bond[10]
  
  # Correlation between R_K and r_t:
  loadings_r <- matrix(c(0,0,0,1,0,0),ncol=1)
  sigma_r <- sqrt(Var_w[1])
  Cov_log_RK_r <- c(t(loadings_RK) %*% Var_w_w_1 %*% loadings_r)
  all_results["$\\mathbb{C}or(R_{K,t,t+1},r_t)$",i] <-
    Cov_log_RK_r*avg_gross_RK/(sigma_r*sd_RK)
  
  # Correlation between R_10 and r_t:
  Cov_log_R10_r <- c(t(loadings_R10) %*% Var_w_w_1 %*% loadings_r)
  all_results["$\\mathbb{C}or(R_{10,t,t+1},r_t)$",i] <-
    Cov_log_R10_r*avg_gross_R10/(sigma_r*sd_R10)
  
  # Correlation between E_t(R_{k,t+1}) and r_t:
  # loadings_r <- matrix(c(1,0,0),ncol=1)
  # sigma_r <- sqrt(Var_w[1])
  # mu_ERK1_r <- mu_ERK1 - matrix(c(1,0,0),nrow=1)
  # stdev_ERK_r <-  sqrt(mu_ERK1_r %*% Var_w %*% t(mu_ERK1_r))
  # Cov <- mu_ERK1_r %*% Var_w %*% loadings_r
  # all_results["$\\mathbb{C}or(\\mathbb{E}_t[R_{K,t,t+1}],r_t)$",i] <- Cov/sigma_r/stdev_ERK_r
  
}


# Prepare LaTeX table: ---------------------------------------------------------
make_entry <- function(x,format.nb){
  output <- paste("$",sprintf(paste("%.",format.nb,"f",sep=""),x),"$",sep="")
  return(output)
}
format.nb <- 2

latex.table <- NULL
flag <- 0
for(i in 1:dim(all_results)[1]){
  this.line <- rownames(all_results)[i]
  for(j in 1:dim(all_results)[2]){
    this.line <- 
      paste(this.line,"&",make_entry(multip_factors[i]*all_results[i,j],all_nb[i]),suffix[i],sep="")
  }
  this.line <- 
    paste(this.line,"\\\\",sep="")
  latex.table <- rbind(latex.table,
                       this.line)
  if(i %in% indic_hline){
    flag <- flag + 1
    if(flag==1){
      latex.table <- rbind(latex.table,
                           "\\hline",
                           "{\\bf Model-implied moments:}\\\\",
                           "\\hline")
    }else{
      latex.table <- rbind(latex.table,
                           "\\hline")
    }
  }
}
latex.table <- rbind(latex.table,
                     "\\hline")

latex.file <- paste("tables/table_PH_stocks.txt",sep="")
write(latex.table, file = latex.file)

```

This final block explores an additional calibration of the preferred-habitat model.

```{r ccc}
# Purpose: Check the additional preferred-habitat calibration and perpetuity price.

# Check perpetuity price:

# Phi[3,3] <- 0.5

# Sigma <- diag(c(diag(model_VV$Sigma),0.04^2))
# Sigma <- .0000001*diag(3)
# Sigma[1:2,1:2] <- model_VV$Sigma
# Sigma[2,2] <- 0
# Sigma[1,1] <- 0
# Sigma[3,3] <- 0.05

# mu_k <- 0

model <- list(gamma = a,mu = mu,
              Phi = Phi,
              Sigma = Sigma,
              alpha = Alpha,
              beta = Beta,
              xi = xi,
              a1 = rbind(model_VV$a1,0),b1 = model_VV$b1,
              omega = 0, mu_K0=mu_k, mu_K1=mu_K1,
              beta_K=.3, alpha_K=alpha_K, xi_K=xi_K)

RES <- solve_PH_TSM(model,max_iter = 300)

# Compute perpetuity price "manually":
psi.parameterization.Q <- list(mu  = RES$muQ,
                               Phi = RES$PhiQ,
                               Sigma = Sigma)

HH <- 400
loadings.before.t.plus.h <-  mu_K1 + matrix(c(-1,0,0),3,1)
aux.loadings <- 
  reverse_MHLT(psi_GaussianVAR,
               u1=mu_K1,
               u2=loadings.before.t.plus.h,
               H=HH,
               psi.parameterization.Q)
# Multiply by exp(-r_t):
Loadings <- matrix(c(-1,0,0),3,HH)
Loadings[,1:HH] <- Loadings[,1:HH] + aux.loadings$A[,,1:HH]

B <- c(aux.loadings$B) + (1:HH)*(- b1 + mu_k)

# select any potential value for state vector:
X <- matrix(c(0.0,0.02,0),ncol=1)

B <- exp(B + t(Loadings) %*% X)
sum(B)
exp(RES$mu_zeta0 + t(RES$mu_zeta1)%*%X)

```
