# Estimation {#Estimation}

## Maximum likelihood estimation with observed variables

Indicate below whether the models' parameterizations are re-estimated (otherwise), use pre-estimated parameters.

```{r IndicEstim_FAKE, eval=FALSE}
# Purpose: Show the switch used when re-estimating the estimation examples.
# Switch to TRUE when re-estimating the examples.
indic_reestim <- TRUE # if FALSE, then use results in folder /stored_results/
```

```{r IndicEstim, include=FALSE}
# Purpose: Set the default estimation examples to use stored estimates.
# ==============================================================================
# ==============================================================================
indic_reestim <- FALSE # if FALSE, then use results in folder /stored_results/
# ==============================================================================
# ==============================================================================
```



### The likelihood surface and basic ingredients of maximum-likelihood estimation

This example introduces the likelihood surface and basic ingredients of maximum-likelihood estimation.

```{r MLE0, warning=FALSE, message=FALSE, fig.width=10, fig.height=4}
# Purpose: This example introduces the likelihood surface and basic ingredients of
# maximum-likelihood estimation
library(DTAM)
library(lubridate) # to handle dates
library(vars)
library(optimx)

# ==============================================================================
# Prepare dataset
# ------------------------------------------------------------------------------

data("SPF", package = "DTAM")
data("Data_Macro_US_quarterly", package = "DTAM")
data("YC_LW", package = "DTAM")

DATA <- merge(SPF,Data_Macro_US_quarterly,by="date")
DATA <- merge(DATA,YC_LW,by="date")


freq <- 4 # quarterly frequency

# Create matrix of state variables
wy <- DATA[,c("CPIAUCSL","yld_3m",  # price index, 3-month rate
              "yld_1y","yld_2y","yld_5y","yld_10y", # nominal yields (LW database)
              "BILL1")] # SPF forecast, 3-month rate, 1 year ahead.
maturities_in_model_periods <- c(1,2,5,10)*freq # maturities of yields
horizons_in_model_periods   <- c(1)*freq # horizon of forecast

# Compute inflation:
TT <- dim(DATA)[1]
wy$CPIAUCSL <- c(rep(NaN,4),log(DATA$CPIAUCSL[5:TT]/
                                  DATA$CPIAUCSL[1:(TT-4)]))

# Make yields consistent with model frequency:
wy[,2:dim(wy)[2]] <- wy[,2:dim(wy)[2]]/(freq*100) 

# Remove entries with NaNs:
date_index <- complete.cases(wy)
wy <- wy[date_index,] # remove missing observations
vec_dates <- DATA$date[date_index]

# State vector:
w <- as.matrix(wy[,1:2])
mean_r <- mean(w[,"yld_3m"]) # keep mean of short-term rate
# Demean state variables (hence no need to estimate mu):
w <- w - t(matrix(apply(w,2,mean),dim(w)[2],dim(w)[1]))

# Observed variables:
y <- as.matrix(wy[,(dim(w)[2]+1):dim(wy)[2]])

n <- dim(w)[2] # dimension of w
# ==============================================================================


# Estimate VAR model (P dynamics; used as starting value)
estVAR <- VAR(w,p=1,type = "none")
Phi <- Acoef(estVAR)[[1]]
mu  <- matrix(0,n,1)
Sigma <- var(residuals(estVAR)) # covariance matrix of residuals
modelP <- list(mu=mu,Phi=Phi,Sigma=Sigma)

# Specification of short-term rates:
xi0 <- mean_r
xi1 <- matrix(0,n,1)
xi1[which(colnames(w)=="yld_3m")] <- 1

# Initial specif of prices of risk:
lambda0 <- matrix(0,n,1)
lambda1 <- matrix(0,n,n)

# ==============================================================================
# Specifiy functions used in the estimation:
# ------------------------------------------------------------------------------

param2model_GATSM <- function(param,n,
                              modelP = NaN){
  # If modelP is not NULL, then it is not updated.
  
  mu <- matrix(0,n)
  
  if(is.na(modelP[1])){
    Phi <- matrix(param[1:n^2],n,n)
    nb <- n^2
    Sigma12 <- matrix(0,n,n)
    Sigma12[!upper.tri(Sigma12)] <- param[(nb+1):(nb+n*(n+1)/2)]
    Sigma <- Sigma12 %*% t(Sigma12)
    nb <- nb + n*(n+1)/2
  }else{
    Phi <- modelP$Phi
    Sigma <- modelP$Sigma
    nb <- 0
  }
  lambda0 <- matrix(param[(nb+1):(nb+n)],ncol=1)
  lambda1 <- matrix(param[(nb+n+1):(nb+n+n^2)],n,n)
  nb <- nb + n + n^2
  Sigma12 <- t(chol(Sigma))
  muQ  <- mu  + Sigma12 %*% lambda0
  PhiQ <- Phi + Sigma12 %*% lambda1
  
  modelP <- list(mu=mu,Phi=Phi,Sigma=Sigma)
  modelQ <- list(mu=muQ,Phi=PhiQ,Sigma=Sigma)
  
  # Same standard deviation for yields and forecasts:
  std_dev <- param[nb+1]
  
  return(list(modelP=modelP,
              modelQ=modelQ,
              lambda0 = lambda0,lambda1 = lambda1,
              std_dev=std_dev))
}

make_StateSpace_GATSM <- function(models,
                                  xi0,xi1,
                                  maturities_in_model_periods,
                                  horizons_in_model_periods=NaN){
  # This function builds the state-space model assocoated with
  # a Gaussian Affine Term Structure models.
  
  modelP <- models$modelP
  modelQ <- models$modelQ
  
  resAB <- compute_AB_classical(xi0,xi1,
                                H=max(maturities_in_model_periods), # maximum maturity
                                psi=psi_GaussianVAR,
                                psi.parameterization=modelQ)
  # Affine representation of yields, for selected maturities:
  A <- resAB$a[,1,maturities_in_model_periods]
  B <- resAB$b[,1,maturities_in_model_periods]
  
  # Short-term rate forecasts:
  resAB_P <- compute_AB_classical(xi0,xi1,
                                  H=max(maturities_in_model_periods,
                                        horizons_in_model_periods), # maximum horizon
                                  psi=psi_GaussianVAR,
                                  psi.parameterization=modelP)
  
  if(!is.na(horizons_in_model_periods[1])){
    # Affine representation of short-term rate forecasts, for selected maturities:
    A_BILL <- resAB_P$a[,1,horizons_in_model_periods]
    B_BILL <- resAB_P$b[,1,horizons_in_model_periods]
  }else{
    A_BILL <- NaN
    B_BILL <- NaN
  }
  
  # Affine representation of yields under P, for selected maturities:
  A_P <- resAB_P$a[,1,maturities_in_model_periods]
  B_P <- resAB_P$b[,1,maturities_in_model_periods]
  
  return(list(A = A, B = B,
              A_P = A_P, B_P = B_P,
              A_BILL = A_BILL, B_BILL = B_BILL,
              resAB = resAB,
              resAB_P = resAB_P))}

loglik_GATSM_observed_factors <- function(param,y,w,
                                          maturities_in_model_periods,
                                          horizons_in_model_periods,
                                          modelP = NaN,
                                          max_lambdalambda_squared = NaN){
  # This function computes the log-likelihood associated with a Gaussian
  # Affine Term Structure model with observed factors (w)
  # y contains observations of yields and forecasts of the short-term yield.
  # If modelP not NaN, then do not estimate P dynamics
  
  n <- dim(w)[2]
  TT <- dim(w)[1]
  
  models <- param2model_GATSM(param,n,modelP)
  
  SS <- make_StateSpace_GATSM(models,
                              xi0,xi1,
                              maturities_in_model_periods,
                              horizons_in_model_periods)
  
  nb_mat <- length(maturities_in_model_periods)
  nb_horiz <- length(horizons_in_model_periods)
  
  yds_model <- w %*% SS$A + t(matrix(SS$B,nb_mat,TT))
  BILL_model <- w %*% SS$A_BILL + t(matrix(SS$B_BILL,nb_horiz,TT))
  
  # Measurement errors:
  eta <- cbind(yds_model,BILL_model) - y # pricing errors
  
  # w's innovations:
  epsilon <- w[2:TT,] - w[1:(TT-1),] %*% t(models$modelP$Phi)
  
  logL_measur  <- log_mvdnorm(eta, diag(rep(models$std_dev^2,nb_mat+nb_horiz)))
  logL_transit <- log_mvdnorm(epsilon, Sigma)
  
  if(is.na(max_lambdalambda_squared)){
    penalty <- 0
  }else{
    lambda0 <- models$lambda0
    lambda1 <- models$lambda1
    Phi <- models$modelP$Phi
    Var_w <- matrix(
      solve(diag(n^2) - Phi %x% Phi) %*% c(models$modelP$Sigma),
      n,n)
    sum_lambdalambda <- 
      sum(models$lambda0^2) + sum(diag(lambda1 %*% Var_w %*% t(lambda1)))
    penalty <- 10*(sum_lambdalambda > max_lambdalambda_squared) * 
      (sum_lambdalambda - max_lambdalambda_squared)^2
  }
  
  return(-(logL_measur + logL_transit) + penalty)
}
# ==============================================================================
```

The following chunk estimates the model while imposing restrictions on $\boldsymbol\lambda_0$ and $\boldsymbol\lambda_1$ (which disciplines the estimation).

### Initialization affects likelihood-based estimation results

This example shows how initialization affects likelihood-based estimation results.

```{r fig_MLE_ini1, warning=FALSE, message=FALSE, fig.width=10, fig.height=4}
# Purpose: This example shows how initialization affects likelihood-based estimation results

max_lambdalambda_squared <- 1 # If NaN, no restrictions imposed on lambda_t
# ==============================================================================

# ==============================================================================
# Estimation by MLE
# ------------------------------------------------------------------------------

if(indic_reestim){
  std_dev <- sd(y)
  for(indic_update_P in c(FALSE,TRUE)){
    
    if(indic_update_P){
      modelP <- NaN
      Sigma12 <- t(chol(Sigma))
      param_est <- c(Phi,Sigma12[!upper.tri(Sigma12)])
    }else{
      modelP <- list(mu = matrix(0,dim(w)[2]),
                     Phi = Phi, Sigma = Sigma)
      param_est <- NULL
    }
    param_est <- c(param_est,lambda0,lambda1,std_dev)
    
    if(!indic_update_P){
      nb_loop <- 2
    }else{
      nb_loop <- 5
    }
    
    for(i in 1:nb_loop){
      print(paste("--- Loop ",i," ---",sep=""))
      for(method in c("Nelder-Mead","nlminb")){
        res.optim <- optimx(par=param_est,
                            fn=loglik_GATSM_observed_factors,
                            y=y,w=w,
                            maturities_in_model_periods=maturities_in_model_periods,
                            horizons_in_model_periods=horizons_in_model_periods,
                            modelP = modelP,
                            max_lambdalambda_squared = max_lambdalambda_squared,
                            method=method,
                            control=list(trace=FALSE,
                                         maxit=ifelse(method=="Nelder-Mead",4000,20),
                                         kkt = FALSE))
        print(paste("--- Log-lik (",method,"): ",round(res.optim$value,2)," ---",sep=""))
        param_est  <- c(as.matrix(res.optim)[1:length(param_est)])
      }
    }
    
    models <- param2model_GATSM(param_est,n,
                                modelP = modelP)
    
    lambda0 <- models$lambda0
    lambda1 <- models$lambda1
    std_dev <- models$std_dev
  }
  save(param_est,
       file=
         paste("stored_results/param_GATSM_obs_varia_lambda_",
               max_lambdalambda_squared,".Rdat",sep=""))
}else{
  load(file=paste("stored_results/param_GATSM_obs_varia_lambda_",
                  max_lambdalambda_squared,".Rdat",sep=""))
}
# ==============================================================================


# ==============================================================================
# Prepare plots
# ------------------------------------------------------------------------------

n <- dim(w)[2]
TT <- dim(w)[1]
models <- param2model_GATSM(param_est,n)
SS <- make_StateSpace_GATSM(models,
                            xi0,xi1,
                            maturities_in_model_periods,
                            horizons_in_model_periods)
nb_mat <- length(maturities_in_model_periods)
nb_horiz <- length(horizons_in_model_periods)
yds_model   <- w %*% SS$A      + t(matrix(SS$B,nb_mat,TT))
BILL_model  <- w %*% SS$A_BILL + t(matrix(SS$B_BILL,nb_horiz,TT))
yds_model_P <- w %*% SS$A_P    + t(matrix(SS$B_P,nb_mat,TT))


par(mfrow=c(1,2))
par(plt=c(.1,.95,.1,.85))
count <- 0
for(maturity in c(2,10)){
  count <- count + 1
  main.t <- paste("(",letters[count],") Maturity: ",maturity," years",sep="")
  
  m <- which(maturities_in_model_periods==freq*maturity)
  vari <- paste("yld_",maturity,"y",sep="")
  plot(vec_dates,freq*100*y[,vari],type="l",lty=1,lwd=2,col="dark grey",
       ylim=c(-1,max(freq*100*y[,vari])),
       xlab="",ylab="",las=1,main=main.t)
  lines(vec_dates,freq*100*yds_model[,m],lty=1,lwd=2)
  lines(vec_dates,freq*100*yds_model_P[,m],lty=3,lwd=2)
  grid()
  
  if(count == 1){
    legend("topright",
           legend=c("Data","Model-implied","Model-implied (Expectation Hypothesis)"),
           lty=c(1,1,3),
           col=c("dark grey","black","black"),
           lwd=c(2,2,2),bty = "n")
  }
}
```

In the following chunk, we remove the restrictions on $\boldsymbol\lambda_0$ and $\boldsymbol\lambda_1$.

### Observed variables with their fitted counterparts under the estimated model

This example compares observed variables with their fitted counterparts under the estimated model.

```{r fig_MLE_obs_variables, warning=FALSE, message=FALSE, fig.width=8, fig.height=4}
# Purpose: This example compares observed variables with their fitted counterparts under the
# estimated model

# ==============================================================================
max_lambdalambda_squared <- NaN # If NaN, no restrictions imposed on lambda_t
# ==============================================================================

# ==============================================================================
# Estimation by MLE
# ------------------------------------------------------------------------------

if(indic_reestim){
  nb_loop <- 20
  
  for(i in 1:nb_loop){
    print(paste("--- Loop ",i," ---",sep=""))
    for(method in c("Nelder-Mead","nlminb")){
      res.optim <- optimx(par=param_est,
                          fn=loglik_GATSM_observed_factors,
                          y=y,w=w,
                          maturities_in_model_periods=maturities_in_model_periods,
                          horizons_in_model_periods=horizons_in_model_periods,
                          modelP = modelP,
                          max_lambdalambda_squared = max_lambdalambda_squared,
                          method=method,
                          control=list(trace=FALSE,
                                       maxit=ifelse(method=="Nelder-Mead",4000,20),
                                       kkt = FALSE))
      print(paste("--- Log-lik (",method,"): ",round(res.optim$value,2)," ---",sep=""))
      param_est  <- c(as.matrix(res.optim)[1:length(param_est)])
    }
  }
  
  save(param_est,
       file=
         paste("stored_results/param_GATSM_obs_varia_lambda_",
               max_lambdalambda_squared,".Rdat",sep=""))
}else{
  load(file=paste("stored_results/param_GATSM_obs_varia_lambda_",
                  max_lambdalambda_squared,".Rdat",sep=""))
}
# ==============================================================================


# ==============================================================================
# Prepare plots
# ------------------------------------------------------------------------------

n <- dim(w)[2]
TT <- dim(w)[1]
models <- param2model_GATSM(param_est,n)
SS <- make_StateSpace_GATSM(models,
                            xi0,xi1,
                            maturities_in_model_periods,
                            horizons_in_model_periods)
nb_mat <- length(maturities_in_model_periods)
nb_horiz <- length(horizons_in_model_periods)
yds_model   <- w %*% SS$A      + t(matrix(SS$B,nb_mat,TT))
BILL_model  <- w %*% SS$A_BILL + t(matrix(SS$B_BILL,nb_horiz,TT))
yds_model_P <- w %*% SS$A_P    + t(matrix(SS$B_P,nb_mat,TT))


par(mfrow=c(1,2))
par(plt=c(.1,.95,.1,.85))
count <- 0
for(maturity in c(2,10)){
  count <- count + 1
  main.t <- paste("(",letters[count],") Maturity: ",maturity," years",sep="")
  
  m <- which(maturities_in_model_periods==freq*maturity)
  vari <- paste("yld_",maturity,"y",sep="")
  plot(vec_dates,freq*100*y[,vari],type="l",lty=1,lwd=2,col="dark grey",
       ylim=c(-1,max(freq*100*y[,vari])),
       xlab="",ylab="",las=1,main=main.t)
  lines(vec_dates,freq*100*yds_model[,m],lty=1,lwd=2)
  lines(vec_dates,freq*100*yds_model_P[,m],lty=3,lwd=2)
  grid()
  
  if(count == 1){
    legend("topright",
           legend=c("Data","Model-implied","Model-implied (Expect. Hyp.)"),
           lty=c(1,1,3),
           col=c("dark grey","black","black"),
           lwd=c(2,2,2),bty = "n")
  }
}
```

## Maximum Likelihood with the Kalman filter

In this section, we take the same data and estimate a model with latent state vector, using the Kalman filter.

In this example, we employ the results of Example \@ref(exm:nominalBth) in the context described by Christensen, Diebold, and Rudebusch (2009). Specifically, we consider a three factor model following a Gaussian VAR:
$$
w_t = \left[\begin{array}{c}X_{1,t}\\X_{2,t}\\X_{3,t}\end{array}\right] = 
\left[\begin{array}{ccc}
\lambda_1 & 0 & 0\\
0&1-\lambda_2&\lambda_2\\
0&0&1-\lambda_2\end{array}\right]
\left[\begin{array}{c}X_{1,t-1}\\X_{2,t-1}\\X_{2,t-1}\end{array}\right] +
\left[\begin{array}{ccc}
\sigma_{11} & 0 & 0\\
\sigma_{21}&\sigma_{22}&0\\
\sigma_{31}&\sigma_{32}&\sigma_{33}\end{array}\right]
\left[\begin{array}{c}\varepsilon_{1,t}\\\varepsilon_{2,t}\\\varepsilon_{3,t}\end{array}\right],
$$
where $\left[\begin{array}{c}\varepsilon_{1,t}\\\varepsilon_{2,t}\\\varepsilon_{3,t}\end{array}\right] \sim \,i.i.d.\, \mathcal{N}(0,Id)$.

The nominal short-term rate is given by $i_t = X_{1,t}+X_{2,t}$.

Christensen, Diebold, and Rudebusch (2009) show that such specifications allow to approximate the widely-used Nelson-Siegel yield curve specification.^[In the context of the Nelson-Siegel model, the zero-coupon yield of maturity $h$ is given by $\beta_0 + \beta_1\left(\frac{1 - \exp(-\lambda h)}{\lambda h}\right)+\beta_2\left(\frac{1 - \exp(-\lambda h)}{\lambda h}-\exp(-\lambda h)\right)$, where $\beta_0$, $\beta_1$, $\beta_2$, and $\lambda$ are parameters.]



### Estimated latent factors from the maximum-likelihood exercise

This example reports estimated latent factors from the maximum-likelihood exercise.

```{r fig_MLE_latent_variables, warning=FALSE, message=FALSE, fig.width=9, fig.height=5}
# Purpose: This example reports estimated latent factors from the maximum-likelihood exercise

# Dataset ----------------------------------------------------------------------

# Create matrix of observables
observables <- DATA[,c("yld_3m","yld_1y","yld_2y","yld_5y","yld_10y", # nominal yields (LW database)
                       "BILL1","BILL10")] # SPF forecast, 3-month rate, 1 and 10 years ahead.
maturities_in_model_periods <- c(.25,1,2,5,10)*freq # maturities of yields
horizons_in_model_periods   <- c(1,10)*freq # horizon of forecast

# Make yields consistent with model frequency:
observables <- as.matrix(observables)/(freq*100) 

# vector of dates:
vec_dates <- DATA$date

# Estimation -------------------------------------------------------------------
# Constraint on unreasonable Sharpe Ratios:
max_lambdalambda_squared <- NaN

# Number of factors
n <- 3
TT <- dim(observables)[1]

param2model_AFNS <- function(param){
  # Parameterize the three-factor model of Christensen, Diebold, and Rudebusch (2009).
  # This model is called Arbitrage-Free Nelson-Siegel (AFNS)
  
  n <- 3 # three factor model
  mu <- matrix(param[1:n],n)
  nb <- n
  
  lambda1P <- exp(param[nb+1])/(1+exp(param[nb+1]))
  lambda2P <- exp(param[nb+2])/(1+exp(param[nb+2]))
  Phi <- diag(c(1-lambda1P,1-lambda2P,1-lambda2P))
  Phi[2,3] <- lambda2P
  nb <- nb + 2
  
  # Phi <- matrix(param[(nb+1):(nb+n^2)],n,n)
  # nb <- nb + n^2
  
  Sigma12 <- matrix(0,n,n)
  Sigma12[!upper.tri(Sigma12)] <- param[(nb+1):(nb+n*(n+1)/2)]
  Sigma <- Sigma12 %*% t(Sigma12)
  nb <- nb + n*(n+1)/2
  
  muQ  <- matrix(0,n,1)
  lambda1Q <- exp(param[nb+1])/(1+exp(param[nb+1]))
  lambda2Q <- exp(param[nb+2])/(1+exp(param[nb+2]))
  PhiQ <- diag(c(1-lambda1Q,1-lambda2Q,1-lambda2Q))
  PhiQ[2,3] <- lambda2Q
  nb <- nb + 2
  
  lambda0 <- solve(Sigma12) %*% (muQ - mu)
  lambda1 <- solve(Sigma12) %*% (PhiQ - Phi)
  
  modelP <- list(mu=mu,Phi=Phi,Sigma=Sigma,Sigma12=Sigma12)
  modelQ <- list(mu=muQ,Phi=PhiQ,Sigma=Sigma,Sigma12=Sigma12)
  
  xi0 <- param[nb+1]
  xi1 <- matrix(c(1,1,0),n,1)
  nb <- nb + 1
  
  # Same standard deviation for yields and forecasts:
  std_dev <- max(exp(param[nb+1]),.0001)
  
  return(list(modelP=modelP,
              modelQ=modelQ,
              xi0 = xi0, xi1 = xi1,
              lambda0 = lambda0,lambda1 = lambda1,
              std_dev=std_dev))
}


loglik_AFNS <- function(param,observables,
                        maturities_in_model_periods,
                        horizons_in_model_periods,
                        max_lambdalambda_squared = NaN){
  # This function computes the log-likelihood associated with a Gaussian
  # Affine Term Structure model with observed factors (w)
  # y contains observations of yields and forecasts of the short-term yield.
  # If modelP not NaN, then do not estimate P dynamics
  
  models <- param2model_AFNS(param)
  n <- dim(models$modelP$Phi)[2]
  
  # Unconditional variance of w:
  Phi <- models$modelP$Phi
  Var_w <- matrix(
    solve(diag(n^2) - Phi %x% Phi) %*% c(models$modelP$Sigma),
    n,n)
  
  SS <- make_StateSpace_GATSM(models,
                              models$xi0,models$xi1,
                              maturities_in_model_periods,
                              horizons_in_model_periods)
  
  nb_mat   <- length(maturities_in_model_periods)
  nb_horiz <- length(horizons_in_model_periods)
  
  # Define matrices needed in the Kalman_filter procedures:
  nu_t <- t(matrix(models$modelP$mu,n,TT))
  H <- models$modelP$Phi
  G <- t(cbind(SS$A,SS$A_BILL))
  mu_t <- cbind(t(matrix(SS$B,nb_mat,TT)),
                t(matrix(SS$B_BILL,nb_horiz,TT)))
  N <- models$modelP$Sigma12
  M <- diag(rep(models$std_dev,dim(observables)[2]))
  Sigma_0 <- Var_w # unconditional variance of w
  rho_0 <- matrix(0,n,1) # unconditional mean of w
  
  filter.res   <- Kalman_filter(observables,nu_t,H,N,mu_t,G,M,Sigma_0,rho_0)
  
  if(is.na(max_lambdalambda_squared)){
    penalty <- 0
  }else{
    lambda0 <- models$lambda0
    lambda1 <- models$lambda1
    sum_lambdalambda <- 
      sum(models$lambda0^2) + sum(diag(lambda1 %*% Var_w %*% t(lambda1)))
    penalty <- 10*(sum_lambdalambda > max_lambdalambda_squared) * 
      (sum_lambdalambda - max_lambdalambda_squared)^2
  }
  
  penalty <- penalty + 100*sum(
    (abs(param)>20)*(abs(param)-20)
  )
  
  loss <- - sum(filter.res$loglik.vector) + penalty
  
  return(loss)
}

mu <- matrix(0,n,1)
lambda1P <- .1
lambda2P <- .1
lambda1Q <- .1
lambda2Q <- .1
Sigma <- .01^2 * diag(n)
Sigma12 <- t(chol(Sigma))
param_est <- c(mu,
               log(lambda1P/(1-lambda1P)),log(lambda2P/(1-lambda2P)),
               Sigma12[!upper.tri(Sigma12)],
               log(lambda1Q/(1-lambda1Q)),log(lambda2Q/(1-lambda2Q)),
               .06/freq,#r_infty
               log(.05))# std_dev

# Initialization:
param <- param_est

if(indic_reestim){
  nb_loop <- 3
  
  for(i in 1:nb_loop){
    print(paste("--- Loop ",i," ---",sep=""))
    for(method in c("nlminb","Nelder-Mead")){
      # for(method in c("nlminb")){
      res.optim <- optimx(par=param_est,
                          fn=loglik_AFNS,
                          observables=observables,
                          maturities_in_model_periods=maturities_in_model_periods,
                          horizons_in_model_periods=horizons_in_model_periods,
                          max_lambdalambda_squared = max_lambdalambda_squared,
                          method=method,
                          control=list(trace=TRUE,
                                       maxit=ifelse(method=="Nelder-Mead",2000,100),
                                       kkt = FALSE)
      )
      print(paste("--- Log-lik: ",round(res.optim$value,2)," ---",sep=""))
      param_est  <- c(as.matrix(res.optim)[1:length(param_est)])
    }
  }
  
  save(param_est,
       file=
         paste("stored_results/param_AFNS_lambda_",
               max_lambdalambda_squared,".Rdat",sep=""))
}else{
  load(file=paste("stored_results/param_AFNS_lambda_",
                  max_lambdalambda_squared,".Rdat",sep=""))
}
# ==============================================================================


# ==============================================================================
# Prepare plots
# ------------------------------------------------------------------------------

TT <- dim(observables)[1]

models <- param2model_AFNS(param_est)
n <- dim(models$modelP$Phi)[2]

# Unconditional variance of w:
Phi <- models$modelP$Phi
Var_w <- matrix(
  solve(diag(n^2) - Phi %x% Phi) %*% c(models$modelP$Sigma),
  n,n)

SS <- make_StateSpace_GATSM(models,
                            models$xi0,models$xi1,
                            maturities_in_model_periods,
                            horizons_in_model_periods)

nb_mat   <- length(maturities_in_model_periods)
nb_horiz <- length(horizons_in_model_periods)

# Define matrices needed in the Kalman_filter procedures:
nu_t <- t(matrix(models$modelP$mu,n,TT))
H <- models$modelP$Phi
G <- t(cbind(SS$A,SS$A_BILL))
mu_t <- cbind(t(matrix(SS$B,nb_mat,TT)),
              t(matrix(SS$B_BILL,nb_horiz,TT)))
N <- models$modelP$Sigma12
M <- diag(rep(models$std_dev,dim(observables)[2]))
Sigma_0 <- Var_w # unconditional variance of w
rho_0 <- matrix(0,n,1) # unconditional mean of w

filter.res   <- Kalman_filter(observables,nu_t,H,N,mu_t,G,M,Sigma_0,rho_0)
ww <- filter.res$r
yds_model   <- ww %*% SS$A      + t(matrix(SS$B,nb_mat,TT))
BILL_model  <- ww %*% SS$A_BILL + t(matrix(SS$B_BILL,nb_horiz,TT))
yds_model_P <- ww %*% SS$A_P    + t(matrix(SS$B_P,nb_mat,TT))

par(mfrow=c(1,2))
par(plt=c(.1,.95,.1,.85))
count <- 0
for(maturity in c(2,10)){
  count <- count + 1
  main.t <- paste("(",letters[count],") Maturity: ",maturity," years",sep="")
  
  m <- which(maturities_in_model_periods==freq*maturity)
  plot(vec_dates,freq*100*observables[,m],type="l",lty=1,lwd=2,col="dark grey",
       ylim=c(0,max(freq*100*yds_model[,m])),
       xlab="",ylab="",las=1,main=main.t)
  lines(vec_dates,freq*100*yds_model[,m],lty=1,lwd=2)
  lines(vec_dates,freq*100*yds_model_P[,m],lty=3,lwd=2)
  if(maturity==10){
    points(vec_dates,DATA$BILL10,pch=19,lwd=1,col="dark grey")
  }
  grid()
  
  if(count == 2){
    legend("bottomleft",
           legend=c("Data","Model-implied","Model-implied (Expect. Hyp.)",
                    "SPF: 10-yr SPF fcst of 3-mth rate"),
           lty=c(1,1,3,NaN),
           pch=c(NaN,NaN,NaN,19),
           col=c("dark grey","black","black","dark grey"),
           lwd=c(2,2,2),bty = "n")
  }
}
```


## Kitagawa-Hamilton

This example estimates a regime-switching stock-return model and visualizes the inferred states.

```{r fig_RS_stocks1,message=FALSE,warning=FALSE,fig.width=8,fig.height=7}
# Purpose: This example estimates a regime-switching stock-return model and visualizes the
# inferred states

library(DTAM)
library(expm)
library(mvtnorm)
library(optimx)
library(tis) # for NBER recession dates

param2model <- function(param,J=2){
  nb <- 0 # number of parameters already used
  # matrix of transition probabilities:
  Pi <- matrix(0,J,J)
  
  diag(Pi) <- exp(param[1:J])/(1+exp(param[1:J]))
  Pi[1,2] <- 1- Pi[1,1]
  Pi[J,J-1] <- 1- Pi[J,J]
  nb <- nb + J
  
  if(J>2){
    par_aux <- param[(nb+1):(nb+J-2)]
    par_aux <- exp(par_aux)/(1+exp(par_aux))
    if(length(par_aux)==1){# in thius case, J=3
      aux <- par_aux*(1-Pi[2,2])
    }else{
      aux <- diag(c(par_aux)*(1-diag(Pi[2:(J-1),2:(J-1)])))
    }
    Pi[2:(J-1),1:(J-2)] <- Pi[2:(J-1),1:(J-2)] + aux
    
    sum_rows <- apply(matrix(Pi[2:(J-1),],J-2,J),1,sum)
    if(length(sum_rows)==1){
      aux <- 1 - sum_rows
    }else{
      aux <- diag(J-2) - diag(c(sum_rows))
    }
    Pi[2:(J-1),3:J] <- Pi[2:(J-1),3:J] + aux
    
    nb <- nb + J - 2
  }
  
  phi <- .999*exp(param[nb+1])/(1+exp(param[nb+1]))
  nb <- nb + 1
  
  mu_r <- matrix(param[(nb+1):(nb+J)],ncol=1)/100
  nb <- nb + J
  
  sigma_r <- matrix(param[(nb+1):(nb+J)],ncol=1)/100
  nb <- nb + J
  
  mu_s2 <- matrix(param[(nb+1):(nb+3*J+1)],ncol=1)/100
  nb <- nb + 3*J + 1
  
  lambda <- matrix(param[(nb+1):(nb+3*J+1)],ncol=1)/100
  nb <- nb + 3*J + 1
  
  
  model <- list(Pi = Pi,
                phi = phi,
                mu_r = mu_r,
                sigma_r = sigma_r,
                mu_s2 = mu_s2,
                lambda = lambda)
  return(model)
}

model2param <- function(model){
  
  Pi <- model$Pi
  
  J <- dim(Pi)[1]
  
  param <- log(diag(Pi)/(1-diag(Pi)))
  
  if(J>2){
    if(J==3){
      aux <- Pi[2,1]/(1-Pi[2,2])
    }else{
      aux <- diag(Pi[2:(J-1),1:(J-2)])/(1-diag(Pi)[2:(J-1)])
    }
    param <- c(param,
               log(aux/.999/(1-aux/.999)))
  }
  
  param <- c(param,
             log(model$phi/(1-model$phi)),
             100*model$mu_r,
             100*model$sigma_r,
             100*model$mu_s2,
             100*model$lambda)
  return(param)
}

solve_model_RS_stocks <- function(model){
  model$mu_s0 <- 0
  J <- dim(model$Pi)[1]
  e_1 <- matrix(c(1,rep(0,3*J)),1+3*J,1)
  model$mu_s1 <- e_1 + psi_RS_stock(model$lambda,model)$a - 
    psi_RS_stock(model$lambda + model$mu_s2,model)$a
  return(model)
}

psi_RS_stock <- function(u,model){
  # Laplace transform of the state vector w_t, with
  # w_t = (r_t,z_t',eps_{r,t}z_t',eps_{s,t}z_t')'
  # u is n x k, where n = 1+3*J. Indeed, dimension of J is 1+3*J
  # k is the number of vectors on which one wants to evaluate the LT of w
  
  J <- (dim(u)[1]-1)/3
  k <- dim(u)[2]
  
  Pi <- model$Pi
  phi <- model$phi
  
  mu_r    <- model$mu_r
  sigma_r <- model$sigma_r
  mu_s2   <- model$mu_s2
  lambda  <- model$lambda
  
  u_r <- matrix(u[1,],nrow=1)
  U_z <- matrix(u[(1+0*J+1):(1+1*J),],J,k)
  U_r <- matrix(u[(1+1*J+1):(1+2*J),],J,k)
  U_s <- matrix(u[(1+2*J+1):(1+3*J),],J,k)
  
  a <- rbind(
    u_r*phi,
    log(Pi %*% exp((matrix(1,J,1)%*%u_r)*(mu_r %*% matrix(1,1,k)) + U_z + 
                     .5*((matrix(1,J,1)%*%u_r) * sigma_r %*% matrix(1,1,k) + U_r)^2 +
                     .5*U_s^2)),
    matrix(0,2*J,k)
  )
  return(list(a=a,b=0))
}

simul_model_RS_stocks <- function(model,TT){
  # This function simulates the regime-switching stock-return model.
  
  J <- dim(model$Pi)[1]
  
  z <- simul_RS(model$Pi,TT)
  
  # Compute unconditional distribution of z:
  Paux <- t(model$Pi)
  for(i in 1:10){
    Paux <- Paux %^% 2
  }
  Ez <- matrix(Paux[,1],ncol=1)
  
  # mean STR:
  Emu_r <- t(model$mu_r) %*% Ez
  Er    <- Emu_r/(1-model$phi)
  
  epsilon_r <- matrix(rnorm(TT),ncol=1)
  epsilon_s <- matrix(rnorm(TT),ncol=1)
  
  w <- cbind(0,z,
             z*(epsilon_r%*%matrix(1,1,J)),
             z*(epsilon_s%*%matrix(1,1,J)))
  
  r <- Er
  for(t in 1:TT){
    r <- z[t,] %*% model$mu_r + phi*r + (z[t,] %*% model$sigma_r)*epsilon_r[t]
    w[t,1] <- r
  }
  
  w_1 <- rbind(NaN,w[1:(TT-1),])
  
  rs <- model_solved$mu_s0 + w_1%*%model_solved$mu_s1 + w%*%model_solved$mu_s2
  
  return(list(w=w,z=z,
              rs=rs,r=w[,1]))
}

filter_RS_stocks <- function(model_solved,Y,indic_smooth=FALSE){
  # Y contains the dataset:
  #  - 1st column: short-term rate r_t
  #  - 2nd column: stock return rs_t
  
  TT <- dim(Y)[1]
  J <- dim(model_solved$Pi)[1]
  
  mu_r_extended    <- model_solved$mu_r    %x% matrix(1,J,1)
  sigma_r_extended <- model_solved$sigma_r %x% matrix(1,J,1)
  
  # Compute the matrix of transition probabilities of z_{t} %x% z_{t-1}:
  PI <- make_Pi_z_kron_z(model_solved$Pi)
  
  # Compute epsilon_r as a function of z_{t} %x% z_{t-1} -----------------------
  r   <- Y[2:TT,1] %*% matrix(1,1,J*J)
  r_1 <- Y[1:(TT-1),1] %*% matrix(1,1,J*J)
  
  sigma_r_epsilon_r <- r - model_solved$phi * r_1 - matrix(1,TT-1,1) %*% t(mu_r_extended)
  epsilon_r <- sigma_r_epsilon_r/(matrix(1,TT-1,1) %*% t(sigma_r_extended))
  
  # Compute epsilon_s as a function of z_{t} %x% z_{t-1} -----------------------
  rs <- Y[2:TT,2] %*% matrix(1,1,J*J)
  
  mu_s1_z_extended <- matrix(1,J,1) %x% matrix(model_solved$mu_s1[2:(J+1)],ncol=1)
  mu_s2_z_extended <- matrix(model_solved$mu_s2[2:(J+1)],ncol=1) %x% matrix(1,J,1)
  
  sigma_sr_extended <- matrix(model_solved$mu_s2[(1+1*J+1):(1+2*J)],ncol=1) %x% matrix(1,J,1)
  sigma_ss_extended <- matrix(model_solved$mu_s2[(1+2*J+1):(1+3*J)],ncol=1) %x% matrix(1,J,1)
  
  sigma_s_epsilon_s <- 
    rs - model_solved$mu_s0 - 
    model_solved$mu_s1[1] * r_1 -
    matrix(1,TT-1,1) %*% t(mu_s1_z_extended) -
    model_solved$mu_s2[1] * r -
    matrix(1,TT-1,1) %*% t(mu_s2_z_extended) -
    (matrix(1,TT-1,1) %*% t(sigma_sr_extended))*epsilon_r
  # epsilon_s <- sigma_s_epsilon_s/(matrix(1,TT-1,1) %*% t(sigma_ss_extended))
  
  Eta <- matrix(NaN,TT-1,J*J)
  for(i in 1:(J*J)){
    sigma <- diag(c(sigma_sr_extended[i]^2,sigma_ss_extended[i]^2))
    sigma_times_epsilon <- cbind(sigma_r_epsilon_r[,i],sigma_s_epsilon_s[,i])
    # Eta[,i] <- dmvnorm(sigma_times_epsilon,sigma=sigma,log = TRUE)
    Eta[,i] <- -
      log(sqrt(2*pi*sigma_sr_extended[i]^2))-.5*sigma_r_epsilon_r[,i]^2/sigma_sr_extended[i]^2-
      log(sqrt(2*pi*sigma_ss_extended[i]^2))-.5*sigma_s_epsilon_s[,i]^2/sigma_ss_extended[i]^2
  }
  
  res_KH <- KH_filter(PI,Eta)
  
  res_KH$ksi_matrix <- res_KH$ksi_matrix %*% (diag(J) %x% matrix(1,J,1))
  
  if(indic_smooth){
    res_KH$ksi_tT_matrix <- KH_smoother(PI,Eta) %*% (diag(J) %x% matrix(1,J,1))
  }
  
  return(res_KH)
}

loglik4optim <- function(param,Y,J=2){
  model <- param2model(param,J=J)
  model_solved <- solve_model_RS_stocks(model)
  res_KH <- filter_RS_stocks(model_solved,Y)
  return(-res_KH$loglik)
}

# Prepare dataset:

P <- Shiller$RealTotReturnPrice
TT <- dim(Shiller)[1]
Shiller$RealTotReturn <- c(NaN,log(P[2:TT]/P[1:(TT-1)]))

DATA <- YC_LW[,c("date","yld_1m")]
DATA$yld_1m <- DATA$yld_1m/1200
names(DATA)[1] <- "Date"
DATA <- merge(DATA,Shiller,by="Date")

TT <- dim(DATA)[1]

Y <- cbind(DATA$yld_1m,DATA$RealTotReturn)
Y <- Y[complete.cases(Y),]
TT <- dim(Y)[1]

# Make initial model:
Pi <- matrix(c(.95,.15,.05,.85),2,2)
mu_s2 <- matrix(c(0,
                  .01,-.03,
                  .01,.01,
                  .01,.03),ncol=1)
lambda <- matrix(c(0,
                   .01,.02,
                   .01,.01,
                   .01,.01),ncol=1)
model <- list(Pi=Pi,
              phi = .9,
              mu_r = matrix(c(.001,-.001),ncol=1),
              sigma_r = matrix(c(.0025,.005),ncol=1),
              mu_s2 = mu_s2,
              lambda = lambda)

param_est <- model2param(model)

if(indic_reestim){
  nb_loop <- 20
  for(i in 1:nb_loop){
    print(paste("--- Loop ",i," ---",sep=""))
    for(method in c("Nelder-Mead","nlminb")){
      res.optim <- optimx(par=param_est,
                          fn=loglik4optim,
                          Y=Y,J=2,
                          method=method,
                          control=list(trace=TRUE,
                                       maxit=ifelse(method=="Nelder-Mead",3000,20),
                                       kkt = FALSE))
      print(paste("--- Log-lik (",method,"): ",round(res.optim$value,2)," ---",sep=""))
      param_est_tempo  <- c(as.matrix(res.optim)[1:length(param_est)])
      if(sum(is.na(param_est_tempo))==0){
        param_est <- param_est_tempo
      }else{
        print("problem")
      }
    }
  }
  save(param_est,file="stored_results/param_RS_stocks.Rdat")
}else{
  load(file="stored_results/param_RS_stocks.Rdat")
}

J <- 2
model <- param2model(param_est,J=J)
model_solved <- solve_model_RS_stocks(model)
res_KH <- filter_RS_stocks(model_solved,Y,indic_smooth = TRUE)
res_KH_smoothed <- 1
par(mfrow=c(1,1))
regime <- 1

vec_dates <- DATA$Date[2:dim(DATA)[1]]

par(mfrow=c(3,1))
par(plt=c(.08,.97,.15,.8))

layout(matrix(1:3, nrow = 3, byrow = TRUE),
       heights = c(2, 1, 1))   # relative heights

proba <- res_KH$ksi_tT_matrix[,regime]
plot(vec_dates,proba,type="l",lwd=2,col="white",
     xaxs = "i",xlab="",ylab="",
     main="(a) Smoothed probability of being in Regime 1",las=1)
grid()
# lines(DATA$Date[2:dim(DATA)[1]],res_KH$ksi_matrix[,regime],col="red",lwd=2)
polygon(c(vec_dates,rev(vec_dates)),c(proba,0*proba),
        border="darkgrey",col="#AAAAAA77")
invisible(make_recessions(col = "#666666FF",MIN=0,MAX=.4))

plot(vec_dates,Y[2:TT,2],type="l",col="white",las=1,
     xaxs = "i",xlab="",ylab="",
     main="(b) Stock returns")
grid()
lines(vec_dates,Y[2:TT,2],lwd=2)
plot(vec_dates,1200*Y[2:TT,1],type="l",col="white",las=1,
     xaxs = "i",xlab="",ylab="",
     main="(c) Short-term nominal rate")
grid()
lines(vec_dates,1200*Y[2:TT,1],lwd=2)

```


### Additional diagnostics for the regime-switching stock model

This example reports additional diagnostics for the regime-switching stock model.

```{r fig_RS_stocks2}
# Purpose: This example reports additional diagnostics for the regime-switching stock model
# Compute expected stock returns
model_solved$n_w <- length(model_solved$mu_s1)
aux <- compute_expect_variance(psi_RS_stock,model_solved)
r <- Y[2:TT,1]
wt <- cbind(r,res_KH$ksi_tT_matrix,matrix(0,TT-1,2*J))
Ewt <- matrix(1,TT-1,1) %*% t(aux$mu) + wt %*% t(aux$Phi)
Ers <- model_solved$mu_s0 + wt %*% model_solved$mu_s1 + Ewt %*% model_solved$mu_s2
Exs <- Ers - r
par(mfrow=c(1,2))
plot(Ers,type="l")
plot(Exs*12,type="l")
```

## Inversion technique plus regime switching

The following example is set at a relatively high frequency (weekly). At this frequency, recursive formulas that involve loops can slow down computations. In the present example, bond yields decompose into two independent components: one driven by a time-homogeneous Markov chain, and the other by a VAR(1) process. Both processes are stable under temporal aggregation, which allows their multi-horizon Laplace transforms to be computed more efficiently by lowering the frequency. The following propositions form the basis of the functions `compute_LT_RS()` (for fast computation of the Markov-switching-based multi-horizon Laplace transforms) and `compute_Gaussian_fast()` (for the Gaussian VAR model).


:::{.theorem #FastPricingVAR name="VAR(1): Fast computation of multi-horizon conditional distribution"}
If 
$$
\mathbf{x}_{t+1} = \boldsymbol\mu + \boldsymbol\Phi\mathbf{x}_{t}+ \boldsymbol\Sigma^{\tfrac{1}{2}}\boldsymbol\varepsilon_{t+1},\quad 
\boldsymbol\varepsilon_{t}\sim\mathcal{N}(\mathbf{0},\mathbf{Id}),
$$

then

$$
\mathbf{x}_{t+h}+\dots+\mathbf{x}_{t+1}\mid \underline{\mathbf{x}_t}\sim 
\mathcal{N}\!\left(\boldsymbol\mu_h + \boldsymbol\Phi_h\mathbf{x}_t,\; \boldsymbol\Sigma_h\right),
$$

with

\[
\begin{aligned}
\boldsymbol\mu_h &= (\mathbf{Id}-\boldsymbol\Phi)^{-1}(h\mathbf{Id} - \boldsymbol\Phi  \mathbf{K}_h)\boldsymbol\mu, \\
\boldsymbol\Phi_h &= \boldsymbol\Phi \mathbf{K}_h, \\
\mbox(vec)(\boldsymbol\Sigma_h) &= h \,\mbox(vec)(\boldsymbol\Sigma^*) 
- \mbox(vec)\!\big(\boldsymbol\Phi \mathbf{K}_h \boldsymbol\Sigma^*\big) 
- \mbox(vec)\!\big( \boldsymbol\Sigma^* \mathbf{K}_h' \boldsymbol\Phi'\big) 
+ \mbox(vec)( \mathcal{S}_h),
\end{aligned}
\]

where $\boldsymbol\Sigma^* = (\mathbf{Id}-\boldsymbol\Phi)^{-1}\boldsymbol\Sigma (\mathbf{Id}-\boldsymbol\Phi')^{-1}$,  
$\mathbf{K}_h = (\mathbf{Id}-\boldsymbol\Phi)^{-1}(\mathbf{Id}-\boldsymbol\Phi^h)$, and

$$
\mbox(vec)(\mathcal{S}_h) = (\boldsymbol\Phi \otimes \boldsymbol\Phi)(\mathbf{Id}-\boldsymbol\Phi \otimes \boldsymbol\Phi)^{-1}
\Big[\mathbf{Id}-(\boldsymbol\Phi \otimes \boldsymbol\Phi)^{h}\Big]\mbox(vec)(\boldsymbol\Sigma^*).
$$

:::
:::{.proof}
We have

$$
\mathbf{x}_{t+h} = (\mathbf{Id} + \boldsymbol\Phi+\dots + \boldsymbol\Phi^{h-1})\boldsymbol\mu 
+ \boldsymbol\Phi^h\mathbf{x}_{t}
+ \boldsymbol\Sigma^{\tfrac{1}{2}}\boldsymbol\varepsilon_{t+h} 
+ \boldsymbol\Phi\boldsymbol\Sigma^{\tfrac{1}{2}}\boldsymbol\varepsilon_{t+h-1}
+\dots + \boldsymbol\Phi^{h-1}\boldsymbol\Sigma^{\tfrac{1}{2}}\boldsymbol\varepsilon_{t+1}.
$$

The variance of $\mathbf{x}_{t+h}+\dots+\mathbf{x}_{t+1}$, conditional on $\underline{\mathbf{x}_t}$, is

\[
\begin{aligned}
& \boldsymbol\Sigma 
+ (\mathbf{Id}+\boldsymbol\Phi)\boldsymbol\Sigma(\mathbf{Id}+\boldsymbol\Phi)' 
+ \dots 
+ (\mathbf{Id}+\boldsymbol\Phi+\dots+\boldsymbol\Phi^{h-1})\boldsymbol\Sigma(\mathbf{Id}+\boldsymbol\Phi+\dots+\boldsymbol\Phi^{h-1})' \\
&= \sum_{i=1}^{h}(\mathbf{Id}-\boldsymbol\Phi)^{-1}(\mathbf{Id}-\boldsymbol\Phi^i)\boldsymbol\Sigma 
\big[(\mathbf{Id}-\boldsymbol\Phi)^{-1}(\mathbf{Id}-\boldsymbol\Phi^i)\big]' \\
&= \sum_{i=1}^{h}(\mathbf{Id}-\boldsymbol\Phi^i)\boldsymbol\Sigma^* (\mathbf{Id}-\boldsymbol\Phi^i)' \\
&= \sum_{i=1}^{h} \boldsymbol\Sigma^* - \boldsymbol\Phi^i\boldsymbol\Sigma^* 
- \boldsymbol\Sigma^*(\boldsymbol\Phi^i)' + \boldsymbol\Phi^i\boldsymbol\Sigma (\boldsymbol\Phi^i)' \\
&= h \boldsymbol\Sigma^* 
- \boldsymbol\Phi (\mathbf{Id}-\boldsymbol\Phi)^{-1}(\mathbf{Id}-\boldsymbol\Phi^h) \boldsymbol\Sigma^* 
- \boldsymbol\Sigma^* (\mathbf{Id}-\boldsymbol\Phi^h)'  (\mathbf{Id}-\boldsymbol\Phi')^{-1} \boldsymbol\Phi' 
+ \mathcal{S}_h,
\end{aligned}
\]
which leads to the result. 
:::



:::{.theorem #FastRSpricing name="Fast computation of multi-horizon LT of a Markov chain"}
Consider an exogenous Markov chain $\mathbf{z}_t$, valued in the columns of the identity matrix of dimension $J \times J$. Its dynamics is characterized by a time-homogeneous matrix of transition probabilities denoted by $\boldsymbol\Pi$ (whose rows sum to one).

We have:
\begin{equation}
\mathbb{E}\left[\exp\left(\mathbf{u}'\{\mathbf{z}_{t+1}+\dots+\mathbf{z}_{t+h}\}\right)\left|\underline{\mathbf{z}_t}\right.\right] = \begin{pmatrix}
1&1&\dots&1
\end{pmatrix}(\mathbf{D}\boldsymbol\Pi')^h\mathbf{z}_t,(\#eq:PDz)
\end{equation}
where $\mathbf{D}=\mbox{diag}[\exp(\mathbf{u})]$.
:::
:::{.proof}
Equation \@ref(eq:PDz) is satisfied for $h=1$ since
$$
\mathbb{E}\left[\exp\left(\mathbf{u}'\mathbf{z}_{t+1}\right)\left|\underline{\mathbf{z}_t}\right.\right]=\mathbb{E}\left[\exp(\mathbf{u})'\mathbf{z}_{t+1}\left|\underline{\mathbf{z}_t}\right.\right]= \exp(\mathbf{u})'\boldsymbol\Pi'\mathbf{z}_t.
$$
Assume it is valid for a given $h>0$, we have:
\begin{eqnarray*}
&&\mathbb{E}\left[\exp\left(\mathbf{u}'\{\mathbf{z}_{t+1}+\dots+\mathbf{z}_{t+h}+\mathbf{z}_{t+h+1}\}\right)\left|\underline{\mathbf{z}_t}\right.\right]\\
&=& \mathbb{E}\left[\mathbb{E}\left[\exp\left(\mathbf{u}'\{\mathbf{z}_{t+1}+\dots+\mathbf{z}_{t+h}+\mathbf{z}_{t+h+1}\}\right)\left|\underline{\mathbf{z}_{t+1}}\right.\right]\left|\underline{\mathbf{z}_t}\right.\right]\\
&=& \mathbb{E}\left[\begin{pmatrix}
1&1&\dots&1
\end{pmatrix}(\mathbf{D}\boldsymbol\Pi')^h\mathbf{z}_{t+1}\exp(\mathbf{u}'\mathbf{z}_{t+1})\left|\underline{\mathbf{z}_t}\right.\right]\\
&=& \begin{pmatrix}
1&1&\dots&1
\end{pmatrix}(\mathbf{D}\boldsymbol\Pi')^h\mathbb{E}\left[\mathbf{z}_{t+1}\exp(\mathbf{u}'\mathbf{z}_{t+1})\left|\underline{\mathbf{z}_t}\right.\right]\\
&=& \begin{pmatrix}
1&1&\dots&1
\end{pmatrix}(\mathbf{D}\boldsymbol\Pi')^h\mbox{diag}[\exp(\mathbf{u})]\boldsymbol\Pi \mathbf{z}_t,
\end{eqnarray*}
which leads to the result.
:::

A direct implication of Theorem \@ref(thm:FastRSpricing) is:

:::{.theorem #FastRSpricing2 name="Fast pricing with Markov chains"}
We consider a scalar process $x_t$ that is such that
$$
\mathbb{E}\left[\exp\left(x_{t}\right)\left|\underline{\mathbf{z}_t},\underline{x_{t-1}}\right.\right] = \exp(\boldsymbol\alpha)'\mathbf{z}_t,
$$
where $\exp(\boldsymbol\alpha)$ is a $J$-dimensional vector whose $i^{th}$ entry is $\exp(\alpha_i)$ ($\alpha_i$ being the $i^{th}$ entry of vector $\boldsymbol\alpha$). If $\mathbf{z}_t$ is exogenous to $x_t$, we have:
$$
\mathbb{E}\left[\exp\left(x_{t+1}+\dots+x_{t+h}\right)\left|\underline{\mathbf{z}_t},\underline{x_t}\right.\right] = \begin{pmatrix}
1&1&\dots&1
\end{pmatrix}\{\mbox{diag}[\exp(\boldsymbol\alpha)]\boldsymbol\Pi'\}^h\mathbf{z}_t.
$$
:::



### Kitagawa-Hamilton inversion in a regime-switching environment

This example illustrates Kitagawa-Hamilton inversion in a regime-switching environment.

```{r fig_KH_inversion0,message=FALSE,warning=FALSE,fig.width=8,fig.height=7}
# Purpose: This example illustrates Kitagawa-Hamilton inversion in a regime-switching
# environment
library(DTAM)
library(expm)
library(mvtnorm)
library(optimx)

bond_prices_RS <- function(model,Maturities_decompo){
  # This function computes the loadings of yields on pricing factors
  #      (Markov chain and Gaussian factors).
  # Modeled yields are given by:
  # R_t = A_yds_z' * z_t + B_yds_s' + A_yds_s' * s_t
  # ----------------------------------------------------------------------------
  # Maturities_decompo defines the maturities of the yields considered.
  #      Denote by k the number of thise yields.
  # ----------------------------------------------------------------------------
  
  mu_r <- model$mu_r
  n_r <- length(mu_r)
  
  # Construct matrix of transition probabilities:
  J <- 3
  Pi <- matrix(0,J*n_r,J*n_r)
  
  p_ES <- logistic(model$alpha_ES + model$beta_ES*mu_r)
  p_TS <- logistic(model$alpha_TS + model$beta_TS*mu_r)
  p_SS <- logistic(model$alpha_SS + model$beta_SS*mu_r)
  p_ST <- logistic(model$alpha_ST + model$beta_ST*mu_r)*(1 - p_SS)
  
  p_EE <- 1-p_ES
  p_TT <- 1-p_TS
  p_SE <- 1-p_SS-p_ST
  
  p_u <- logistic(model$alpha_u + model$beta_u*mu_r)
  p_d <- logistic(model$alpha_d + model$beta_d*mu_r)
  
  # If current regime is Easing --------------------------------------------------
  Pi[1:n_r,1:n_r] <- diag(p_EE*(1-p_d))
  # Rate cut (25 bps:)
  Pi[2:n_r,1:(n_r-1)] <- Pi[2:n_r,1:(n_r-1)] + 
    diag(p_EE[2:n_r]*p_d[2:n_r]*(1-model$kappa_d))
  # Rate cut (50 bps:)
  Pi[3:n_r,1:(n_r-2)] <- Pi[3:n_r,1:(n_r-2)] + 
    diag(p_EE[3:n_r]*p_d[3:n_r]*model$kappa_d)
  # Correction for impossibility to have rate cut when the rate is low:
  Pi[1,1] <- p_EE[1]
  Pi[2,1] <- p_EE[2] - Pi[2,2]
  # Change in regime towards S:
  Pi[1:n_r,n_r + (1:n_r)] <- diag(p_ES)
  
  # If current regime is Status Quo ----------------------------------------------
  Pi[n_r + (1:n_r),1*n_r + (1:n_r)] <- diag(p_SS)
  Pi[n_r + (1:n_r),0*n_r + (1:n_r)] <- diag(p_SE)
  Pi[n_r + (1:n_r),2*n_r + (1:n_r)] <- diag(p_ST)
  
  # If current regime is Tightening ----------------------------------------------
  Pi[2*n_r + (1:n_r),2*n_r + (1:n_r)] <- diag(p_TT*(1-p_u))
  # Rate hike (25 bps:)
  Pi[2*n_r + (1:(n_r-1)),2*n_r + (2:n_r)] <- 
    Pi[2*n_r + (1:(n_r-1)),2*n_r + (2:n_r)] + 
    diag(p_TT[1:(n_r-1)]*p_u[1:(n_r-1)]*(1-model$kappa_u))
  # Rate hike (50 bps:)
  Pi[2*n_r + (1:(n_r-2)),2*n_r + (3:n_r)] <- 
    Pi[2*n_r + (1:(n_r-2)),2*n_r + (3:n_r)] + 
    diag(p_TT[1:(n_r-2)]*p_u[1:(n_r-2)]*model$kappa_u)
  # Correction for impossibility to have rate hikes when the rate is high:
  Pi[J*n_r,J*n_r] <- p_TT[n_r]
  Pi[J*n_r-1,J*n_r] <- p_TT[n_r-1] - Pi[J*n_r-1,J*n_r-1]
  # Change in regime towards S:
  Pi[2*n_r + (1:n_r),n_r + (1:n_r)] <- diag(p_TS)
  
  res_pricing <- compute_LT_RS(alpha = matrix(1,J,1) %x% matrix(-mu_r,ncol=1),Pi = Pi,
                               Maturities_decompo = Maturities_decompo,indic_add_current = TRUE)
  
  A_yds_z <- -log(res_pricing$A)/(matrix(1,J*n_r,1) %*% matrix(res_pricing$H+1,nrow=1))
  
  # Process Gaussian part:
  res_AB_Gauss <- compute_Gaussian_fast(model,Maturities_decompo)
  nb_mat <- dim(Maturities_decompo)[1]
  H <- res_AB_Gauss$H + 1
  # Add current date:
  res_AB_Gauss$PHI <- res_AB_Gauss$PHI + array(diag(2),c(2,2,nb_mat))
  A_yds_s <- matrix(NaN,2,nb_mat)
  B_yds_s <- matrix(NaN,1,nb_mat)
  for(i in 1:nb_mat){
    A_yds_s[,i] <- t(res_AB_Gauss$PHI[,,i]) %*% (model$delta1) / H[i]
    B_yds_s[i] <- (t(res_AB_Gauss$MU[,1,i]) %*% (model$delta1) -
                     t(model$delta1) %*% res_AB_Gauss$SIGMA[,,i] %*% model$delta1)/2 /H[i]
  }
  return(list(Pi = Pi,
              H = res_pricing$H,
              A_yds_z = A_yds_z,
              A_yds_s = A_yds_s,
              B_yds_s = B_yds_s))
}

filter_model_RS <- function(model,Y,Maturities_decompo,
                            vec_std_dev,indic_smoother=FALSE){
  # This function estimates the probabilities of being in the different
  #      monetary policy regimes depending on the sample Y.
  #      Note that one date (the first one) is lost due to the auto-regressive
  #      nature of s_t.
  # There are nine regimes, as the extended Markovian vector is
  #      Z_t = z_{m,t} x z_{m,t-1}.
  #      Hence, since z_{m,t} is (1) Easing, (2) Status Quo, (3) Tightening,
  #      We have: Z_t = e_1 => z_{m,t} = Easing, z_{m,t-1} = Easing.
  #               Z_t = e_2 => z_{m,t} = Easing, z_{m,t-1} = Status Quo
  #               Z_t = e_3 => z_{m,t} = Easing, z_{m,t-1} = Tightening
  #               Z_t = e_4 => z_{m,t} = Status Quo, z_{m,t-1} = Easing
  #               etc.
  # Maturities_decompo defines the maturities of the yields considered.
  #      Denote by k the number of thise yields.
  # vec_std_dev is of dimension kx1; it defines the standard deviations of the
  #      measurement errors. Some of them have to be equal
  #      to zero (number = number of Gaussian factors).
  # Y contains 1+k yields (where k is the dimension of Maturities_decompo).
  #      The first column of Y is the target rate (which defines z_{r,t})
  # vec_std_dev is a vector of 0 and 1. The zero indicate the position of the
  #      yields measured without pricing error.
  
  Y <- as.matrix(Y)
  n_r <- length(mu_r)
  TT <- dim(Y)[1]
  J <- 3
  nb_mat <- dim(Maturities_decompo)[1]
  
  mu_r <- model$mu_r
  
  res_pricing <- bond_prices_RS(model,Maturities_decompo)
  Pi <- res_pricing$Pi
  A_yds_z <- res_pricing$A_yds_z
  A_yds_s <- res_pricing$A_yds_s
  B_yds_s <- res_pricing$B_yds_s
  
  p_ES <- logistic(model$alpha_ES + model$beta_ES*mu_r)
  p_SS <- logistic(model$alpha_SS + model$beta_SS*mu_r)
  p_TS <- logistic(model$alpha_TS + model$beta_TS*mu_r)
  p_ST <- logistic(model$alpha_ST + model$beta_ST*mu_r)*(1 - p_SS)
  
  p_EE <- 1-p_ES
  p_TT <- 1-p_TS
  p_SE <- 1-p_SS-p_ST
  
  # Determine matrix of transition probabilities of z_{m,t} x z_{m,t-1} in the
  #  data (this depends on the prevailing target rate.)
  # step 1: Determine index of prevailing target rate:
  step <- model$mu_r[2] - model$mu_r[1]
  index_rates <- round(Y[,"rbar"]/step + 1)
  # step 2:
  PI <- array(0,c(3,3,TT))
  PI[1,1,] <- p_EE[index_rates]
  PI[1,2,] <- p_ES[index_rates]
  PI[2,1,] <- p_SE[index_rates]
  PI[2,2,] <- p_SS[index_rates]
  PI[2,3,] <- p_ST[index_rates]
  PI[3,2,] <- p_TS[index_rates]
  PI[3,3,] <- p_TT[index_rates]
  # Convert
  PI_9x9 <- make_Pi_z_kron_z(PI)
  
  # Compute values of Gaussian factor, depending on regime. --------------------
  # The regime variable is Z_t = z_{m,t} x z_{m,t-1}. Hence 9 regimes.
  matur_0error <- which(vec_std_dev==0)
  nb_matur_0error <- length(matur_0error)
  if(nb_matur_0error != dim(model$Phi)[1]){
    print("Inconsistency between number of yields perfectly observed and dimension of Phi.")
    return(0)
  }
  B_s_0error <- B_yds_s[1,matur_0error]
  A_s_0error <- A_yds_s[,matur_0error]
  # Determine model-implied yields, for each of the 3 regimes, on each date t:
  A_Z_t <- array(0,c(TT,nb_mat,J))
  for(j in 1:J){
    A_yds_z_aux <- A_yds_z[((j-1)*n_r+1):(j*n_r),]
    A_Z_t[,,j] <- A_yds_z_aux[index_rates,]
  }
  # For each regime, compute the values of the Gaussian factors:
  S_t <- array(0,c(TT,nb_matur_0error,J))
  A1_1 <- solve(A_s_0error)
  for(j in 1:J){
    S_t[,,j] <-   (Y[,matur_0error+1] - A_Z_t[,matur_0error,j] - 
                     t(matrix(B_s_0error,nb_matur_0error,TT))) %*% A1_1
  }
  
  # For each of the J*J regimes, compute epsilon_s (and log-likelihood): ---------------
  loglik_epsilon_s <- matrix(0,TT,J*J)
  for(zm in 1:J){
    s_t <- S_t[,,zm]
    for(zm_1 in 1:J){
      s_t_1 <- rbind(NaN,S_t[1:(TT-1),,zm_1])
      epsilon_s <- s_t - 
        t(matrix(model$mu,nb_matur_0error,TT)) - s_t_1 %*% t(model$Phi)
      loglik_epsilon_s[,(zm-1)*J+zm_1] <- 
        dmvnorm(epsilon_s,sigma = model$Sigma,log = TRUE)
    }
  }
  
  
  # For each of the J*J regimes, compute pricing errors (and log-likelihood):-----------
  
  Modeled_yields <- array(0,c(TT,J,nb_mat))
  matur_with_error <- which(vec_std_dev!=0)
  
  loglik_pricing_error <- matrix(0,TT,J*J)
  for(j in 1:J){
    modeled_yields <- t(matrix(B_yds_s,nb_mat,TT)) + S_t[,,j] %*% A_yds_s + A_Z_t[,,j]
    pricing_error <- Y[,matur_with_error+1] - modeled_yields[,matur_with_error]
    # Note in previous line: "+1" because the first column of Y is the target.
    loglik_pricing_error[,(J*(j-1)+1):(j*J)] <- 
      dmvnorm(pricing_error,
              sigma = diag(c(model$std_pricing_error*vec_std_dev[matur_with_error]^2)),
              log = TRUE)
    # Store modeled yields (for final outputs):
    Modeled_yields[,j,] <- modeled_yields
  }
  
  
  # Compute part of log-likelihood associated with dynamics of z_{r,t}: ----------------
  index_rates_1 <- c(NaN,index_rates[1:(TT-1)])
  
  # Some dates correspond to an "impossible" change in rate (larger than 50 bps),
  # proceed as if change = 50 bps:
  index_rates_modif <- index_rates
  large_positive_move <- which(index_rates-index_rates_1>2)
  index_rates_modif[large_positive_move] <- index_rates_1[large_positive_move]+2
  large_negative_move <- which(index_rates-index_rates_1< -2)
  index_rates_modif[large_negative_move] <- index_rates_1[large_negative_move]-2
  
  loglik_dynamics_z_r <- matrix(0,TT,J*J)
  z_r <- matrix(0,TT,n_r)
  z_r[index_rates]
  for(zm in 1:J){
    for(zm_1 in 1:J){
      Pi_sub <- Pi[(n_r*(zm_1-1)+1):(n_r*zm_1),(n_r*(zm-1)+1):(n_r*zm)]
      # Proba_numer is the vector of proba to have (z_{r,t},z_{m,t})
      #  given (z_{r,t-1},z_{m,t-1})
      Proba_numer <- Pi_sub[(index_rates_modif-1)*n_r + index_rates_1]
      # Prob_denom is the vector of proba to have z_{m,t} given (z_{r,t-1},z_{m,t-1})
      Pi_sub_sum <- apply(Pi_sub,1,sum)
      Proba_denom <- Pi_sub_sum[index_rates_1]
      Proba_denom[Proba_denom==0] <- 1 # Just to avoid NaN
      # Compute ratio (this is P(z_{r,t}|z_{r,t-1},z_{m,t},z_{m,t-1})):
      loglik_dynamics_z_r[,(zm-1)*J+zm_1] <- log(Proba_numer/Proba_denom)
      loglik_dynamics_z_r[is.infinite(loglik_dynamics_z_r)] <- -100000
    }
  }
  
  # ----------------------------------------------------------------------------
  # Compute resulting log-likelihood, conditional on Z_t, adding the determinant
  #   of the Jacobian matrix associated with the relationship between
  #   yields observed without errors, and Gaussian factors s_t.
  loglik_per_regime <- log(abs(det(A1_1))) + 
    loglik_epsilon_s + loglik_pricing_error + loglik_dynamics_z_r
  
  # Use the Kitagawa-Hamilton filter: ------------------------------------------
  res_KH <- KH_filter(PI_9x9,loglik_per_regime[2:TT,])
  ksi_matrix <- res_KH$ksi_matrix
  loglik_vec <- res_KH$loglik_vec
  loglik     <- res_KH$loglik
  # Probabilities associated with z_{m,t} (and not z_{m,t} x z_{m,t-1}):
  ksi_tt <- ksi_matrix %*% (diag(J) %x% matrix(1,J,1))
  
  # Compute estimates of Gaussian factors: -------------------------------------
  S_t_hat <- matrix(0,TT-1,dim(S_t)[2])
  for(j in 1:J){
    S_t_hat <- S_t_hat + S_t[2:TT,,j] * 
      (apply(ksi_matrix[,(J*(j-1)+1):(j*J)],1,sum) %*% matrix(1,1,dim(S_t)[2]))
  }
  
  # If requested, use the Kitagawa-Hamilton smoother: --------------------------
  if(indic_smoother){
    ksi_tT <- KH_smoother(PI_9x9,loglik_per_regime[2:TT,])
    # Probabilities associated with z_{m,t} (and not z_{m,t} x z_{m,t-1}):
    ksi_tT <- ksi_tT %*% (diag(J) %x% matrix(1,J,1))
  }else{
    ksi_tT <- NaN
  }
  
  return(list(A_yds_z = A_yds_z,
              A_yds_s = A_yds_s,
              B_yds_s = B_yds_s,
              A_Z_t = A_Z_t,
              Pi = Pi,
              H = res_pricing$H+1,
              PI_9x9 = PI_9x9,
              loglik_vec = loglik_vec,
              loglik = loglik,
              index_rates = index_rates,
              S_t_hat = S_t_hat,
              ksi_tT = ksi_tT,
              Modeled_yields = Modeled_yields
  ))
}


# Prepare dataset: -------------------------------------------------------------

# Compute approximate policy rate:
YC_US_weekly$rbar <- YC_US_weekly$DFEDTAR

indic_where_use_DW <- which(is.na(YC_US_weekly$rbar)&
                              (!is.na(YC_US_weekly$DFEDTARU))&
                              (!is.na(YC_US_weekly$DFEDTARL)))
YC_US_weekly$rbar[indic_where_use_DW] <- 
  0.5*YC_US_weekly$DFEDTARU[indic_where_use_DW] +
  0.5*YC_US_weekly$DFEDTARL[indic_where_use_DW]
YC_US_weekly$rbar <- round(4*YC_US_weekly$rbar)/4


Maturities_decompo <- matrix(0,6,6)
Maturities_decompo[1,] <- c(3, 0, 0, 0, 0, 0)
Maturities_decompo[2,] <- c(0, 4, 0, 0, 0, 0)
Maturities_decompo[3,] <- c(1, 0, 2, 0, 0, 0)
Maturities_decompo[4,] <- c(1, 0, 0, 2, 0, 0)
Maturities_decompo[5,] <- c(1, 0, 0, 0, 2, 0)
Maturities_decompo[6,] <- c(2, 0, 0, 0, 1, 2)

Y <- YC_US_weekly[,c("rbar","DTB3","DTB6","THREEFY1","THREEFY2","THREEFY3")]
Maturities_decompo <- Maturities_decompo[1:5,1:5]
vec_std_dev <- c(1,0,1,1,0)

Maturities_decompo <- matrix(0,4,4)
Maturities_decompo[1,] <- c(12, 0, 0, 0)
Maturities_decompo[2,] <- c(1, 2, 0, 0)
Maturities_decompo[3,] <- c(1, 0, 2, 0)
Maturities_decompo[4,] <- c(1, 0, 0, 2)

Y <- YC_US_weekly[,c("rbar","DTB3","DTB6","THREEFY1","THREEFY2")]
vec_std_dev <- c(1,0,1,0)


vec_dates <- YC_US_weekly$date[complete.cases(Y)]

Y <- Y[complete.cases(Y),]/52/100 # weekly returns
TT <- dim(Y)[1]



logistic <- function(theta){
  return(exp(theta)/(1+exp(theta)))
}
invlogistic <- function(theta){
  return(log(theta/(1-theta)))
}

rmax <- .10
step <- .0025
mu_r <- seq(0,rmax,by=step)/52 # weekly return
n_r <- length(mu_r)


# Specify initial model (used as starting point during estimation) -------------

alpha_d <- -2
beta_d <- 245
alpha_u <- -2
beta_u <- -380

alpha_ES <- -1
beta_ES <- -900
alpha_TS <- -5
beta_TS <- 2400

alpha_SS <- 2.3
beta_SS <- -420
p_SS <- logistic(alpha_SS + beta_SS*mu_r)
alpha_ST <- 1.4
beta_ST <- - 800
p_ST <- (1-p_SS)*logistic(alpha_ST + beta_ST*mu_r)
p_SE <- 1 - p_SS - p_ST

kappa_u <- .12
kappa_d <- .08

# Gaussian factors:
lambda <- .1
Phi <- diag(2)
Phi[1,1] <- .975
Phi[2,2] <- .986
Phi[1,2] <- .05
Sigma  <- diag(2)
rho <- 0.000000001
Sigma[1,2] <- rho
Sigma[2,1] <- rho

Vs <- matrix(solve(diag(4)- Phi %x% Phi) %*% c(Sigma),2,2)
std_s <- sqrt(diag(Vs))

std_dev_by_s <- .5*sd(Y[,"rbar"])

model <- list(
  alpha_d = alpha_d,
  beta_d = beta_d,
  alpha_u = alpha_u,
  beta_u = beta_u,
  alpha_ES = alpha_ES,
  beta_ES = beta_ES,
  alpha_SS = alpha_SS,
  beta_SS = beta_SS,
  alpha_TS = alpha_TS,
  beta_TS = beta_TS,
  alpha_ST = alpha_ST,
  beta_ST = beta_ST,
  kappa_u = kappa_u,
  kappa_d = kappa_d,
  mu_r = mu_r,
  mu = matrix(0,2,1),
  Phi = Phi,
  Sigma = Sigma,
  delta0 = mean(Y[,"DTB3"]-Y[,"rbar"]),
  delta1 = matrix(c(std_dev_by_s/std_s[1],0),2,1),
  std_pricing_error = .25/52/100
)


param2model <- function(param,model_ini){
  # construct matrix of transition probabilities
  
  model <- model_ini
  
  model$Phi <- diag(2)
  model$Phi[1,1] <- logistic(param[1])
  model$Phi[2,2] <- logistic(param[2])
  model$Phi[1,2] <- exp(param[3])
  
  model$alpha_d  = param[4]
  model$beta_d   = param[5]*1000
  model$alpha_u  = param[6]
  model$beta_u   = param[7]*1000
  model$alpha_ES = param[8]
  model$beta_ES  = param[9]*1000
  model$alpha_SS = param[10]
  model$beta_SS  = param[11]*1000
  model$alpha_TS = param[12]
  model$beta_TS  = param[13]*1000
  model$alpha_ST = param[14]
  model$beta_ST  = param[15]*1000
  model$kappa_u = logistic(param[16])
  model$kappa_d = logistic(param[17])
  
  model$std_pricing_error <- exp(param[18])
  
  Vs <- matrix(solve(diag(4)- model$Phi %x% model$Phi) %*% c(model$Sigma),2,2)
  std_s <- sqrt(diag(Vs))
  model$delta1 <- matrix(c(std_dev_by_s/std_s[1],0),2,1)
  
  return(model)
}

model2param <- function(model){
  
  param <- matrix(NaN,18,1)
  
  param[1]   <- invlogistic(model$Phi[1,1])
  param[2]  <- invlogistic(model$Phi[2,2])
  param[3]  <- log(model$Phi[1,2])
  
  model$alpha_d  -> param[4]
  model$beta_d/1000   -> param[5]
  model$alpha_u  -> param[6]
  model$beta_u/1000   -> param[7]
  model$alpha_ES -> param[8]
  model$beta_ES/1000  -> param[9]
  model$alpha_SS -> param[10]
  model$beta_SS/1000  -> param[11]
  model$alpha_TS -> param[12]
  model$beta_TS/1000  -> param[13]
  model$alpha_ST -> param[14]
  model$beta_ST/1000  -> param[15]
  
  invlogistic(model$kappa_u) -> param[16]
  invlogistic(model$kappa_d) -> param[17]
  
  log(model$std_pricing_error) -> param[18]
  
  Vs <- matrix(solve(diag(4)- model$Phi %x% model$Phi) %*% c(model$Sigma),2,2)
  std_s <- sqrt(diag(Vs))
  
  # invlogistic(model$delta1*std_s/(1/52/100)) -> param[17:18]
  
  return(param)
}

loglik4optim <- function(param,Y,Maturities_decompo,
                         vec_std_dev,model_ini){
  model <- param2model(param,model_ini = model_ini)
  res <- filter_model_RS(model,Y,Maturities_decompo,
                         vec_std_dev)
  
  return(-res$loglik)
}

model_ini <- model
param_est <- c(model2param(model))

if(indic_reestim){
  
  nb_loop <- 5
  
  for(i in 1:nb_loop){
    print(paste("--- Loop ",i," ---",sep=""))
    # for(method in c("nlminb")){
    for(method in c("nlminb","Nelder-Mead")){
      res.optim <- optimx(par=param_est,
                          fn=loglik4optim,
                          Y=Y,Maturities_decompo=Maturities_decompo,
                          vec_std_dev=vec_std_dev,
                          model_ini = model_ini,
                          method=method,
                          control=list(trace=TRUE,
                                       maxit=ifelse(method=="Nelder-Mead",1000,20),
                                       kkt = FALSE))
      print(paste("--- Log-lik (",method,"): ",round(res.optim$value,2)," ---",sep=""))
      param_est  <- c(as.matrix(res.optim)[1:length(param_est)])
    }
  }
  save(param_est,file= "stored_results/param_RS_inversion.Rdat")
}else{
  # In that case, use paraemters estimated previously:
  load(file= "stored_results/param_RS_inversion.Rdat")
}

# Prepare plot -----------------------------------------------------------------
# Estimated model:
model <- param2model(param_est,model_ini = model_ini)
res <- filter_model_RS(model,Y,Maturities_decompo,
                       vec_std_dev,indic_smoother = TRUE)


layout_matrix <- matrix(c(1, 1,
                          2, 3), 
                        nrow = 2, byrow = TRUE)

layout(layout_matrix)

p_E <- res$ksi_tT[,1]
p_T <- res$ksi_tT[,3]
p_S <- 1 - p_E - p_T

Modeled_yields <- res$Modeled_yields[2:TT,1,] * (matrix(p_E,ncol=1) %*% 
                                                   matrix(1,1,dim(res$Modeled_yields)[3])) +
  res$Modeled_yields[2:TT,2,] * (matrix(p_S,ncol=1) %*% 
                                   matrix(1,1,dim(res$Modeled_yields)[3])) +
  res$Modeled_yields[2:TT,3,] * (matrix(p_T,ncol=1) %*% 
                                   matrix(1,1,dim(res$Modeled_yields)[3]))

par(plt=c(.075,.95,.1,.85))
max_y <- 10
plot(vec_dates[2:TT],rep(0,TT-1),ylim=c(0,max_y),col="white",
     xlab="",ylab="",las=1,main = "(a) Target rate and monetary-policy regime",
     yaxt="n")
axis(side=2, at=seq(0,10,by=2),
     labels=paste(seq(0,10,by=2),"%",sep=""),
     col.axis="black",las=1)

polygon(c(vec_dates[2:TT],rev(vec_dates[2:TT])),max_y*c(p_E,0*rev(p_E)),col="darkgrey",
        border=NaN)
polygon(c(vec_dates[2:TT],rev(vec_dates[2:TT])),max_y*c(p_E,rev(1-p_T)),col="lightgrey",
        border=NaN)
lines(vec_dates[2:TT],52*100*Y[2:TT,"rbar"],lwd=2,ylim=c(0,max_y))

scale_factor <- max_y
axis(side=4, at=pretty(c(0,max_y)),
     labels=pretty(c(0,max_y)) / scale_factor,
     col.axis="black",las=1)

legend("topright",
       legend=c("Target rate (lhs)",
                "Proba. of Easing regime (rhs)",
                "Proba. of Status-Quo regime (rhs)",
                "Proba. of Tightening regime (rhs)"),
       lty=c(1, NA, NA, NA),       # line type
       pch=c(NA, 22, 22, 22),     # filled squares
       pt.bg = c(NA, "darkgrey", "lightgrey", "white"), # fill
       pt.cex = 1.8,
       col   = c("black", "black", "black", "black"),   # border
       lwd   = c(2,1,1,1),
       bg    = "white")

# Compute model-implied yields for the different regimes:
nb_mat <- length(vec_std_dev)
TT <- dim(Y)[1]
modeled_yields_E <- t(matrix(res$B_yds_s,nb_mat,TT-1)) + res$S_t_hat %*% res$A_yds_s + 
  res$A_Z_t[2:TT,,1]
modeled_yields_S <- t(matrix(res$B_yds_s,nb_mat,TT-1)) + res$S_t_hat %*% res$A_yds_s + 
  res$A_Z_t[2:TT,,2]
modeled_yields_T <- t(matrix(res$B_yds_s,nb_mat,TT-1)) + res$S_t_hat %*% res$A_yds_s + 
  res$A_Z_t[2:TT,,3]


par(plt=c(.15,.95,.2,.85))

count <- 0
for(t in c(580,1450)){
  count <- count + 1
  ylim <- c(
    min(modeled_yields_E[t,],modeled_yields_S[t,],modeled_yields_T[t,],
        as.matrix(Y[t,2:dim(Y)[2]])) - .15/52/100,
    max(modeled_yields_E[t,],modeled_yields_S[t,],modeled_yields_T[t,],
        as.matrix(Y[t,2:dim(Y)[2]])) + .5/52/100
  )
  
  plot(res$H*12/52,0*res$H,ylim=52*100*ylim,col="white",las=1,
       xlab="maturity (in months)",ylab="",
       main=paste("(",letters[count+1],") Yield curves on ",vec_dates[t+1],sep=""),
       xaxt="n")
  axis(side=1, at=res$H*12/52,
       labels=res$H*12/52,
       col.axis="black",las=1)
  
  lines(res$H*12/52,52*100*modeled_yields_E[t,],lwd=2,col="grey",lty=1)
  lines(res$H*12/52,52*100*modeled_yields_S[t,],lwd=2,col="grey",lty=2)
  lines(res$H*12/52,52*100*modeled_yields_T[t,],lwd=2,col="grey",lty=3)
  
  points(res$H*12/52,52*100*Modeled_yields[t,],pch=2,lwd=2,col="darkgrey")
  points(res$H*12/52,52*100*Y[t+1,2:dim(Y)[2]],pch=19)
  
  if(count == 2){
    legend("topleft",
           legend=c("Data","Model",
                    "Under Easing regime","Under Status-Quo regime","Under Tightening regime"),
           lty=c(NaN,NaN,1:3),
           pch=c(19,2,NaN,NaN,NaN),
           col=c("black","dark grey","dark grey","dark grey","dark grey"),
           lwd=2,bg     = "white")
  }
}

```

### Inversion output and state probabilities

This example shows the inversion output and state probabilities under the estimated model.

```{r fig_KH_inversion1,message=FALSE,warning=FALSE,fig.width=8,fig.height=4}
# Purpose: This example shows the inversion output and state probabilities under the estimated
# model

# Compute state-transition probabilities implied by the estimated model.
p_ES <- logistic(model$alpha_ES + model$beta_ES*mu_r)
p_SS <- logistic(model$alpha_SS + model$beta_SS*mu_r)
p_TS <- logistic(model$alpha_TS + model$beta_TS*mu_r)
p_ST <- logistic(model$alpha_ST + model$beta_ST*mu_r)*(1 - p_SS)

p_EE <- 1-p_ES
p_TT <- 1-p_TS
p_SE <- 1-p_SS-p_ST

p_u <- logistic(model$alpha_u + model$beta_u*mu_r)
p_d <- logistic(model$alpha_d + model$beta_d*mu_r)

# The first panel summarizes policy-regime changes.
# Plot regime-switching and target-rate transition probabilities.
par(mfrow=c(1,2))
par(plt=c(.25,.95,.15,.9))
plot(model$mu_r*52*100,p_ES,type="l",ylim=c(0,.75),col="white",
     xlab=expression(paste(bar(r),", in percent",sep="")),
     ylab="Proba. of change in monetary-policy regime")
grid()
lines(model$mu_r*52*100,p_ES,col="black",lwd=2,lty=1)
lines(model$mu_r*52*100,p_TS,col="black",lwd=2,lty=2)
lines(model$mu_r*52*100,p_SE,col="darkgrey",lwd=2,lty=1)
lines(model$mu_r*52*100,p_ST,col="darkgrey",lwd=2,lty=2)

legend("topleft",
       legend=c("Easing to SQ","Tightening to SQ",
                "SQ to Easing","Tightening to SQ"),
       lty=c(1,2,1,2),
       col=c("black","black","darkgrey","darkgrey"),
       lwd=2,bg     = "white")

# The second panel summarizes target-rate hikes and cuts.
plot(model$mu_r*52*100,p_ES,type="l",ylim=c(0,.75),col="white",
     xlab=expression(paste(bar(r),", in percent",sep="")),
     ylab="Proba. of change in target rate")
grid()
lines(model$mu_r*52*100,p_u,col="black",lwd=2,lty=1)
lines(model$mu_r*52*100,p_d,col="black",lwd=2,lty=2)

legend("topleft",
       legend=c(expression(paste(Delta,bar(r)," > 0 (hike)",sep="")),
                expression(paste(Delta,bar(r)," < 0 (cut)",sep=""))),
       lty=c(1,2,1,2),
       col=c("black","black","darkgrey","darkgrey"),
       lwd=2,bg     = "white")

```
