# Imperfect information and rational expectations {#ImperfectInfoRE}

<!-- ```{r ECB1, eval=FALSE} -->
<!-- library(ecb) -->
<!-- maturities <- 1:10 -->
<!-- list_of_tickers <- paste("YC.B.U2.EUR.4F.G_N_A.SV_C_YM.SR_",maturities,"Y",sep="") -->
<!-- all_yields <- data.frame() -->
<!-- count <- 0 -->
<!-- for(ticker in list_of_tickers){ -->
<!--   count <- count + 1 -->
<!--   x <- get_data(ticker) -->
<!--   x_reduced <- data.frame(date=x$obstime, -->
<!--                           yield=x$obsvalue) -->
<!--   names(x_reduced)[2] <- paste("Y",maturities[count],sep="") -->
<!--   if(count==1){ -->
<!--     all_yields <- x_reduced -->
<!--   }else{ -->
<!--     all_yields <- merge(all_yields,x_reduced,by="date") -->
<!--   } -->
<!-- } -->
<!-- ``` -->

<!-- ```{r ECB2, echo=FALSE} -->
<!-- ``` -->

## Imperfect information

This example introduces learning dynamics under imperfect information.

```{r fig_Imperfect_Info1,message=FALSE,warning=FALSE,fig.width=7,fig.height=4}
# Purpose: This example introduces learning dynamics under imperfect information
library(DTAM)

set.seed(123)

# Bansal and Yaron (2004) specification:
mu_c    <- 0 #.0015
phi   <- .979
sigma <- 0.0078
varphi_e <- .044

mu <- matrix(c(0,mu_c),2,1)
Phi <- matrix(0,2,2)
Phi[1,1] <- phi
Phi[2,1] <- 1
Sigma12 <- matrix(0,2,2)
Sigma12[1,1] <- varphi_e*sigma
Sigma12[2,2] <- sigma
A <- matrix(c(0,1),1,2)
B <- 0
Omega12 <- 0.0000001

model <- list(mu=mu,Phi=Phi,Sigma12=Sigma12,
              A=A,B=B,Omega12=Omega12,
              Lambda0=0*Phi,Lambda1=0*Phi)
res <- solve_learning(model = model)
# Should be equivalent given no Lambda0 and Lambda1:
# res <- make_stationary_filter(mu,Phi,Sigma12,A,B,Omega12)


# Check analytical formula:
# p11 <- sigma^2/2*(sqrt((1 - phi^2 - varphi_e^2)^2+4*varphi_e^2) - (1 - phi^2 - varphi_e^2))
# cbind(p11,P[1,1])
# P_check <- matrix(c(p11,0,0,0),2,2)
# Pstar_check <- Phi %*% P_check %*% t(Phi) + Sigma12 %*% t(Sigma12)

model_ww <- list(mu=res$mu_ww,Phi=res$Phi_ww,Sigma12=res$Sigma12_ww)

TT <- 200
ww <- simul_GaussianVAR(model_ww,nb.sim = TT)

dc <- ww[,1:2] %*% t(A)

stdv_w <- sqrt(diag(res$P))
lower_bound <- ww[,3] - 2*stdv_w[1]
upper_bound <- ww[,3] + 2*stdv_w[1]

# Prepare plot:

par(plt=c(.1,.95,.1,.95))

col_dc  <- "#11111177"
col_CI  <- "#77777744"
col_wt  <- "black"
col_wtt <- "black"

plot(dc,col="white",
     ylab="",xlab="",
     las=1,ylim=c(min(dc),1.0*max(dc)))
polygon(c(1:TT,TT:1),c(lower_bound,rev(upper_bound)),
        col=col_CI,border=NaN)

grid()
lines(dc,lwd=1,col=col_dc)
lines(ww[,1],lwd=2,col=col_wt)
lines(ww[,3],lty=3,lwd=3,col=col_wtt)

legend("topright",
       c(expression(paste(Delta,c[t],sep="")),
         expression(x[t]),
         expression(x[t*'|'*t]),
         expression(paste(x[t*'|'*t],", 95% conf. interv.",sep=""))),
       lwd=c(1,2,3,NaN),
       lty=c(1,1,3,NaN),
       pch=c(NaN,NaN,NaN,15),
       pt.cex=c(NaN,NaN,NaN,1.5),
       col=c(col_dc,col_wt,col_wtt,col_CI),
       seg.len = 2)

# Compare with KF:
# Y_t <- matrix(dc,ncol=1)
# nu_t <- matrix(0,TT,2)
# H <- Phi
# N <- Sigma12
# mu_t <- matrix(0,TT,1)
# G <- A
# M <- matrix(Omega12)
# Sigma_0 <- diag(2)
# rho_0 <- matrix(0,2,1)
# res_KF <- Kalman_filter(Y_t,nu_t,H,N,mu_t,G,M,Sigma_0,rho_0)
# lines(res_KF$r[,1],col="red",lwd=2,lty=3)


```

This short block computes auxiliary objects used in the imperfect-information examples that follow.

```{r results='hide'}
# Purpose: Compute the low-frequency contribution to consumption-growth variance.
# Computation of share of variance of dc accounted for by low-frequency component:

n <- length(model$mu)
Vww <- matrix(solve(diag(4*4) - model_ww$Phi %x% model_ww$Phi) %*%
                c(model_ww$Sigma12 %*% t(model_ww$Sigma12)),4,4)
print(Vww[1,1]/Vww[2,2])

```

### Filtered beliefs and true states under imperfect information

This example compares filtered beliefs and true states under imperfect information.

```{r fig_Imperfect_Info2,message=FALSE,warning=FALSE,fig.width=8,fig.height=4}
# Purpose: This example compares filtered beliefs and true states under imperfect information

H <- 60 # max horizon

# Prepare IRF in the context of imperfectly observed long-run risk component
Phi_ww <- res$Phi_ww
Sigma_ww <- res$Sigma12_ww


par(mfrow=c(1,2))
par(plt=c(.2,.95,.2,.8))


for(i in 1:2){
  
  x <- matrix(Sigma_ww[,i],ncol=1)
  IRF <- x
  for(t in 1:H){
    x <- Phi_ww %*% x
    IRF <- cbind(IRF,x)
  }
  IRF <- 100*IRF
  
  if(i==1){
    main.t <- expression(paste("(a) Shock on ",epsilon[x*','*t],sep=""))
  }else{
    main.t <- expression(paste("(b) Shock on ",epsilon[c*','*t],sep=""))
  }
  
  plot(0:H,IRF[1,],col="white",las=1,
       xlab="horizon",ylim=c(0,max(IRF[c(1,3),])),
       ylab="",
       main=main.t)
  grid()
  lines(IRF[1,],lwd=2,col=col_wt)
  # lines(IRF[2,],lwd=1,col=col_dc)
  lines(IRF[3,],lwd=3,col=col_wtt,lty=2)
  
  legend("topright",
         c(
           # expression(paste(Delta,c[t],sep="")),
           expression(x[t]),
           expression(x[t*'|'*t])),
         lwd=c(2,3),
         lty=c(1,2),
         col=c(col_wt,col_wtt),
         seg.len = 2)
  
}



```

### Imperfect information in a long-run-risk environment

This example embeds imperfect information in a long-run-risk environment.

```{r fig_Imperfect_BY, fig.width=7, fig.height=4}
# Purpose: This example embeds imperfect information in a long-run-risk environment

library(DTAM)

# ------------------------------------------------------------------------------
# Bansal and Yaron (2004): Homoskedastic case
# (using psi_GaussianVAR() for Laplace transform)
# ------------------------------------------------------------------------------

# Parameterization from Bansal and Yaron (2004):
EIS   <- 1.5
gamma <- 10
delta <- .998
mu    <- .0015
mu_d  <- .0015
phi   <- .979
beta  <- 3
sigma <- 0.0078
varphi_e <- .044
varphi_d <- 4.5

model_basic <- list(
  mu = matrix(c(0,mu,mu_d),3,1),
  Phi = matrix(c(phi,1,beta,rep(0,6)),3,3),
  Sigma = diag(sigma^2*c(varphi_e,1,varphi_d)^2),
  rho = 1/EIS,
  phi = phi,
  gamma = gamma,
  delta = delta,
  sigma = sigma,
  mu_c0 = 0,
  mu_c1 = c(0,1,0),
  n_w = 3
)


# all_EIS <- c(.5,.999,1.5,NaN)
all_EIS <- c(1,2,NaN)
all_col <- c("black","black","black","darkgrey","darkgrey","darkgrey")
all_lty <- c(1,2,3,1,2,3)
all_lwd <- c(2,2,1,2,2,1)

all_gamma <- c(seq(.10001,1.3,length.out=40),1.5,2,3)

all_maxSR <- matrix(0,2*length(all_EIS),length(all_gamma))

count_EIS <- 0

for(std_mesurement_noise in c(.0000001,100)){
  
  # Modify model to take into account imperfect observation ----------------------
  Omega12 <- diag(c(.0000001,std_mesurement_noise))
  Omega <- Omega12 %*% t(Omega12)
  A <- matrix(c(model_basic$mu_c1,c(1,0,0)),nrow=2,byrow = TRUE)
  B <- c(model_basic$mu_c0,0)
  VAR_ImperfectInfo <- 
    make_stationary_filter(model_basic$mu,
                           model_basic$Phi,
                           t(chol(model_basic$Sigma)),A,B,Omega12,nb_iter = 200)
  P     <- VAR_ImperfectInfo$P
  Pstar <- VAR_ImperfectInfo$Pstar
  K     <- VAR_ImperfectInfo$K
  Sigma_wtt <- K %*% (A %*% Pstar %*% t(A) + Omega) %*% t(K)
  
  model <- model_basic
  model$Sigma <- Sigma_wtt
  # ------------------------------------------------------------------------------
  
  model_solved <- solve_EZ_SDF(model,psi_GaussianVAR,z_bar_ini=1,
                               nb_loop_z_bar = 20,nb_loop_mu_z1 = 20)
  # Check formula (compare with analytical BY formula):
  # A1 <- (1 - 1/EIS)/(1 - model_solved$kappa1*model$phi)
  
  
  for(EIS in all_EIS){
    count_EIS <- count_EIS + 1
    count_gamma <- 0
    model$rho <- 1/EIS
    for(gamma in all_gamma){
      count_gamma <- count_gamma + 1
      model$gamma <- gamma
      if(is.na(EIS)){
        model$rho <- model$gamma
      }
      
      if((!is.na(EIS))&(EIS==1)){
        res <- solve_EZ_UnitEIS(model,psi = psi_GaussianVAR,Phi = model$Phi)
      }
      if((!is.na(EIS))&(EIS!=1)){
        res <- solve_EZ_SDF(model,psi_GaussianVAR,z_bar_ini=5,
                            nb_loop_z_bar = 20,nb_loop_mu_z1 = 20)
      }
      if(is.na(EIS)){
        res <- solve_CRRA(model,psi_GaussianVAR)
      }
      
      psi_w_2alpha <- psi_GaussianVAR(2*res$alpha,model)
      psi_w_alpha  <- psi_GaussianVAR(res$alpha,model)
      w <- res$Ew
      maxSR <- sqrt(exp(t(psi_w_2alpha$a - 2*psi_w_alpha$a) %*% w +
                          psi_w_2alpha$b - 2*psi_w_alpha$b)-1)
      all_maxSR[count_EIS,count_gamma] <- maxSR*sqrt(12)
    }
  }
}

par(plt=c(.15,.95,.2,.99))
plot(all_gamma,(all_maxSR[1,]),ylim=c(0,.2),type="l",
     lwd=all_lwd[1],las=1,col=all_col[1],lty=all_lty[1],
     xlab=expression(paste("coefficient of RRA (",gamma,")",sep="")),
     ylab="maximum Sharpe ratio (annualized)")
grid()
for(i in 2:dim(all_maxSR)[1]){
  lines(all_gamma,(all_maxSR[i,]),lwd=all_lwd[i],
        col=all_col[i],lty=all_lty[i])
}

for(i in 1:(length(all_EIS)-1)){
  j <- which.min(abs(all_gamma - 1/all_EIS[i]))
  y <- all_maxSR[length(all_EIS),j]
  points(1/all_EIS[i],y,pch=19,cex=1.4)
}

legend("topleft",
       c(expression(paste("EIS = 1",sep="")),
         expression(paste("EIS = 1.5",sep="")),
         expression(paste("EIS = 1/",gamma," (CCAPM case)",sep=""))),
       lwd=all_lwd,
       lty=all_lty,
       col=all_col,
       seg.len = 3
)

legend("topright",
       c(expression(paste("Full information",sep="")),
         expression(paste("Limited information",sep=""))),
       lwd=2,
       pch=15,pt.cex = 1.2,
       lty=c(NaN,NaN),
       col=c("black","darkgrey"),
       seg.len = 3
)


```





<!-- ```{r tablePCA, echo=FALSE} -->
<!-- nb_factors <- 4 -->
<!-- PCA <- prcomp(all_yields[,2:dim(all_yields)[2]]) -->
<!-- # Put factor loadings: -->
<!-- PCA_table <- round(t(PCA$rotation[,1:nb_factors]),2) -->
<!-- col.names <- rownames(PCA$rotation) -->
<!-- row.names <- colnames(PCA$rotation)[1:nb_factors] -->
<!-- # Add share of variance: -->
<!-- PCA_table <- cbind(PCA_table, -->
<!--                    round(cumsum(PCA$sdev^2)[1:nb_factors]/sum(PCA$sdev^2),3)) -->
<!-- colnames(PCA_table) <- c(col.names,"Var. share") -->
<!-- knitr::kable(PCA_table, -->
<!--              caption = "This is the table caption") -->
<!-- ``` -->


<!-- ```{r essai} -->
<!-- latex_table <- NULL -->
<!-- latex_table <- rbind("\\begin{tabular}{ccc}", -->
<!-- "\\toprule", -->
<!-- "Column 1 & Column 2 & Column 3 \\\\", -->
<!-- "\\midrule", -->
<!-- "A & 1 & 10 \\\\", -->
<!-- "B & 2 & 20 \\\\", -->
<!-- "C & 3 & 30 \\\\", -->
<!-- "\\bottomrule", -->
<!-- "\\end{tabular})") -->
<!-- name.of.file <- "table_param_essai" -->
<!-- latex.file <- paste(name.of.file,".txt", sep="") -->
<!-- write(latex_table, paste("",latex.file,sep="")) -->
<!-- ``` -->


### How optimistic beliefs affect asset prices and macro dynamics

This example studies how optimistic beliefs affect asset prices and macro dynamics.

```{r fig_Imperfect_Optimism, fig.width=8, fig.height=8}
# Purpose: This example studies how optimistic beliefs affect asset prices and macro dynamics

library(DTAM)

max_h <- 30

# ------------------------------------------------------------------------------
# Parameter values (set yours here)
# ------------------------------------------------------------------------------
phi       <- 0.8        # persistence of output gap
sigma_x   <- 0.01
rho       <- 0.9
xi        <- 0.4
mu_g      <- 0.02
sigma_g   <- 0.005
sigma_nu  <- 0.01

all_xi <- c(0,.3)

col_xi <- c("grey","black")
lwd_xi <- c(3,2)
types_info <- c(1,2)

# ------------------------------------------------------------------------------
# State-space objects
# ------------------------------------------------------------------------------
# State vector w_t = (g_t, x_t, x_{t-1})'
n <- 3

names_var <- c(expression(g[t]),
               expression(x[t]),
               expression(paste("log(",C[t],")",sep="")))
names_shock <- c(
  expression(epsilon[g*','*t]),
  expression(epsilon[x*','*t]),
  expression(eta[t])
)
nb_var <- length(names_var)-1
nb_shocks <- length(names_shock)

all_IRFs <- array(NaN,c(2*n,1+max_h,nb_shocks,length(all_xi)))

count_xi <- 0
for(xi in all_xi){
  
  count_xi <- count_xi + 1
  
  # Lambda_0 = 0
  Lambda0 <- matrix(0, nrow = n, ncol = n)
  
  # Lambda_1
  Lambda1 <- matrix(0, nrow = n, ncol = n)
  Lambda1[1,1] <- xi              # feedback from g_{t-1|t-1}
  
  # mu vector
  mu <- matrix(c((1 - rho)*mu_g, 0, 0), ncol = 1)
  
  # Phi matrix
  Phi <- matrix(c(
    rho - xi,   0,     0,
    0,          phi,   0,
    0,          1,     0
  ), nrow = n, byrow = TRUE)
  
  # Sigma^(1/2)
  Sigma12 <- matrix(c(
    sigma_g, 0,       
    0,       sigma_x, 
    0,       0), nrow = n, byrow = TRUE)
  
  # Full Sigma (if needed)
  Sigma <- Sigma12 %*% t(Sigma12)
  
  # ------------------------------------------------------------------------------
  # Measurement equation y_t = B + A w_t + Omega^(1/2) eta_t
  # ------------------------------------------------------------------------------
  A <- t(matrix(c(1,  1, -1,
                  1,  0,  0),n,2))       # Δc_t = g_t + x_t - x_{t-1}
  B <- matrix(0,2,1)
  Omega12 <- matrix(c(0,sigma_nu),2,1)
  Omega <- Omega12^2
  
  # ------------------------------------------------------------------------------
  # Bundle into a model list (optional)
  # ------------------------------------------------------------------------------
  model <- list(
    phi = phi,
    rho = rho,
    xi = xi,
    mu_g = mu_g,
    sigma_x = sigma_x,
    sigma_g = sigma_g,
    Lambda0 = Lambda0,
    Lambda1 = Lambda1,
    mu = mu,
    Phi = Phi,
    Sigma12 = Sigma12,
    Sigma = Sigma,
    A = A,
    B = B,
    Omega12 = Omega12,
    Omega = Omega
  )
  
  res <- solve_learning(model)
  # res <- make_stationary_filter(model$mu,model$Phi,model$Sigma12,
  #                               model$A,model$B,model$Omega12)
  
  
  for(shock in 1:nb_shocks){
    x <- c(res$Sigma12_ww[,shock])
    allx <- matrix(x,length(x),1)
    for(t in 1:max_h){
      x <- res$Phi_ww %*% x
      allx <- cbind(allx,x)
    }
    all_IRFs[,,shock,count_xi] <- allx
  }
}


par(plt=c(.15,.95,.1,.7))

par(mfrow=c(nb_var+1,nb_shocks))

for(i_variable in 1:(nb_var+1)){
  for(i_shock in 1:nb_shocks){
    if(i_variable<=nb_var){
      
      main.t <- bquote("(" * .(letters[i_shock]) * "." * .(i_variable) * ") "
                       * "Response of " * .(names_var[[i_variable]])
                       * " to shock " * .(names_shock[[i_shock]]))
      
      plot(0:max_h,rep(0,max_h+1),
           xlab="time after shock",ylab="",
           type="l",lwd=2,las=1,col="white",
           ylim=c(min(all_IRFs[i_variable,,i_shock,],
                      all_IRFs[n+i_variable,,i_shock,]),
                  max(all_IRFs[i_variable,,i_shock,],
                      all_IRFs[n+i_variable,,i_shock,])),
           main=main.t)
      grid()
      for(i_xi in 1:length(all_xi)){
        lines(0:max_h,all_IRFs[i_variable,,i_shock,i_xi],
              lty=types_info[1],
              lwd=lwd_xi[i_xi],
              col=col_xi[i_xi])
        lines(0:max_h,all_IRFs[n+i_variable,,i_shock,i_xi],
              lwd=lwd_xi[i_xi],
              lty=types_info[2],
              col=col_xi[i_xi])
      }
    }else{
      # Add effect on cumulated consumption
      vector <- matrix(c(1,1,-1,0,0,0),nrow=1)
      dc_xi1 <- vector %*% all_IRFs[,,i_shock,1]
      dc_xi2 <- vector %*% all_IRFs[,,i_shock,2]
      cum_dc_xi1 <- cumsum(dc_xi1)
      cum_dc_xi2 <- cumsum(dc_xi2)
      
      main.t <- bquote("(" * .(letters[i_shock]) * "." * .(i_variable) * ") "
                       * "Response of " * .(names_var[[i_variable]])
                       * " to shock " * .(names_shock[[i_shock]]))
      
      plot(0:max_h,rep(0,max_h+1),
           xlab="time after shock",ylab="",
           type="l",lwd=2,las=1,col="white",
           ylim=c(min(cum_dc_xi1,cum_dc_xi2),
                  max(cum_dc_xi1,cum_dc_xi2)),
           main=main.t)
      grid()
      lines(0:max_h,cum_dc_xi1,
            lty=1,
            lwd=lwd_xi[1],
            col=col_xi[1])
      lines(0:max_h,cum_dc_xi2,
            lty=1,
            lwd=lwd_xi[2],
            col=col_xi[2])
    }
    
    legend_labels <- sapply(all_xi, function(x)
      bquote(xi == .(round(x, 2))))
    
    if((i_shock==1)&(i_variable==1)){
      legend("topright",
             legend = c(legend_labels,
                        expression(paste("True variable (",w[t],")",sep="")),
                        expression(paste("Filtered variable (",w[t*'|'*t],")",sep=""))),
             lwd=c(lwd_xi,1,1),
             lty=c(1,1,types_info),
             col=c(col_xi,"black","black"),
             seg.len = 2)
    }
  }
}



```

## Rational Expectations


This example presents the benchmark BCM imperfect-information term-structure model.

```{r fig_BCM, fig.width=9, fig.height=8}
# Purpose: This example presents the benchmark BCM imperfect-information term-structure model

# Model of Bekaert, Cho, and Moreno (2010, JMCB): ------------------------------

# Parameter values from Table 1
zeta   <- 0.611     # δ in BCM
kappa  <- 0.064     # κ 
gamma  <- 3.156     # utility curvature
mu     <- 0.424     # µ
rho    <- 0.723     # ρ (policy smoothing)
beta_pi<- 1.525     # β (in policy rule, response to inflation)
beta_y <- 0.001     # β (in policy rule, response to ygap)
phi_yn <- 0.958
omega  <- 0.877     # ω
phi    <- 0.134
phi_1  <- 0.500     # φ₁
phi_2  <- 0.470     # φ₂
phi_3  <- 0.010 #1 - phi_1 - phi_2  # φ₃ = 1−φ₁−φ₂ ≈ 0.008
sigma_AS  <- 1.249/100
sigma_IS  <- 0.671/100
sigma_MP  <- 2.177/100
sigma_yn  <- 1.380/100
sigma_pistar <- 0.730/100

alpha_IS <- 0.01
r_bar <- .03
alpha_MP <- (1-rho)*r_bar

n <- 5

# Build matrices
Gamma0 <- matrix(0, 5, 5)
Gamma0[1, ] <- c(1, -kappa, 0,  kappa, 0)
Gamma0[2, ] <- c(0,    1,  phi, 0,     0)
Gamma0[3, ] <- c(0, -(1-rho)*beta_y, 1, (1-rho)*beta_y, (1-rho)*beta_pi)
Gamma0[4, ] <- c(0, 0, 0, 1, 0)
Gamma0[5, ] <- c(-phi_3, 0, 0, 0, 1)

Gamma1 <- matrix(0, 5, 5)
Gamma1[1,1] <- 1 - zeta
Gamma1[2,2] <- 1 - mu
Gamma1[3,3] <- rho
Gamma1[4,4] <- phi_yn      # your varphi_{y^n}
Gamma1[5,5] <- phi_2

C <- matrix(c(0, alpha_IS, alpha_MP, 0, 0), 5, 1)

Psi0 <- matrix(0, 5, 5)
Psi0[1,1] <- zeta
Psi0[2,1] <- phi
Psi0[2,2] <- mu
Psi0[3,1] <- (1-rho)*beta_pi
Psi0[5,5] <- phi_1

Psi1 <- 0*Psi0

Pi <- diag(c(sigma_AS, sigma_IS, sigma_MP, sigma_yn, sigma_pistar))

# Solve
# Your solver:
model <- list(mu = solve(Gamma0) %*% C,
              Phi = solve(Gamma0) %*% Gamma1,
              Psi0 = solve(Gamma0) %*% Psi0,
              Psi1 = solve(Gamma0) %*% Psi1,
              Sigma12 = solve(Gamma0) %*% Pi)
res <- solve_RE(model)

# Unconditional mean of w:
Ew <- solve(diag(n) - res$Phi_tilde) %*% res$mu_tilde

# Prepare IRFs: -----------

max_h <- 40

nb_shocks <- dim(Pi)[2]

par(mfrow=c(n,nb_shocks))
par(plt=c(.33,.98,.15,.8))

names_var <- c(expression(pi[t]),
               expression(y[t]),
               expression(r[t]),
               expression(y[t]^n),
               expression(pi[t]^'*'))
names_shock <- c(expression(epsilon[AS*','*t]),
                 expression(epsilon[IS*','*t]),
                 expression(epsilon[MP*','*t]),
                 expression(epsilon[y^n*','*t]),
                 expression(epsilon[pi^"*"*','*t]))

ylab_IRF <- function(i_var){
  bquote("Response of " * .(names_var[[i_var]]) * "...")
}

title_IRF <- function(i_shock) {
  bquote("... to " * .(names_shock[[i_shock]]))
}

for(shock in 1:nb_shocks){
  x <- c(res$Sigma12_tilde[,shock])
  allx <- matrix(x,n,1)
  for(t in 1:max_h){
    x <- res$Phi_tilde %*% x
    allx <- cbind(allx,x)
  }
  for(i in 1:n){
    par(mfg=c(i,shock))
    
    ylab_val <- if (shock == 1) ylab_IRF(i) else ""
    main_val <- if (i == 1) title_IRF(shock) else ""
    
    plot(0:max_h,100*allx[i,],type="l",lwd=2,las=1,col="white",
         xlab="",ylab=ylab_val,main=main_val,
         cex.lab = 1.4,cex.main=1.4)
    grid()
    lines(0:max_h,100*allx[i,],lwd=2)
  }
}


```


### Additional implications of the BCM imperfect-information model

This example reports additional implications of the BCM imperfect-information model.

```{r fig_BCM2, fig.width=9, fig.height=6}
# Purpose: This example reports additional implications of the BCM imperfect-information model

# Build the risk-neutral dynamics implied by the SDF loading.
lambda0 <- solve(res$Sigma12_tilde) %*%
  matrix(c(0, sigma_IS/phi, 0, 0, 0), n, 1) -
  matrix(c(1, gamma, 0, 0, 0), n, 1)

model_P <- list(mu = res$mu_tilde,
                Phi = res$Phi_tilde,
                Sigma = res$Sigma12_tilde %*% t(res$Sigma12_tilde))
model_Q <- model_P
model_Q$mu <- res$mu_tilde + res$Sigma12_tilde %*% lambda0

resAB <- compute_AB_classical(xi0=0,xi1=matrix(c(0,0,1,0,0),n,1),
                              H=40, # maximum maturity
                              psi=psi_GaussianVAR,
                              psi.parameterization=model_Q)

# Compute level, slope, and curvature loadings.
loadings <- matrix(NaN,3,n)
loadings[1,] <- resAB$a[,1,c(1,4,20)] %*% matrix(rep(1/3,3),3,1)
loadings[2,] <- resAB$a[,1,20] - resAB$a[,1,1] 
loadings[3,] <- resAB$a[,1,20] + resAB$a[,1,1] - 2*resAB$a[,1,4]

names_var <- c(expression(paste("level",sep="")),
               expression(paste("slope",sep="")),
               expression(paste("curvature",sep="")))

ylab_IRF <- function(i_var){
  bquote("Response of " * .(names_var[[i_var]]) * "...")
}

title_IRF <- function(i_shock) {
  bquote("... to " * .(names_shock[[i_shock]]))
}

par(mfrow=c(length(names_var),nb_shocks))
par(plt=c(.33,.98,.15,.8))

# Plot impulse responses of yield-curve factors to each structural shock.
for(shock in 1:nb_shocks){
  x <- c(res$Sigma12_tilde[,shock])
  allx <- matrix(x,n,1)
  for(t in 1:max_h){
    x <- res$Phi_tilde %*% x
    allx <- cbind(allx,x)
  }
  allx <- loadings %*% allx
  
  for(i in 1:dim(allx)[1]){
    par(mfg=c(i,shock))
    
    ylab_val <- if (shock == 1) ylab_IRF(i) else ""
    main_val <- if (i == 1) title_IRF(shock) else ""
    
    plot(0:max_h,100*allx[i,],type="l",lwd=2,las=1,col="white",
         xlab="",ylab=ylab_val,main=main_val,
         cex.lab = 1.4,cex.main=1.4)
    grid()
    lines(0:max_h,100*allx[i,],lwd=2)
  }
}


```

### Rational Expectation framework with longer horizon

We consider a system in which the $n$–dimensional state vector $\mathbf{w}_t$ evolves according to
\[
\mathbf{w}_t
= \boldsymbol\mu
+ \boldsymbol\Phi\,\mathbf{w}_{t-1}
+ \boldsymbol\Psi_0\,\mathbb{E}(\mathbf{w}_{t+2}\mid\underline{\mathbf{w}_t})
+ \boldsymbol\xi_t,
\qquad
\boldsymbol\xi_t \sim i.i.d.\;\mathcal{N}(\mathbf{0},\boldsymbol\Sigma_\xi).
\]

There are no one–period–ahead expectations and no contemporaneous expectational terms:
\[
\boldsymbol\Psi_1 = \mathbf{0}, \qquad
\mathbb{E}(\mathbf{w}_{t+1}\mid\underline{\mathbf{w}_t}) \text{ does not enter the equation.}
\]

To cast the problem into the standard first-order form, we introduce the auxiliary forward-looking variable
\[
\widetilde{\mathbf{w}}_t = \mathbb{E}(\mathbf{w}_{t+1}\mid\underline{\mathbf{w}_t}),
\]
and define the augmented state vector
\[
\mathbf{W}_t = (\mathbf{w}_t',\,\widetilde{\mathbf{w}}_t')'.
\]

Using the law of iterated expectations,
\[
\mathbb{E}(\widetilde{\mathbf{w}}_{t+1}\mid\underline{\mathbf{w}_t})
= \mathbb{E}(\mathbf{w}_{t+2}\mid\underline{\mathbf{w}_t}),
\]
the system becomes
\[
\mathbf{W}_t
=
\begin{pmatrix}
\boldsymbol\mu \\
\mathbf{0}
\end{pmatrix}
+
\begin{pmatrix}
\boldsymbol\Phi & \mathbf{0} \\
\mathbf{0} & \mathbf{0}
\end{pmatrix}
\mathbf{W}_{t-1}
+
\begin{pmatrix}
\mathbf{0} & \boldsymbol\Psi_0 \\
\mathbf{Id}_n & \mathbf{0}
\end{pmatrix}
\mathbb{E}(\mathbf{W}_{t+1}\mid\underline{\mathbf{W}_t})
+
\begin{pmatrix}
\boldsymbol\xi_t \\
\mathbf{0}
\end{pmatrix}.
\]

This representation has a standard structure involving conditional expectations involving a single horizon.


This block computes a two-period-ahead rational-expectations benchmark for comparison.

```{r RE2periodsahead}
# Purpose: Build the two-period-ahead rational-expectations benchmark.

# ==========================
# Original model dimensions
# ==========================
n  <- 2               # dimension of w_t
mu <- matrix(c(0,0), n, 1)

Phi  <- matrix(c(0.5,0.1,
                 0.0,0.6), n, n)

Psi0 <- matrix(c(0.2,0,
                 0,  0.1), n, n)

Sigma12 <- diag(n)

# ==========================
# Augment the system
# ==========================
# Augmented state dimension: 2n
N <- 2 * n

# Augmented mũ
mu_tilde <- rbind(mu, matrix(0, n, 1))

# Augmented Phĩ
Phi_tilde <- rbind(
  cbind(Phi, matrix(0, n, n)),
  cbind(matrix(0, n, n), matrix(0, n, n))
)

# Augmented Ψ̃0
Psi0_tilde <- rbind(
  cbind(matrix(0, n, n), Psi0),
  cbind(diag(n),         matrix(0, n, n))
)

# No Ψ1 term
Psi1_tilde <- matrix(0, N, N)

# Shock matrix Σ̃½ = [Σ½; 0]
Sigma12_tilde <- rbind(Sigma12, matrix(0, n, n))

# Build model list for solve_RE
model_aug <- list(
  mu      = mu_tilde,
  Phi     = Phi_tilde,
  Psi0    = Psi0_tilde,
  Psi1    = Psi1_tilde,
  Sigma12 = Sigma12_tilde
)

# Solve
sol <- solve_RE(model_aug)

# Check

sim.w <- simul_GaussianVAR(sol,nb.sim=100)
E.w <- sim.w %*% t(sol$Phi_tilde)

plot(sim.w[,1],type="l")
lines(sim.w[,3],col="darkgrey")
lines(E.w[,1],col="black",lty=2)

```




### The imperfect-information framework to an Ireland-style New Keynesian model

This example applies the imperfect-information framework to an Ireland-style New Keynesian model.

```{r fig_Ireland, fig.width=10, fig.height=10}
# Purpose: This example applies the imperfect-information framework to an Ireland-style New
# Keynesian model

# Ireland (2004) --------------------------------------------------------------
library(geigen)
library(DTAM)



# R code for Ireland (2004) system:
# w_t = (a_t, e_t, g_t, r_t, x_t, pi_t)'
# Γ0 w_t = Γ1 w_{t-1} + Ψ0 E_t w_{t+1} + Π ε_t



# Parameters (symbolic placeholders) -------------------------------------------

rho_a   <- 0.9470
rho_e   <- 0.9625
rho_g   <- 0.2536
rho_x   <- 0.0347
rho_pi  <- 0.3597

omega    <- 0.0617
alpha_x  <- 0.0836
alpha_pi <- 0.0000

psi      <- 0.1
beta     <- 0.99

sigma_a  <- 0.0405
sigma_e  <- 0.0012
sigma_z  <- 0.0109
sigma_r  <- 0.0031

# --------- BUILD MATRICES ------------------------------------------------------------

n <- 6

Gamma0 <- matrix(0, n, n)
Gamma0[1, ] <- c(1,                0,              0,     0,       0,      0)
Gamma0[2, ] <- c(0,                1,              0,     0,       0,      0)
Gamma0[3, ] <- c(-omega,           0,              1,     0,      -1,      0)
Gamma0[4, ] <- c(0,                0,         -rho_g,     1,  -rho_x, -rho_pi)
Gamma0[5, ] <- c(-(1-omega)*(1-rho_a), 0,      0,     1,       1,      0)
Gamma0[6, ] <- c(0,                1,              0,     0,   -psi,     1)

Gamma1 <- matrix(0, n, n)
Gamma1[1, ] <- c(rho_a, 0,       0,   0,         0,           0)
Gamma1[2, ] <- c(0,     rho_e,   0,   0,         0,           0)
Gamma1[3, ] <- c(-omega, 0,       0,   0,       -1,           0)
Gamma1[4, ] <- c(0,     0,       0,   1,         0,           0)
Gamma1[5, ] <- c(0,     0,       0,   0,    alpha_x,        0)
Gamma1[6, ] <- c(0,     0,       0,   0,         0,  beta*alpha_pi)

Psi0 <- matrix(0, n, n)
Psi0[5, ] <- c(0, 0, 0, 0, (1-alpha_x), 1)
Psi0[6, ] <- c(0, 0, 0, 0, 0, beta*(1-alpha_pi))

Psi1 <- 0*Psi0

Pi <- matrix(0, n, 4)
Pi[1,1] <- sigma_a
Pi[2,2] <- sigma_e
Pi[3,3] <- sigma_z
Pi[4,4] <- sigma_r
# rows 5–6 remain zero

C <- matrix(0, 6, 1)

# Return as list
model_matrices <- list(
  Gamma0 = Gamma0,
  Gamma1 = Gamma1,
  Psi0   = Psi0,
  C      = C,
  Pi     = Pi
)

# Solve
# Your solver:
model <- list(mu = solve(Gamma0) %*% C,
              Phi = solve(Gamma0) %*% Gamma1,
              Psi0 = solve(Gamma0) %*% Psi0,
              Psi1 = solve(Gamma0) %*% Psi1,
              Sigma12 = solve(Gamma0) %*% Pi)
res <- solve_RE(model)


# Prepare IRFs:

# Check
# Gamma0 %*% res$Phi_tilde - Gamma1 - Psi0 %*% res$Phi_tilde %*% res$Phi_tilde


max_h <- 40

nb_shocks <- dim(Pi)[2]

par(mfrow=c(nb_shocks,n))

par(plt=c(.15,.95,.1,.7))

names_var <- c("a","e","g","r","x","pi")
names_shock <- c("a","e","z","r")

for(shock in 1:nb_shocks){
  x <- c(res$Sigma12_tilde[,shock])
  allx <- matrix(x,n,1)
  for(t in 1:max_h){
    x <- res$Phi_tilde %*% x
    allx <- cbind(allx,x)
  }
  for(i in 1:n){
    plot(0:max_h,allx[i,],type="l",lwd=2,
         main=paste("Resp. of ",names_var[i],
                    " to shock ",names_shock[shock],sep=""))
  }
}

```

## Imperfect information and rational expectations

This example studies another imperfect-information New Keynesian specification and its term-structure implications.

```{r fig_CGG1, fig.width=10, fig.height=10}
# Purpose: This example studies another imperfect-information New Keynesian specification and
# its term-structure implications



library(DTAM)

max_h <- 30

col_info <- c("grey","black")
lwd_info <- c(4,2)
types_t_or_tt <- c(1,2)

# --- Parameters ------------------------------------------------------------
zeta       <- .5
kappa      <- .2  # κ
gamma      <- 3   # γ
rho        <- .8  # ρ
phi_x      <- .2  # φ_x
phi_pi     <- 1.5 # φ_π

rho_rstar  <- .95 # ρ_r_star
rho_pistar <- .95 # ρ_pi_star

rbar_star  <- .02 # r_bar_star
pibar_star <- .02 # pi_bar_star

sigma_AS   <- .01  # σ_AS
sigma_IS   <- .01 # σ_IS
sigma_MP   <- .01 # σ_MP
sigma_rstar  <- .01 * sqrt(1-rho_rstar^2) # σ_r_star
sigma_pistar <- .01 * sqrt(1-rho_pistar^2) # σ_pi_star

sigma_nu   <- .02

# --- Define Θ ------------------------------------------------------------

Theta <- rbind(
  c(1, -kappa,                0,                 0,                     0),
  c(0,  1,        1/gamma,        -1/gamma,       0),
  c(0, -(1-rho)*phi_x, 1, -(1-rho), -(1-rho)*(1-phi_pi)),
  c(0, 0, 0, 1, 0),
  c(0, 0, 0, 0, 1)
)

Theta_inv <- solve(Theta)

# --- μ ------------------------------------------------------------

mu_vec <- c(0,
            0,
            0,
            (1 - rho_rstar) * rbar_star,
            (1 - rho_pistar) * pibar_star)

mu <- Theta_inv %*% matrix(mu_vec, ncol = 1)

# --- Φ ------------------------------------------------------------

Phi_raw <- rbind(
  c(zeta,0,0,0,0),
  c(0,0,0,0,0),
  c(0,0,rho,0,0),
  c(0,0,0,rho_rstar,0),
  c(0,0,0,0,rho_pistar)
)

Phi <- Theta_inv %*% Phi_raw

# --- Γ₀ ------------------------------------------------------------

Gamma0_raw <- rbind(
  c((1-zeta), 0, 0, 0, 0),
  c(1/gamma, 1, 0, 0, 0),
  c(0, 0, 0, 0, 0),
  c(0, 0, 0, 0, 0),
  c(0, 0, 0, 0, 0)
)

Gamma0 <- Theta_inv %*% Gamma0_raw

# --- Ψ₀ ------------------------------------------------------------

Psi0_raw <- rbind(
  c(0,0,0,0,0),
  c(0,0,0,0,0),
  c((1-rho)*phi_pi, 0, 0, 0, 0),
  c(0,0,0,0,0),
  c(0,0,0,0,0)
)

Psi0 <- Theta_inv %*% Psi0_raw

# --- Σ^{1/2} ------------------------------------------------------------

Sigma_half_raw <- diag(c(sigma_AS,
                         sigma_IS,
                         sigma_MP,
                         sigma_rstar,
                         sigma_pistar))

Sigma_half <- Theta_inv %*% Sigma_half_raw

# Lambda0, Lambda1, Psi1, and Gamma1 are zero matrices -------------------------

Lambda0 <- matrix(0, 5, 5)
Lambda1 <- matrix(0, 5, 5)
Psi1    <- matrix(0, 5, 5)
Gamma1  <- matrix(0, 5, 5)

# --- Measurement-equation matrices for imperfect information
A <- matrix(c(1,0,0,0,0,
              0,1,0,0,0,
              0,0,1,0,0),
            nrow = 3, byrow = TRUE)

B <- matrix(0, nrow = 3, ncol = 1)

Omega12 <- matrix(0, nrow = 3, ncol = 1)
Omega12[2] <- sigma_nu

# --- Final output ------------------------------------------------------------

# Specify model with imperfect information and rational expectations (IIRE):
model_IIRE <- list(
  Theta        = Theta,
  mu           = mu,
  Phi          = Phi,
  Gamma0       = Gamma0,
  Psi0         = Psi0,
  Sigma12      = Sigma_half,
  Lambda0      = Lambda0,
  Lambda1      = Lambda1,
  Psi1         = Psi1,
  Gamma1       = Gamma1,
  Omega12      = Omega12,
  A            = A,
  B            = B)

n <- dim(Phi)[1]
nb_shocks <- n

names_var <- c(expression(pi[t]),
               expression(x[t]),
               expression(r[t]),
               expression(r[t]*'*'),
               expression(pi[t]*'*'))
names_shock <- c(
  expression(epsilon[AS*','*t]),
  expression(epsilon[IS*','*t]),
  expression(epsilon[MP*','*t]),
  expression(epsilon[r*'*'*','*t]),
  expression(epsilon[pi*'*'*','*t]))


all_indic_FI <- c(TRUE,FALSE)

all_IRFs <- array(NaN,c(2*n,1+max_h,nb_shocks,length(all_indic_FI)))

count_info_type <- 0
for(indic_FI in all_indic_FI){
  count_info_type <- count_info_type + 1
  
  if(!indic_FI){
    # Imperfect information
    model <- model_IIRE
    model$Psi0 <- model$Psi0 + model$Gamma0
    model$Gamma0 <- 0*model$Gamma0
  }else{
    # Full information
    model <- model_IIRE
    model$A <- diag(n)
    model$B <- matrix(0,n,1)
    model$Omega12 <- matrix(0,n,1)
  }
  
  res <- solve_Learning_RE(model)
  
  for(shock in 1:nb_shocks){
    x <- c(res$Sigma12_ww[,shock])
    allx <- matrix(x,length(x),1)
    for(t in 1:max_h){
      x <- res$Phi_ww %*% x
      allx <- cbind(allx,x)
    }
    all_IRFs[,,shock,count_info_type] <- allx
  }
}

all_IRFs <- 100*all_IRFs


par(plt=c(.22,1,.15,.75))

par(mfrow=c(n,nb_shocks))

for(i_variable in 1:n){
  for(i_shock in 1:nb_shocks){
    
    main.t <- bquote("(" * .(letters[i_shock]) * "." * .(i_variable) * ") "
                     * "Resp. of " * .(names_var[[i_variable]])
                     * " to " * .(names_shock[[i_shock]]))
    
    ylim <- c(min(all_IRFs[i_variable,,i_shock,],
                  all_IRFs[n+i_variable,,i_shock,]),
              max(all_IRFs[i_variable,,i_shock,],
                  all_IRFs[n+i_variable,,i_shock,]))
    ylim[1] <- min(ylim[1],-.001)
    ylim[2] <- max(ylim[2],.001)
    
    plot(0:max_h,rep(0,max_h+1),
         xlab="time after shock",ylab="",
         type="l",lwd=2,las=1,col="white",
         ylim=ylim,
         main=main.t,cex.main=1.3)
    grid()
    for(i_indic_FI in 1:length(all_indic_FI)){
      lines(0:max_h,all_IRFs[i_variable,,i_shock,i_indic_FI],
            lty=types_t_or_tt[1],
            lwd=lwd_info[i_indic_FI],
            col=col_info[i_indic_FI])
      lines(0:max_h,all_IRFs[n+i_variable,,i_shock,i_indic_FI],
            lwd=lwd_info[i_indic_FI],
            lty=types_t_or_tt[2],
            col=col_info[i_indic_FI])
    }
    
    if((i_shock==1)&(i_variable==1)){
      legend("topright",
             legend = c(expression(paste("Full Info. (",
                                         w[t]," = ",w[t*'|'*t],")",sep="")),
                        expression(paste("Imperfect info., ",
                                         w[t],sep="")),
                        expression(paste("Imperfect info., ",
                                         w[t*'|'*t],sep=""))),
             lwd=c(lwd_info[1],lwd_info[2],lwd_info[2]),
             lty=c(1,types_t_or_tt[1],types_t_or_tt[2]),
             col=c(col_info[1],col_info[2],col_info[2]),
             seg.len = 2)
    }
  }
}


```
